### 2021-09-06 author-title-etc.tex
Update on Overleaf.

%!TEX root = main.tex

%%%%% THIS IS THE SECTION WHERE THE AUTHOR PUTS IN ALL OF THEIR TITLE AND AFFILIATION
%%%%% INFORMATION AND  A FEW OTHER THINGS
\newcommand{\myTitle}{{\sc gscramble:} Simulation of admixed individuals without reuse of genetic material}
\title{\myTitle}

% redefine this to make author list. Note affiliation symbols are done manually.
% All caps tends to look better
\newcommand{\myAuthors}{Eric C. Anderson$^{*,\S}$, Rachel M. Giglio$^{\dagger}$, Timothy J. Smyser$^{\dagger}$}
\author{Eric C. Anderson$^{*,\S}$, Rachel M. Giglio$^{\dagger}$, Timothy J. Smyser$^{\dagger}$}

% redefine to make the affiliation list.  Note symbols are done manually
\newcommand{\myAffiliations}{$^*$Fisheries Ecology Division,
    Southwest Fisheries Science Center, National Marine Fisheries Service, NOAA,
    110 McAllister Road,
    Santa Cruz, CA 95060, USA.,
    $^\dagger$  ADDRESS FOR APHIS/NWRC}
\renewcommand{\AuthorAddresses}{\myAffiliations}

\renewcommand{\KeyWords}{genetic stock identification, mixture deconvolution, permutation testing, power analysis, blah, blah}

\renewcommand{\CorrespondingAuthor}{Eric C. Anderson, eric.anderson@noaa.gov}

% the email address for the corresponding author
\newcommand{\myEmailAddress}{eric.anderson@noaa.gov}
\newcommand{\myEmailFootnote}{$^\S$}

% here you can put your very own copyright notice
\newcommand{\myCopyright}{\copyright US Federal Government work in the public domain in the USA}

% here you can put a running title (a short title that goes on the left of the
% even pages)
\newcommand{\myRunningTitle}{Linkage-aware power assessment for admixture proportions}
\renewcommand{\RunningTitle}{\myRunningTitle}

% and here you put the running author (short listing of authors for the
% upper right header on the odd pages)
\newcommand{\myRunningAuthor}{Anderson {\em et al.}}

%%%% DONE WITH AUTHOR/TITLE/ETC INFORMATION DEFINITION

### 2021-09-06 body-text-of-paper.tex
Update on Overleaf.

%!TEX root = main.tex

\section*{Introduction}

In just a few years, next-generation sequencing (NGS) technologies have revolutionized the study of evolution and ecology in both model and non-model organisms, and have become established as standard tools in molecular ecology.

Not really what we will say here.

Don't forget to talk about Anthony Almudevar's visionary method for GSI power assessement \citep{almudevar2000exact}

\section*{Methods}

Something here.

\subsection*{Genomic Simulation Pedigrees}

For the purposes here, we define blah, blah.

\subsection*{Another subsection}

\subsection*{And a third}

Blah, blah blobbity blah.

\begin{figure*}
\begin{center}
\includegraphics[width=0.8\textwidth]{images/gsp4-700.pdf}
\end{center}
\caption[]{Stick this thing in somewhere}
\end{figure*}

Not visible in the static plots are a number of dynamic features in the Shiny app display.
Values of filtering criteria for a single locus can be changed in real time in the app, and the results are updated in the adjacent genotype biplot.
Furthermore, it is possible to highlight points (representing individual samples) in the plot for one locus, and then observe the placement of those same selected samples at other loci.
This is particularly well-suited to identifying samples with aberrant genotypes across multiple loci---a good indicator of contamination or other issues with the sample.

\section*{Discussion}
Hey! Let's discuss this!

\section*{Acknowledgements}
We thank the members of the Molecular Ecology and Genetic Analysis (MEGA) team at NOAA SWFSC's Fisheries Ecology Division, etc. etc. .

### 2021-09-06 data-accessibility.tex
Update on Overleaf.

\section*{Data Accessibility}

{\sc microhaplot} is available as an R package with these associated links:
\begin{itemize}
\item Stable version available on CRAN (when available): \url{https://CRAN.R-project.org/package=gscramble}
\item Development version and entire revision history on GitHub: \url{https://github.com/eriqande/gscramble}
\item Online version of all package documentation: \url{https://eriqande.github.io/gscramble}
\item Data repository with full data from pig, Chinook, and bird (maybe?) data sets \url{https://dryad.something.or.other}
\end{itemize}
\mbox{}

### 2023-10-17 body-text-of-paper.tex
Clean up to make it a GitHub repo

@@
 
-In just a few years, next-generation sequencing (NGS) technologies have revolutionized the study of evolution and ecology in both model and non-model organisms, and have become established as standard tools in molecular ecology.
+Genetic clustering algorithms are a commonly used approach to describe genetic structure among populations, reflecting the cumulative effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.
+The capacity of such algorithms to differentiate populations is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker set and sample sizes from the respective populations.
+Based on an initial description of genetic structure, ecologists are frequently interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location and genetic origins of a given animal \citep{paetkau1995microsatellite,wilson2003bayesian}.
+Describing such patterns of movement have many implications, influencing evolutionary processes, disease dynamics \citep{huestis2019windborne}, or patterns of invasive species expansion.
 
-Not really what we will say here.
+	In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals with complex ancestry patterns (i.e., the proportion of an individual’s genome assigned to sampled populations, sensu \citealt{pritchard2000inference}) that demonstrate mixed associations to multiple populations.
+Various processes could contribute to such mixed associations.
+Most simplistically, individuals may represent the influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately representing contributions from multiple populations included in the study.
+However, alternative processes could also create similarly complex ancestry patterns.
+For example, individuals may have genomic associations to unsampled populations as immigrants or as descendant of immigrant from unsampled populations \citep{pritchard2000inference,beerli2004effect}.
+Yet a limitation of clustering algorithms is that exogenous immigrants can only be associated to populations represented within the sample.
+Similarly, individuals with mixed associations may reflect a discrepancy between the genetic structure as described by the analysis and the underlying ecological process creating patterns of genetic differentiation (e.g., when clinal systems governed by isolation by distance are described as discrete genetic clusters).
+Finally, such patterns may represent a limitation of the statistical power of the genetic data, as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure.
+The challenge for researchers then becomes correctly identifying the ecological process by which similar, complex ancestry patterns were created.
 
-Don't forget to talk about Anthony Almudevar's visionary method for GSI power assessement \citep{almudevar2000exact}
+Pairing empirical data with closely derived simulation-based approaches has been long recognized as a tool to help guide the interpretation of empirical data and define the limitations of the available data \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
+Mirroring empirical datasets with forward-in-time simulations often implement sampling with replacement (e.g.
+citations).
+However, sampling with replacement implicitly increases the pairwise genetic distance (i.e., FST) among populations and inflates the perceived ability to correctly assign individuals back to source populations \cite{almudevar2000exact,anderson2008improved}.
+\textcolor{red}{Seems like 2 additional sentences here would be appropriate to draw from a study that has demonstrated this process.}  The effects of overestimating our ability to resolve populations creates individuals that discretely align with simulated populations.
+These issues could be addressed with simulations that fix observed genetic differences among populations by sampling from the empirical data without replacement (citations for simulation frameworks that use this approach).
+  Beyond simply evaluating the resolution of discrete populations in a simulated framework, researchers may also be interested in how individuals of admixed ancestry are genetically characterized.
+Recognizing that connectivity among genetically distinct populations may be relatively rare, researchers can gain additional insights into the frequency of immigration by also identifying descendants of migrants among study populations (directly [F1 hybrids] or advanced hybrids reflecting the effect of backcrossing across multiple generations [BC1, \ldots, BCX]).
+The process of recombination during meiosis creates a distribution of expected associations to parental populations among advanced hybrid classes.
+Treating linked loci as independent can lead to the false interpretation that the association of advanced hybrid classes to parental population occur strictly as discrete distributions.
+Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of linkage among loci available for analysis.
+	In response to these needs, we developed gscramble – an approach to simulate genotypes that minimizes bias associated with forward-in-time simulations by sampling alleles from respective populations without replacement, thus preserving the genetic differences observed in the empirical dataset.
+Further, by simulating individual genotypes based on user-defined pedigrees, gscramble allows for the simulation of admixed genotypes of varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
+By integrating species-specific recombination rates, gscramble simulates biologically informed genotypes that mirror empirical data.
+We develop and illustrate gscramble with use of single nucleotide polymorphic (SNP) genotypes.
+
+Can we say something about Anthony Almudevar's visionary method for GSI power assessment \citep{almudevar2000exact}
 
@@
 
-Not visible in the static plots are a number of dynamic features in the Shiny app display.
-Values of filtering criteria for a single locus can be changed in real time in the app, and the results are updated in the adjacent genotype biplot.
-Furthermore, it is possible to highlight points (representing individual samples) in the plot for one locus, and then observe the placement of those same selected samples at other loci.
-This is particularly well-suited to identifying samples with aberrant genotypes across multiple loci---a good indicator of contamination or other issues with the sample.
-
 \section*{Discussion}
@@
 \section*{Acknowledgements}
-We thank the members of the Molecular Ecology and Genetic Analysis (MEGA) team at NOAA SWFSC's Fisheries Ecology Division, etc. etc. .
+We thank etc. etc.   Contribution number  mHAVEAGAS-003

### 2023-10-20 author-title-etc.tex
add matt to authors.  tiny edits

@@
 % All caps tends to look better
-\newcommand{\myAuthors}{Eric C. Anderson$^{*,\S}$, Rachel M. Giglio$^{\dagger}$, Timothy J. Smyser$^{\dagger}$}
-\author{Eric C. Anderson$^{*,\S}$, Rachel M. Giglio$^{\dagger}$, Timothy J. Smyser$^{\dagger}$}
+\newcommand{\myAuthors}{Eric C. Anderson$^{*,\dagger,\ddag,\S}$, Rachel M. Giglio$^{\#}$, Matthew G. DeSaix$^{\ddag,\#}$ Timothy J. Smyser$^{\#}$}
+\author{Eric C. Anderson$^{*,\dagger,\ddag,\S}$, Rachel M. Giglio$^{\#}$, Matthew G. DeSaix$^{\ddag,\#}$ Timothy J. Smyser$^{\#}$}
 
@@
     Santa Cruz, CA 95060, USA.,
-    $^\dagger$  ADDRESS FOR APHIS/NWRC}
+    $^\dagger$Dept. of Fish, Wildlife, and Conservation Biology, Colorado State University, Fort Collins, CO, USA,
+    $^\ddag$Dept. of Biology, Colorado State University, Fort Collins, CO, USA,
+    $^\#$National Wildlife Research Center, United States Department of Agriculture, Wildlife Services,
+    Fort Collins, CO, USA
+}
 \renewcommand{\AuthorAddresses}{\myAffiliations}

### 2023-10-20 body-text-of-paper.tex
add matt to authors.  tiny edits

@@
 Genetic clustering algorithms are a commonly used approach to describe genetic structure among populations, reflecting the cumulative effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.
-The capacity of such algorithms to differentiate populations is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker set and sample sizes from the respective populations.
+The ability of such algorithms to differentiate populations is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker set and sample sizes from the respective populations.
 Based on an initial description of genetic structure, ecologists are frequently interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location and genetic origins of a given animal \citep{paetkau1995microsatellite,wilson2003bayesian}.

### 2023-10-21 body-text-of-paper.tex
Matt and Tim's writing

@@
 The ability of such algorithms to differentiate populations is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker set and sample sizes from the respective populations.
-Based on an initial description of genetic structure, ecologists are frequently interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location and genetic origins of a given animal \citep{paetkau1995microsatellite,wilson2003bayesian}.
-Describing such patterns of movement have many implications, influencing evolutionary processes, disease dynamics \citep{huestis2019windborne}, or patterns of invasive species expansion.
+Based on an initial description of genetic structure, ecologists are frequently interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.
+Genetic movement has many implications, influencing evolutionary processes, maintenance of genetic variation through gene flow, disease dynamics \citep{huestis2019windborne}, or patterns of invasive species expansion CITE.
 
@@
 
-Pairing empirical data with closely derived simulation-based approaches has been long recognized as a tool to help guide the interpretation of empirical data and define the limitations of the available data \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
-Mirroring empirical datasets with forward-in-time simulations often implement sampling with replacement (e.g.
-citations).
-However, sampling with replacement implicitly increases the pairwise genetic distance (i.e., FST) among populations and inflates the perceived ability to correctly assign individuals back to source populations \cite{almudevar2000exact,anderson2008improved}.
-\textcolor{red}{Seems like 2 additional sentences here would be appropriate to draw from a study that has demonstrated this process.}  The effects of overestimating our ability to resolve populations creates individuals that discretely align with simulated populations.
+	A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to pair the data with closely derived simulations where the truth is known \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
+	Techniques to simulate individual genotypes frequently use sampling with replacement from the observed allele frequencies.
+	However, sampling with replacement implicitly increases the pairwise genetic distance (e.g., FST) among populations and inflates the perceived ability to correctly assign individuals back to source populations \cite{almudevar2000exact,anderson2008improved}.
+	\textcolor{red}{Seems like 2 additional sentences here would be appropriate to draw from a study that has demonstrated this process.}  
 	These issues could be addressed with simulations that fix observed genetic differences among populations by sampling from the empirical data without replacement (citations for simulation frameworks that use this approach).
+
 	Beyond simply evaluating the resolution of discrete populations in a simulated framework, researchers may also be interested in how individuals of admixed ancestry are genetically characterized.  
@@
 	Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of linkage among loci available for analysis.
-	In response to these needs, we developed gscramble – an approach to simulate genotypes that minimizes bias associated with forward-in-time simulations by sampling alleles from respective populations without replacement, thus preserving the genetic differences observed in the empirical dataset.
+
+	In response to these needs, we developed gscramble – a simulation approach that samples alleles from respective populations without replacement, thus maintaining the genetic differences in the empirical dataset.
 	Further, by simulating individual genotypes based on user-defined pedigrees, gscramble allows for the simulation of admixed genotypes of varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
@@
 	We develop and illustrate gscramble with use of single nucleotide polymorphic (SNP) genotypes.
-
-Can we say something about Anthony Almudevar's visionary method for GSI power assessment \citep{almudevar2000exact}
 
@@
 
-\subsection*{And a third}
+\subsection*{Empirical Example}
 
-Blah, blah blobbity blah.
+Genetic samples were collected from invasive wild pigs across a SPATIAL EXTENT in southeastern Missouri that were lethally removed through population control efforts conducted as a component of the Missouri Feral Hog Elimination Partnership
+by US Department of Agriculture – Animal Plant Health Inspection Services – Wildlife Services, Missouri Department of Conservation, and other cooperative state agencies CHECK. 
+DNA was extracted from hair or tissue (collected from animals at the time of euthanasia) with commercially available magnetic bead recovery kits (MagMax DNA, Thermo Fisher Scientific).  
+Genotyping was performed with GeneSeek’s Genomic Profiler for Porcine HD which provides 62,128 biallelic SNP loci that are mapped across the 18 autosomal chromosomes. 
+We implemented standard genotype quality control metrics, specifically, we pruned genotypes with call rates <95\% and minor allele frequencies <5\%.
+We implemented linkage disequilibrium (LD) pruning using a window size of 50 loci and step size of 5 loci to remove markers above a linkage threshold of R2 > 0.5.
+
+GSCRAMBLE REFERENCE GROUPS
+GSCRAMBLE PEDIGREES
+ADMIXTURE GSCRAMBLE VS EMPIRICAL
+FST OF EMPIRICAL AND GSCRAMBLE CORE POPULATIONS
 

### 2023-10-21 body-text-of-paper.tex
Add the Introductory motivating example

--- 2023-10-20 body-text-of-paper.tex
@@
 
-Can we say something about Anthony Almudevar's visionary method for GSI power assessment \citep{almudevar2000exact}
-
 \section*{Methods}
 
-Something here.
+\subsection*{An introductory motivating example}
+
+Before we proceed to a description of the methodology used by gscramble,
+we demonstrate a case in which using a standard sampling-with-replacement
+approach to simulate genotypes leads to the incorrect conclusion that a set of genetic
+markers is sufficient to accurately estimate admixture fractions between two populations.
+To construct this simple example, we used the R programming language (CITE) to
+simulate allele frequencies for $L$ biallelic
+loci in a single population from a $\mathrm{Beta}(1, 8)$ distribution and then simulated
+two samples, $A$ and $B$ of size $N$ diploid genotypes from this population by drawing each of the three
+possible genotypes at each locus from their
+expected genotype frequencies under Hardy-Weinberg equilibrium.  In this case,
+each set of $N$ individuals is a sample from exactly the same population, so
+there would be no basis for performing population assignment between $A$ and $B$.
+
+However, we treated samples $A$ and $B$ as if they were sampled from two potentially different
+populations, and simulated $9n$ new individuals: $n$ new individuals at each of the 9 values of
+$q_A$, the admixture fraction for population $A$, in  $\{0, 0.125, 0.25, 0.375, 0.5, 0.625, 0.75, 0.875, 1.0\}$.
+We then added these $9n$ to the original $2N$ individuals ($N$ from sample $A$ and $N$ from $B$)
+and analyzed the genotypes using ADMIXTURE with $K=2$, doing a completely
+unsupervised analysis or doing a supervised analysis in which the original $N$ samples from each
+population were provided as learning samples.
+
+For each combination $L \in \{10^2, 10^3, 10^4, 10^5\}$, 
+$N \in \{25, 50, 100, 250\}$ and $n\in\{3,12,24\}$ we conducted multiple replicates
+of the entire process of
+simulating samples $A$ and $B$, simulating the $9n$ additional individuals, and
+then analyzing them with ADMIXTURE.  We conducted sufficient replicates so that,
+for each combination of $L$, $N$, and $n$ we had simulated 480 new individuals
+at each $q_A$ value; thus, $160$ replicates when $n = 3$, $40$ when $n=12$, and
+$20$ when $n=24$.
+
+We hypothesized that ADMIXTURE would return estimates of $q_A$ that were
+centered around the true values, even though there is no basis for that, because
+the simulations were done in a na\"{i}ve way that effectively assumes that the population
+allele frequencies are identical to those observed in the samples themselves. 
+We further expected that ADMIXTURE would return more accurate estimates of
+$q_A$ when $N$ was smaller and the number of loci was larger.
+We summarized the results by plotting boxplots of the 480
+ADMIXTURE-inferred $q_A$ values, for each combination of $L$, $N$, and $n$.
 
@@
 
+\section*{Results}
+
+\subsection*{An introductory motivating example}
+
+As expected, the perceived ability to estimate $q_A$ with ADMIXTURE was higher
+for larger values of $L$ and for smaller values of $N$ (Figure~\ref{fig:bias-sims}).  
+%%%
+\begin{figure*}
+\includegraphics[width=\textwidth]{figures/bias-sims-unsup-and-sup.pdf}
+\caption{ADMIXTURE estimates of $q_A$ as described in
+{\em An introductory motivating example} In these results, the ability to estimate $q_A$
+is a product of a bias.
+}
+\label{fig:bias-sims}
+\end{figure*}
+%%%
+It is interesting to note
+that, for numbers of loci,  $L$, that are 100 (or smaller) the effect is observed only when
+sample size, $N$, is as small as 25---and even then the effect is slight.  However, when the
+number of markers increases to 100,000---values that are becoming commonplace with the availability of chip-
+or sequencing-based approaches---it is apparent that the sampling-with-replacement approach
+induces an extreme bias.  With 100,000 markers, even if sample sizes are as large as 250 individuals,
+using na\"{i}ve simulations improperly indicate that admixture fractions of individuals can be
+accurately estimated, even when there are no genetic differences, whatsoever between the populations that
+samples $A$ and $B$ are drawn.
+
 \section*{Discussion}

### 2023-10-21 body-text-of-paper.tex

--- 2023-10-21 body-text-of-paper.tex
@@
 
-Something here.
+\subsection*{An introductory motivating example}
+
+Before we proceed to a description of the methodology used by gscramble,
+we demonstrate a case in which using a standard sampling-with-replacement
+approach to simulate genotypes leads to the incorrect conclusion that a set of genetic
+markers is sufficient to accurately estimate admixture fractions between two populations.
+To construct this simple example, we used the R programming language (CITE) to
+simulate allele frequencies for $L$ biallelic
+loci in a single population from a $\mathrm{Beta}(1, 8)$ distribution and then simulated
+two samples, $A$ and $B$ of size $N$ diploid genotypes from this population by drawing each of the three
+possible genotypes at each locus from their
+expected genotype frequencies under Hardy-Weinberg equilibrium.  In this case,
+each set of $N$ individuals is a sample from exactly the same population, so
+there would be no basis for performing population assignment between $A$ and $B$.
+
+However, we treated samples $A$ and $B$ as if they were sampled from two potentially different
+populations, and simulated $9n$ new individuals: $n$ new individuals at each of the 9 values of
+$q_A$, the admixture fraction for population $A$, in  $\{0, 0.125, 0.25, 0.375, 0.5, 0.625, 0.75, 0.875, 1.0\}$.
+We then added these $9n$ to the original $2N$ individuals ($N$ from sample $A$ and $N$ from $B$)
+and analyzed the genotypes using ADMIXTURE with $K=2$, doing a completely
+unsupervised analysis or doing a supervised analysis in which the original $N$ samples from each
+population were provided as learning samples.
+
+For each combination $L \in \{10^2, 10^3, 10^4, 10^5\}$,
+$N \in \{25, 50, 100, 250\}$ and $n\in\{3,12,24\}$ we conducted multiple replicates
+of the entire process of
+simulating samples $A$ and $B$, simulating the $9n$ additional individuals, and
+then analyzing them with ADMIXTURE.  We conducted sufficient replicates so that,
+for each combination of $L$, $N$, and $n$ we had simulated 480 new individuals
+at each $q_A$ value; thus, $160$ replicates when $n = 3$, $40$ when $n=12$, and
+$20$ when $n=24$.
+
+We hypothesized that ADMIXTURE would return estimates of $q_A$ that were
+centered around the true values, even though there is no basis for that, because
+the simulations were done in a na\"{i}ve way that effectively assumes that the population
+allele frequencies are identical to those observed in the samples themselves.
+We further expected that ADMIXTURE would return more accurate estimates of
+$q_A$ when $N$ was smaller and the number of loci was larger.
+We summarized the results by plotting boxplots of the 480
+ADMIXTURE-inferred $q_A$ values, for each combination of $L$, $N$, and $n$.
 
@@
 
+\section*{Results}
+
+\subsection*{An introductory motivating example}
+
+As expected, the perceived ability to estimate $q_A$ with ADMIXTURE was higher
+for larger values of $L$ and for smaller values of $N$ (Figure~\ref{fig:bias-sims}).
+%%%
+\begin{figure*}
+\includegraphics[width=\textwidth]{figures/bias-sims-unsup-and-sup.pdf}
+\caption{ADMIXTURE estimates of $q_A$ as described in
+{\em An introductory motivating example} In these results, the ability to estimate $q_A$
+is a product of a bias.
+}
+\label{fig:bias-sims}
+\end{figure*}
+%%%
+It is interesting to note
+that, for numbers of loci,  $L$, that are 100 (or smaller) the effect is observed only when
+sample size, $N$, is as small as 25---and even then the effect is slight.  However, when the
+number of markers increases to 100,000---values that are becoming commonplace with the availability of chip-
+or sequencing-based approaches---it is apparent that the sampling-with-replacement approach
+induces an extreme bias.  With 100,000 markers, even if sample sizes are as large as 250 individuals,
+using na\"{i}ve simulations improperly indicate that admixture fractions of individuals can be
+accurately estimated, even when there are no genetic differences, whatsoever between the populations that
+samples $A$ and $B$ are drawn.
+
 \section*{Discussion}

### 2023-10-22 body-text-of-paper.tex
Further edits on Intro and Methods

--- 2023-10-21 body-text-of-paper.tex
@@
 Based on an initial description of genetic structure, ecologists are frequently interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.
-Genetic movement has many implications, influencing evolutionary processes, maintenance of genetic variation through gene flow, disease dynamics \citep{huestis2019windborne}, or patterns of invasive species expansion CITE.
+Genetic movement has many implications, such as influencing evolutionary processes, maintaining genetic variation through gene flow, and driving disease dynamics \citep{huestis2019windborne} and affecting patterns of invasive species expansion CITE.
 
-	In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals with complex ancestry patterns (i.e., the proportion of an individual’s genome assigned to sampled populations, sensu \citealt{pritchard2000inference}) that demonstrate mixed associations to multiple populations.
+	In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of mixed ancestry---a trait often summarized as the proportion of an individual’s genome originating from each of $K$ subpopulations, sensu \citet{pritchard2000inference}.
 Various processes could contribute to such mixed associations.
-Most simplistically, individuals may represent the influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately representing contributions from multiple populations included in the study.
-However, alternative processes could also create similarly complex ancestry patterns.
-For example, individuals may have genomic associations to unsampled populations as immigrants or as descendant of immigrant from unsampled populations \citep{pritchard2000inference,beerli2004effect}.
-Yet a limitation of clustering algorithms is that exogenous immigrants can only be associated to populations represented within the sample.
-Similarly, individuals with mixed associations may reflect a discrepancy between the genetic structure as described by the analysis and the underlying ecological process creating patterns of genetic differentiation (e.g., when clinal systems governed by isolation by distance are described as discrete genetic clusters).
-Finally, such patterns may represent a limitation of the statistical power of the genetic data, as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure.
-The challenge for researchers then becomes correctly identifying the ecological process by which similar, complex ancestry patterns were created.
+Most directly, individuals may represent the influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately representing contributions from multiple sampled populations.
+However, similar patterns may result from functionally different patterns of population structuring such as when .
+For example, when clinal systems governed by isolation by distance are described as discrete genetic clusters.
+Furthermore, inferred patterns of admixture may represent a limitation of the statistical power of the genetic data, as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure.
+The challenge for researchers then becomes correctly identifying the ecological and/or statistical processes by which observed complex ancestry patterns were created.
 
-	A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to pair the data with closely derived simulations where the truth is known \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
-	Techniques to simulate individual genotypes frequently use sampling with replacement from the observed allele frequencies.
-	However, sampling with replacement implicitly increases the pairwise genetic distance (e.g., FST) among populations and inflates the perceived ability to correctly assign individuals back to source populations \cite{almudevar2000exact,anderson2008improved}.
-	\textcolor{red}{Seems like 2 additional sentences here would be appropriate to draw from a study that has demonstrated this process.}  
-	These issues could be addressed with simulations that fix observed genetic differences among populations by sampling from the empirical data without replacement (citations for simulation frameworks that use this approach).
+A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to conduct
+parallel analyses on simulated data where the truth is known, structured to mirror the empirical data set
+as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
+Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the 
+observed allele frequencies (see CITE, CITE, CITE).
+However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed samples, and 
+inflates the perceived ability to correctly assign individuals back to source populations. \citet{anderson2008improved}
+documented this type of artifact in the context of simulations to assess the power for
+genetic stock identification (GSI), a type of clustering application in which individuals are constrained to have
+all their ancestry from a single subpopulation. We demonstrate, below, that similar artifacts occur when simulating data to 
+assess how accurately admixture fractions can be estimated.  We refer to this phenomenon as resampling-induced
+power inflation, or RIPI for short.  
+In this study, we additionally develop and illustrate an approach by which RIPI can be eliminated
+from simulations by constraining the genetic differences among simulated populations to match
+those of populations from which the samples were obtained.
 
-	Beyond simply evaluating the resolution of discrete populations in a simulated framework, researchers may also be interested in how individuals of admixed ancestry are genetically characterized.  
-	Recognizing that connectivity among genetically distinct populations may be relatively rare, researchers can gain additional insights into the frequency of immigration by also identifying descendants of migrants among study populations (directly [F1 hybrids] or advanced hybrids reflecting the effect of backcrossing across multiple generations [BC1, \ldots, BCX]).
+Beyond simply evaluating the resolution of discrete populations in a simulated framework, researchers may also be 
+interested in how individuals of admixed ancestry are genetically characterized.
+Recognizing that connectivity among genetically distinct populations may be relatively rare, additional insight into the 
+frequency of immigration can be gained by 
+identifying the descendants of migrants in addition to those individuals that may be direct migrants.
+
+direct migrants but also identifying the influence of migration over multiple generations (i.e., F1 hybrids or advanced hybrids reflecting the effect of backcrossing across multiple generations [BC1, \ldots,
+BCX]).
 	The process of recombination during meiosis creates a distribution of expected associations to parental populations among advanced hybrid classes.
@@
 
-Something here.
+\subsection*{An introductory motivating example}
+
+Before we proceed to a description of the methodology used by gscramble,
+we demonstrate a case in which using a standard sampling-with-replacement
+approach induces RIPI when estimating admixture fractions.
+To construct this simple example, we used the R programming language \citep{Rcoreteam} to
+simulate allele frequencies for $L$ biallelic
+loci in a single population from a $\mathrm{Beta}(1, 8)$ distribution and then simulated
+two samples, $A$ and $B$, each of size $N$ diploid genotypes. The genotypes in $A$ and $B$ were simulated
+from the same allele frequencies assuming Hardy-Weinberg equilibrium.
+
+In this case,
+each set of $N$ individuals is a sample from exactly the same population, so
+there is no basis for performing population assignment between $A$ and $B$;
+however, we treated samples $A$ and $B$ as if they were sampled from two potentially different
+populations, and simulated $9n$ new individuals: $n$ new individuals at each of the 9 values of
+$q_A$, the admixture fraction for population $A$, in  $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$.
+We simulated these genotypes according to the "admixture model" used in {\em structure} \citep{pritchard2000inference}.
+Briefly, at each locus, independently, the origin
+of each gene copy was simulated to be from
+population $A$ with probability $q_A$ and from population $B$ with probability $1-q_A$.  Subsequently,
+the allelic types of gene copies from $A$ ($B$), at a locus, were simulated by sampling with replacement
+from the alleles at that locus in sample $A$ ($B$).
+We then added these $9n$ simulated individuals to the $N$ original individuals  from $A$ and $N$ from $B$
+and analyzed all the genotypes using ADMIXTURE with $K=2$. On each data set we performed a supervised analysis in which the original $N$ samples from each
+population were provided as learning samples, and also an
+unsupervised analysis when the origin of the samples from $A$ and $B$ were regarded as unknown.
+
+For each combination of $L \in \{10^2, 10^3, 10^4, 10^5\}$,
+$N \in \{25, 50, 100, 250\}$ and $n\in\{3,12,24\}$ we conducted multiple replicates
+of the entire process of
+simulating samples $A$ and $B$, simulating the $9n$ additional individuals, and
+then analyzing them with ADMIXTURE.  We conducted $R$ replicates so that,
+for each combination of $L$, $N$, and $n$ we had simulated $Rn=480$ new individuals
+at each $q_A$ value; thus, $R=160$ replicates when $n = 3$, $R=40$ when $n=12$, and
+$R=20$ when $n=24$.
+
+Because the simulations were done in a na\"{i}ve way that effectively assumes that the population
+allele frequencies are identical to those observed in the samples themselves, we hypothesized that ADMIXTURE would return estimates of $q_A$ that were
+centered around the true values, even though samples $A$ and $B$ were drawn from the same
+population.
+We further expected that ADMIXTURE would return more accurate estimates of
+$q_A$ when $N$ was smaller and the number of loci was larger.
+We summarized the results by plotting boxplots of the ADMIXTURE-inferred $q_A$ values,
+ for the 480 newly simulated individuals (and for the original samples, $A$ and $B$, in the
+ case of unsupervised clustering),
+for each combination of $L$, $N$, and $n$.
 
@@
 
-For the purposes here, we define blah, blah.
+Our approach for simulating admixed individuals without inflating perceived power for
+the estimation of their admixture fractions is motivated by the traditional cross-validation
+approach for evaluating the power of mixture proportion
+estimation in genetic stock identification \citep{Moran&AndersonRubias}.
+In that cross-validation approach, individuals are removed from the reference
+samples and placed in a mixture sample of "unknowns", whose origin
+is inferred using the remaining individuals in the reference samples.  A key
+feature of this approach is that no new genetic material is being created by
+sampling with replacement; however the genetic material of all individuals
+in the original samples is being used---either within the remaining individuals in the
+reference samples or within the individuals in the mixture. This latter condition
+is not essential, but it does indicate that the methodology is using all the
+available individuals for estimating power through simulation.
 
-\subsection*{Another subsection}
+Our scheme extends the traditional cross-validation approach
+by simulating the descent of chunks of chromosome through a user-specified
+pedigree to create admixed individuals.  By imposing certain constraints
+on the pedigree, our framework ensures that all the genetic material
+within the original individuals is used to either create admixed individuals or to
+remain within reference individuals (i.e., those that are purely of one
+putative subopulation), and yet no new genetic material is
+created by sampling with replacement from the original samples.  We call the structure
+we use to do this a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
+a GSP includes a set of founders (individuals that have no parents specified in the
+pedigree), and it includes non-founders; however, we introduce an additional node
+type in the GSP that represents samples taken from the non-founders. These samples
+are the simulated admixed individuals.  A GSP must have no inbreeding loops, since
+any inbreeding loop indicates that more than one copy of a gene in an ancestor has
+been created amongst its descendants, which is a type of sampling without
+replacement.  Furthermore, in a GSP, the amount of genetic material taken from the
+samples at the sample nodes should be equal to the amount of material in the founders,
+to ensure that all available data is being used to estimate power for the estimation of
+admixture fractions.
 
-\subsection*{Empirical Example}
+Figure~\ref{fig:gsp1} shows an example GSP for the simple case
+of simulating $F_2$ individuals. The figure and its caption
+should be studied, as the following section
+uses terminology that is defined in the figure caption.
+%%%%%%%%%%%
+\begin{figure}
+\newcommand{\gspcapone}{\footnotesize 
+(I NEED TO UPDATE THE FIGURE TO MATCH THE DESCRIPTION)
+A genomic simulation pedigree (GSP) for simulating $F_2$ individuals.
+Squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
+individuals 1--4.  The two triangles above each of those founders denotes, in color
+and text, the population of origin of each of the two genomes within each founder.
+Thus, individuals~1 and~2 are founders purely from population~$A$, and~3 and~4
+are purely from population~$B$.  The pink hexagon labeled s7 is the sample node.
+It represents samples that are taken from non-founder node~7, which, from the structure
+of the pedigree, clearly represents an $F_2$ hybrid between populations $A$ and $B$.  The red numbers
+on edges above and leading into each non-founder node indicate the number of
+gametes that get segregated from the parent node (at the top of the edge) to the
+daughter node (at the bottom of the edge).
+This emphasizes that,
+in dropping genes through a GSP, in contrast
+to a traditional pedigree, each parent
+node may segregate multiple gametes to a daughter node.  See the text for
+further explanation. 
+Each non-founder node $x$ will have exactly two edges coming into it from above
+(from its two parents), we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$. 
+For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
+A non-founder individual may have more or less than two edges coming downward
+out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
+non-founder daughter nodes, we denote the number of gametes segregated
+through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example,
+non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
+Finally, the purple number adjacent to the edge leading into the
+sample node indicates how many sample individuals are drawn from individual
+node 7. Any non-founder individual node can have a maximum of one edge leading
+to a sample node.  We denote the number of samples emanating from node $x$
+node as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
+created from a single simulation on this GSP.
+}
+\begin{center}
+\includegraphics[width=\columnwidth]{images/gsp4-700.pdf}
+\end{center}
+\caption[\gspcapone]{\gspcapone}
+\label{fig:gsp1}
+\end{figure}
+%%%%%%%%%%%%
+It is important to understand that, apart from the founder nodes, the individual
+nodes in a GSP do not necessarily represent only a single individual.  Specifically,
+since we are preserving all genetic material, multiple gametes may get segregated
+through each non-founder node in a GSP.   Genetic material is segregated through
+a GSP using these steps, done upon each node in an order such that the steps
+are run on each node's parents before being run on the node itself:
+\begin{enumerate}
+\item Each genome in a triangle above a founder node is assorted into
+two gametes with recombination occurring between each chromosome
+with probability $\frac{1}{2}$ and within each chromosome according to
+a user-specified recombination map.  
+\item Each non-founder individual node, $x$, will have exactly two edges coming
+into it---one from each parent.  Each of the $G^+_{x,1}$ gametes coming in from the
+first edge is united with a single one of the $G^+_{x,2}$ gametes coming
+in on the second edge (in a random order). 
+\item If a non-founder individual node $x$ has an edge to a sample node,
+a randomly chosen $S_x$ of the united pairs of gametes are delivered
+to the sample node to constitute the $S_x$ samples taken from individual
+node $x$. Any remaining united pairs of gametes are treated as in (4) below.
+\item United pairs of gametes (or those remaining after delivery
+to a sample node) in each non-founder individual node $x$ are segregated into
+gametes with recombination, and each of these gametes is assigned, in
+random order, to the gamete edges proceeding downward out of the node,
+with the number of gametes assigned to the $i\thh$ downward edge being
+$G^-_{x,i}$. 
+\end{enumerate}
+At the end of this process, the samples simulated (as united pairs of gametes)
+into the sample nodes will contain all of the genetic material found in the founders (and
+no more than that)
+but that material will have been segregated and recombined according to the pedigree.  Thus the
+samples obtained from a GSP can be simulated with the expected admixture fractions
+for any hybrid category that can be specified by a pedigree.  They have also been simulated
+in a way that 1) respects physical linkage---each individual inherits segments of the genome
+simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
+genetic material is entirely without replacement so it will not induce any bias in clustering
+power assessment; and 3) no genetic material has been lost---all the genetic material of the founders
+is represented in the samples, maximizing the number of individual and markers available to
+use in downstream analyses.
+
+In the gscramble package there is a convenience function called {\footnotesize\tt create\_GSP()} that
+will create GSPs for samples up to and including second-generation backcrosses; however, gscramble is designed
+to handle user-specified GSPs of arbitrary complexity.  When creating one's own GSP, it is useful
+to keep in mind a handful of requirements for a GSP to be valid.
+Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder individual nodes, and $\nonfounds$ for all non-founder
+nodes that are not parents of other indivduals, but are adjacent to a sample node, $\nonfounde$ for all
+non-founder nodes that have individual node daughters, but no sample nodes, and $\nonfoundse$ for all
+non-founder nodes that are parents of individual nodes and also adjacent to sample-nodes, the necessary conditions 
+for a valid GSP are:
+\begin{equation}
+\begin{aligned}
+\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders|   \\ 
+\sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
+G^+_{x,1} &= G^+_{x,2} & \forall x \in \nonfound \\
+\sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\ 
+\sum_{i=1}^{e_x} G^-_{x,i} + S_x/2 &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfoundse.  \\ 
+\end{aligned}
+\end{equation}
+
+\subsection*{Empirical Examples}
+
+\subsubsection*{Rainbow/cutthroat trout hybrids in California}
+
+Boing.
+
+\subsubsection*{Feral pigs in Missouri}
 
@@
 
+\section*{Results}
+
+\subsection*{An introductory motivating example}
+
+As expected, the perceived ability to estimate $q_A$ with ADMIXTURE was higher
+for larger values of $L$ and for smaller values of $N$ (Figure~\ref{fig:bias-sims}).
+%%%
 \begin{figure*}
-\begin{center}
-\includegraphics[width=0.8\textwidth]{images/gsp4-700.pdf}
-\end{center}
-\caption[]{Stick this thing in somewhere}
+\newcommand{\biassimscap}{\footnotesize ADMIXTURE estimates of $q_A$ as described in
+{\em An introductory motivating example}. In these results, the apparent ability to estimate $q_A$
+is a bias resulting from the use of sampling with replacement to simulate new, admixed genotypes to
+test in ADMIXTURE.  In each figure, the different columns represent different numbers of markers
+from 10 to 100,000, while different rows represent different original sample sizes, $N$, taken. Colors
+of the boxplots indicate how many new individuals, $n$, of each $q_A$ value were simulated during each
+replicate.}
+\includegraphics[width=0.96\textwidth]{figures/bias-sims-unsup-and-sup.pdf}
+\caption[\biassimscap]{\biassimscap}
+\label{fig:bias-sims}
 \end{figure*}
+%%%
+It is interesting to note
+that, for numbers of loci,  $L$, that are 100 (or smaller) the effect is observed only when
+sample size, $N$, is as small as 25---and even then the effect is slight.  However, when the
+number of markers increases to 100,000---values that are becoming commonplace with the availability of chip-
+or sequencing-based approaches---it is apparent that the sampling-with-replacement approach
+induces an extreme bias.  With 100,000 markers, even if sample sizes ($N$) are as large as 250 individuals,
+using na\"{i}ve simulations improperly indicates that admixture fractions of individuals can be
+accurately estimated, even when there are no genetic differences, whatsoever between the populations that
+samples $A$ and $B$ are drawn.
 
 \section*{Discussion}
-Hey! Let's discuss this!
+
+Somewhere here I want to mention that the Anderson et al. (2008) approach for simulating
+mixtures of genotypes to assess how well one can estimate mixture proportions---the fraction
+of individuals from each of a set of baseline or reference populations---allows the user to
+simulate, if desired, more genotypes than occur in the reference samples.  This is because
+a trick is done in the inference stage to remove the correlation between the simulated individual
+and the reference sample, relative to the true population, by removing the simulated individual's
+genotypes from its own reference population when calculating the likelihood that its genotype
+originated from each of the reference populations.  To do the same when doing simulations for
+assessing power using ADMIXTURE or STRUCTURE would require hacking the underlying code
+of those programs, which is not desirable.  So, for these cases we have used the sampling
+without replacement approach.
 

### 2023-10-22 body-text-of-paper.tex
Tim's intro edits

@@
 
-	In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of mixed ancestry---a trait often summarized as the proportion of an individual’s genome originating from each of $K$ subpopulations, sensu \citet{pritchard2000inference}.
+	In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of mixed ancestry--- 
+	characterized as individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference.}
+	a trait often summarized as the proportion of an individual’s genome originating from each of $K$ subpopulations, sensu \citet{pritchard2000inference}.
 Various processes could contribute to such mixed associations.
 Most directly, individuals may represent the influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately representing contributions from multiple sampled populations.
-However, similar patterns may result from functionally different patterns of population structuring such as when .
-For example, when clinal systems governed by isolation by distance are described as discrete genetic clusters.
-Furthermore, inferred patterns of admixture may represent a limitation of the statistical power of the genetic data, as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure.
+However, similar patterns may result from functionally different patterns of population structuring.
+For example, such patterns might be expected when clinal systems governed by isolation by distance are described as discrete genetic clusters.
+	NOTE I think a line here about exogenous immigrants might be worthwhile to set up 2 biological processes paired with a statistical process, something like END NOTE
+	Alternatively, individuals of mixed ancestry could reflect immigration from outlying populations (e.g., exogenous immigration), with true source populations not sufficiently
+	represented in the sample to be among the characterized $K$ subpopulations.
+	Furthermore, the characterization of admixed individuals may reflect a limitation of the statistical power of the genetic data, 
+	as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure. 
 	 The challenge for researchers then becomes correctly identifying the ecological and/or statistical processes by which observed complex ancestry patterns were created.
@@
 genetic stock identification (GSI), a type of clustering application in which individuals are constrained to have
-all their ancestry from a single subpopulation. We demonstrate, below, that similar artifacts occur when simulating data to 
+all their ancestry from a single subpopulation. 
+, a type of clustering application with the limitation that individuals are fully assigned to single subpopulations (ignoring the possibility of mixed ancestry).
+We demonstrate, below, that similar artifacts occur when simulating data to 
 assess how accurately admixture fractions can be estimated.  We refer to this phenomenon as resampling-induced
 power inflation, or RIPI for short.  
+NOTE I would proposed holding on to the next sentence as we build the idea of mixed ancestry in the next paragraph END NOTE
 In this study, we additionally develop and illustrate an approach by which RIPI can be eliminated
@@
 Beyond simply evaluating the resolution of discrete populations in a simulated framework, researchers may also be 
-interested in how individuals of admixed ancestry are genetically characterized.
+interested in how individuals of admixed ancestry are genetically characterized given the statistical power of the available dataset and the analysis methods used.
 Recognizing that connectivity among genetically distinct populations may be relatively rare, additional insight into the 
-frequency of immigration can be gained by 
-identifying the descendants of migrants in addition to those individuals that may be direct migrants.
-
-direct migrants but also identifying the influence of migration over multiple generations (i.e., F1 hybrids or advanced hybrids reflecting the effect of backcrossing across multiple generations [BC1, \ldots,
+frequency of immigration can be gained by integrating across generations to identify both direct migrants and descendants of migrants
+ (i.e., F1 hybrids or advanced hybrids reflecting the effect of backcrossing across multiple generations [BC1, \ldots, 
 BCX]).
-	The process of recombination during meiosis creates a distribution of expected associations to parental populations among advanced hybrid classes.
+	However, the process of recombination during meiosis creates a distribution of expected associations to parental populations among advanced hybrid classes.
 	Treating linked loci as independent can lead to the false interpretation that the association of advanced hybrid classes to parental population occur strictly as discrete distributions.

### 2023-10-22 body-text-of-paper.tex
Eric edited methods a bit

--- 2023-10-22 body-text-of-paper.tex
@@
 \newcommand{\gspcapone}{\footnotesize 
-(I NEED TO UPDATE THE FIGURE TO MATCH THE DESCRIPTION)
+(I NEED TO UPDATE THE FIGURE TO MATCH THE DESCRIPTION AND MAKE
+THE LABELS BIGGER IN THE SQUARES).
 A genomic simulation pedigree (GSP) for simulating $F_2$ individuals.
@@
 since we are preserving all genetic material, multiple gametes may get segregated
-through each non-founder node in a GSP.   Genetic material is segregated through
+through each non-founder node in a GSP\@.   Genetic material is segregated through
 a GSP using these steps, done upon each node in an order such that the steps
@@
 \begin{enumerate}
-\item Each genome in a triangle above a founder node is assorted into
-two gametes with recombination occurring between each chromosome
+\item Each genome in a triangle above a founder node represents a gamete.
+These gametes are united within the founder individual and then
+following recombination between each chromosome
 with probability $\frac{1}{2}$ and within each chromosome according to
-a user-specified recombination map.  
-\item Each non-founder individual node, $x$, will have exactly two edges coming
+a user-specified recombination map, 
+ one gamete is assorted down each edge from the founder node to its daughter.
+\item Each non-founder node, $x$, will have exactly two edges coming
 into it---one from each parent.  Each of the $G^+_{x,1}$ gametes coming in from the
@@
 in on the second edge (in a random order). 
-\item If a non-founder individual node $x$ has an edge to a sample node,
+\item If a non-founder node $x$ has an edge to a sample node,
 a randomly chosen $S_x$ of the united pairs of gametes are delivered
@@
 node $x$. Any remaining united pairs of gametes are treated as in (4) below.
-\item United pairs of gametes (or those remaining after delivery
-to a sample node) in each non-founder individual node $x$ are segregated into
+\item United pairs of gametes (including those remaining after delivery
+to a sample node) in each non-founder node $x$ are segregated into
 gametes with recombination, and each of these gametes is assigned, in
@@
 simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
-genetic material is entirely without replacement so it will not induce any bias in clustering
-power assessment; and 3) no genetic material has been lost---all the genetic material of the founders
-is represented in the samples, maximizing the number of individual and markers available to
+genetic material is entirely without replacement so it will not induce RIPI; and 3) no genetic material has been lost---all the genetic material of the founders
+is represented in the samples, maximizing the number of individuals and markers available to
 use in downstream analyses.
@@
 to keep in mind a handful of requirements for a GSP to be valid.
-Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder individual nodes, and $\nonfounds$ for all non-founder
-nodes that are not parents of other indivduals, but are adjacent to a sample node, $\nonfounde$ for all
-non-founder nodes that have individual node daughters, but no sample nodes, and $\nonfoundse$ for all
-non-founder nodes that are parents of individual nodes and also adjacent to sample-nodes, the necessary conditions 
-for a valid GSP are:
+Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
+nodes that are not parents of other non-founders, but are adjacent to a sample node, $\nonfounde$ for all
+non-founder nodes that have non-founder node daughters, but no sample nodes, and $\nonfoundse$ for all
+non-founder nodes that are parents of non-founder nodes and also adjacent to sample-nodes, the necessary conditions 
+for a valid GSP, in addition to having no inbreeding loops, are:
 \begin{equation}
 \begin{aligned}
-\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders|   \\ 
+\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders| &  \\ 
 \sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
@@
 \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\ 
-\sum_{i=1}^{e_x} G^-_{x,i} + S_x/2 &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfoundse.  \\ 
+ S_x/2 + \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfoundse.  \\ 
 \end{aligned}
 \end{equation}
+Before simulating from a GSP, gscramble checks to ensure the GSP
+satisfies these validity conditions.
+
+\subsection*{Permutation procedures}
+
+gscramble uses a GSP to segregate segments of genetic material from the founders
+to the samples.  At the end of segment simulation, the genome of each sample
+is recorded as a mosaic of segments that derived from different founders.  In order
+to create a data set of markers in each sample, it is necessary only to identify the alleles
+that occurred on each segment within the founders and propagate those alleles to the
+corresponding segments within the samples.  At the stage of assigning alleles to
+founders, gscramble provides a variety of options for to permute alleles amongst
+individuals within samples from which the founders are drawn.  By using permutation,
+variability in the data is created, but without sampling with replacement, and hence
+not inducing RIPI.
 

### 2023-10-25 body-text-of-paper.tex
A number of edits
Added permutations figure

@@
 	In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of mixed ancestry---
-	characterized as individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference.}
+	characterized as individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference}.
 	a trait often summarized as the proportion of an individual’s genome originating from each of $K$ subpopulations, sensu \citet{pritchard2000inference}.
@@
 
-	In response to these needs, we developed gscramble – a simulation approach that samples alleles from respective populations without replacement, thus maintaining the genetic differences in the empirical dataset.
-	Further, by simulating individual genotypes based on user-defined pedigrees, gscramble allows for the simulation of admixed genotypes of varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
-	By integrating species-specific recombination rates, gscramble simulates biologically informed genotypes that mirror empirical data.
-	We develop and illustrate gscramble with use of single nucleotide polymorphic (SNP) genotypes.
+	In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without replacement, thus maintaining the genetic differences in the empirical dataset.
+	Further, by simulating individual genotypes based on user-defined pedigrees, \gscramble{} allows for the simulation of admixed genotypes of varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
+	By integrating species-specific recombination rates, \gscramble{} simulates biologically informed genotypes that mirror empirical data.
+	We develop and illustrate \gscramble{} with use of single nucleotide polymorphic (SNP) genotypes.
 
@@
 
-Before we proceed to a description of the methodology used by gscramble,
+Before we proceed to a description of the methodology used by \gscramble{},
 we demonstrate a case in which using a standard sampling-with-replacement
@@
 each set of $N$ individuals is a sample from exactly the same population, so
-there is no basis for performing population assignment between $A$ and $B$;
+there is clearly no basis for performing population assignment between $A$ and $B$;
 however, we treated samples $A$ and $B$ as if they were sampled from two potentially different
@@
 approach for evaluating the power of mixture proportion
-estimation in genetic stock identification \citep{Moran&AndersonRubias}.
-In that cross-validation approach, individuals are removed from the reference
+estimation in genetic stock identification (e.g., the cross-validation approach
+desribed in \citealt{moran2019bayesian}).
+In the cross-validation approach, individuals are removed from the reference
 samples and placed in a mixture sample of "unknowns", whose origin
@@
 in the original samples is being used---either within the remaining individuals in the
-reference samples or within the individuals in the mixture. This latter condition
-is not essential, but it does indicate that the methodology is using all the
-available individuals for estimating power through simulation.
+reference samples or within the individuals in the mixture.
 
-Our scheme extends the traditional cross-validation approach
+Our scheme, described here, extends the traditional cross-validation approach
 by simulating the descent of chunks of chromosome through a user-specified
@@
 putative subopulation), and yet no new genetic material is
-created by sampling with replacement from the original samples.  We call the structure
-we use to do this a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
+created by sampling with replacement from the original samples.
+
+The call the structure
+we use to do this sort of cross-validation simulation of admixed individuals
+a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
 a GSP includes a set of founders (individuals that have no parents specified in the
@@
 are the simulated admixed individuals.  A GSP must have no inbreeding loops, since
-any inbreeding loop indicates that more than one copy of a gene in an ancestor has
-been created amongst its descendants, which is a type of sampling without
+any inbreeding loop indicates that more than one copy of a gene in an ancestor may
+be created amongst its descendants, which is a type of sampling without
 replacement.  Furthermore, in a GSP, the amount of genetic material taken from the
@@
 \newcommand{\gspcapone}{\footnotesize
-(I NEED TO UPDATE THE FIGURE TO MATCH THE DESCRIPTION AND MAKE
-THE LABELS BIGGER IN THE SQUARES).
 A genomic simulation pedigree (GSP) for simulating $F_2$ individuals.
 Squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
-individuals 1--4.  The two triangles above each of those founders denotes, in color
+individuals 1--4.  The inverted triangles above each of those founders denotes, in color
 and text, the population of origin of each of the two genomes within each founder.
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{images/gsp4-700.pdf}
+\includegraphics[width=\columnwidth]{figures/F2.pdf}
 \end{center}
@@
 since we are preserving all genetic material, multiple gametes may get segregated
-through each non-founder node in a GSP\@.   Genetic material is segregated through
+through each non-founder node in a GSP\@.
+
+Genetic material is segregated through
 a GSP using these steps, done upon each node in an order such that the steps
@@
 \begin{enumerate}
-\item Each genome in a triangle above a founder node represents a gamete.
+\item Each genome in a triangle above a founder node represents a gamete
+carrying a single, haploid copy of the genome.
 These gametes are united within the founder individual and then
@@
 
-In the gscramble package there is a convenience function called {\footnotesize\tt create\_GSP()} that
-will create GSPs for samples up to and including second-generation backcrosses; however, gscramble is designed
+In the \gscramble{} package there is a convenience function called {\footnotesize\tt create\_GSP()} that
+will create GSPs for samples up to and including second-generation backcrosses; however, \gscramble{} is designed
 to handle user-specified GSPs of arbitrary complexity.  When creating one's own GSP, it is useful
@@
 \end{equation}
-Before simulating from a GSP, gscramble checks to ensure the GSP
-satisfies these validity conditions.
+Before simulating from a GSP, \gscramble{} checks to ensure the GSP
+satisfies these validity conditions.  The simulation of genomic segments
+from founders to samples in a GSP is performed by the \gscramble{} function
+{\footnotesize\tt segregate()}, which requires as input a list of GSPs through which
+to segregate genomic segments, and, for each such GSP, a table that maps the
+population labels on each GSP (in the inverted triangles atop the founders) to the
+different sampled groups of individuals.
 
@@
 
-gscramble uses a GSP to segregate segments of genetic material from the founders
+\gscramble{} uses a GSP to segregate segments of genetic material from the founders
 to the samples.  At the end of segment simulation, the genome of each sample
@@
 corresponding segments within the samples.  At the stage of assigning alleles to
-founders, gscramble provides a variety of options for to permute alleles amongst
+founders, \gscramble{} provides a variety of ways to permute alleles amongst
 individuals within samples from which the founders are drawn.  By using permutation,
@@
 not inducing RIPI.
+
+To execute these permutations, it is necessary to have a matrix of genotypes for
+the individuals, and also, each population must belong to exactly one
+population.  Permutation of alleles then proceeds within populations, hence the frequencies
+of alleles from each population remain the same as those observed
+in the samples themselves. Following permutation, there remain individuals 1,2,\ldots,
+in each population, but the alleles that they carry will be different after permutation.  These
+individuals are assigned in order to be the founders from each population in the GSP. Subsequently,
+after genetic material is segregated to the samples, the alleles found on each genomic segment in those permuted
+founders are then propagated to the samples.  It is worth noting that this approach allows for simulation
+with large genomic data sets, with hundreds of thousands to millions of markers, since segregation
+through the pedigree is not done via by segments, and not on a locus-by-locus basis. The only operations
+involving all the alleles in all the individuals are the permutation steps and indexing into the segments,
+which may be done efficiently with matrices in R.
+
+Permutation of alleles is handled within the \gscramble{} function {\footnotesize\tt segments2markers()}
+which takes as input the segments segregated down a GSP, a matrix of sampled genotypes, and a specification
+of the group that each sampled individuals belongs to.
+%%%%%%%
+\newcommand{\permcap}{\footnotesize An illustration of how \gscramble{} permutes data under
+different options.  {\bf (a)}~A picture
+representing an original data set of 12 diploids sampled from three populations and typed at 15 markers. 
+Different colors represent different sampled individuals. Individuals 1--5 are shades of blue and green, indicating that they 
+are from a single population.  Individuals 6--9 are from another population, (shades of red and orange), while 10--12
+are from a different population and are colored shades of violet. Thick vertical black lines separate the different 
+populations.  Markers are shown in rows. There are 6 markers on the first chromosome, 5 on the second, and 4 on the 
+third.
+Each chromosome is delimited by a thick, horizontal black line.
+The "a" and the "b" in the colored cells denote the first and second gene copies at each marker.
+{\bf (b)}~Examples of how the {\tt\footnotesize preserve\_individuals} and {\tt\footnotesize preserve\_haplotypes} options of the {\tt\footnotesize segments2markers()}
+function affect the permutation of 
+the genetic data of the original data set.  Panel headers show the option values while the colors and "a" and "b" values
+in the colored cells show the original position of each gene copy in the original sample.  Note that permutation is always
+done within the populations.  When {\tt\footnotesize preserve\_individuals = TRUE},  then just the 
+individuals in their entirety are permuted.  When {\tt\footnotesize preserve\_individuals = "BY\_CHROM"}, each pair
+of a single chromosome within an individual remains together when permuted.
+When {\tt\footnotesize preserve\_individuals = FALSE}, gene copies are freely permuted between individuals.
+The option {\tt\footnotesize preserve\_haplotypes = TRUE} causes each haplotype on a chromosome to be preserved
+while permuting. This option should only be used when the original data set is phased (the phasing is
+reflected in the input order of gene copies at each locus).
+}
+\begin{figure*}
+\includegraphics[width=\textwidth]{figures/permutations-plot.png}
+\caption[\permcap]{\permcap}
+\end{figure*}
+%%%%%%%
 
@@
 
-GSCRAMBLE REFERENCE GROUPS
-GSCRAMBLE PEDIGREES
-ADMIXTURE GSCRAMBLE VS EMPIRICAL
-FST OF EMPIRICAL AND GSCRAMBLE CORE POPULATIONS
+\gscramble{} REFERENCE GROUPS
+\gscramble{} PEDIGREES
+ADMIXTURE \gscramble{} VS EMPIRICAL
+FST OF EMPIRICAL AND \gscramble{} CORE POPULATIONS
 

### 2023-10-27 body-text-of-paper.tex
Edits on permutation stuff.

@@
 to the samples.  At the end of segment simulation, the genome of each sample
-is recorded as a mosaic of segments that derived from different founders.  In order
+is recorded as a mosaic of segments that derive from different founders.  In order
 to create a data set of markers in each sample, it is necessary only to identify the alleles
@@
 founders, \gscramble{} provides a variety of ways to permute alleles amongst
-individuals within samples from which the founders are drawn.  By using permutation,
-variability in the data is created, but without sampling with replacement, and hence
-not inducing RIPI.
+the collection of individuals from which the founders are drawn.  By using permutation,
+variability in the data is created, but, since it is done without sampling with replacement, it does
+not induce RIPI.
 
-To execute these permutations, it is necessary to have a matrix of genotypes for
-the individuals, and also, each population must belong to exactly one
-population.  Permutation of alleles then proceeds within populations, hence the frequencies
-of alleles from each population remain the same as those observed
-in the samples themselves. Following permutation, there remain individuals 1,2,\ldots,
-in each population, but the alleles that they carry will be different after permutation.  These
-individuals are assigned in order to be the founders from each population in the GSP. Subsequently,
+These permutations are executed upon the genotypes provided by the user.
+\gscramble{} accepts genetic data as a matrix with $N$ rows (one for each individual) and
+$2L$ columns (two for each locus). If the data have been phased, then the phased alleles from the
+first haplotype should occupy the
+first entry for each locus within an individual, while alleles phased together on the second hapltotype should occupy the
+second entry for each locus. 
+There can be several populations represented in the data set, but
+each individual must belong to exactly one
+population.
+
+Permutation of alleles proceeds within populations, such that
+the frequencies of alleles from each population remain the same as those observed
+in the samples themselves. Following permutation, there remain individuals
+in each population, but the alleles that they carry are different after permutation.  These
+individuals are assigned, in order, to be the founders from each population in the GSP. Subsequently,
 after genetic material is segregated to the samples, the alleles found on each genomic segment in those permuted
@@
 with large genomic data sets, with hundreds of thousands to millions of markers, since segregation
-through the pedigree is not done via by segments, and not on a locus-by-locus basis. The only operations
+through the pedigree is done by segments, and not on a locus-by-locus basis. The only operations
 involving all the alleles in all the individuals are the permutation steps and indexing into the segments,
@@
 which takes as input the segments segregated down a GSP, a matrix of sampled genotypes, and a specification
-of the group that each sampled individuals belongs to.
+of the group that each sampled individual belongs to.  The permutation performed when running
+{\footnotesize\tt segments2markers()} is controlled by two main options.  The first, {\tt\footnotesize preserve\_individuals},
+can take three different values.  If {\tt\footnotesize preserve\_individuals = TRUE}, then entire individuals are permuted,
+such that all the alleles within an individual remain together.  If {\tt\footnotesize preserve\_individuals = "BY\_CHROM"}, then
+the alleles within an individual found on a homologous pair of chromosomes remain together, but the different
+chromosome pairs of an individual are independently permuted into new positions in the data set.  This form of permutation
+maintains the LD between markers on the same chromosome, which may be desirable in some contexts.
+If {\tt\footnotesize preserve\_individuals = FALSE} then each gene copy within an individual is permuted separately,
+providing the maximal scrambling of the data.   The second option, {\tt\footnotesize preserve\_haplotypes}, can take two
+different values, {{\tt\footnotesize FALSE} (the default) and {\tt\footnotesize TRUE}.  This option, which keeps alleles together
+on the same haplotype, should be set to true only
+if the original genotypes are phased into separate haplotypes.
+
 %%%%%%%
@@
 different options.  {\bf (a)}~A picture
-representing an original data set of 12 diploids sampled from three populations and typed at 15 markers. 
+representing an original data set of $N=12$ diploids sampled from three populations and typed at $L=15$ markers. 
 Different colors represent different sampled individuals. Individuals 1--5 are shades of blue and green, indicating that they 
 are from a single population.  Individuals 6--9 are from another population, (shades of red and orange), while 10--12
-are from a different population and are colored shades of violet. Thick vertical black lines separate the different 
+are from a third population and are colored shades of violet. Thick vertical black lines separate the different 
 populations.  Markers are shown in rows. There are 6 markers on the first chromosome, 5 on the second, and 4 on the 
@@
 Each chromosome is delimited by a thick, horizontal black line.
-The "a" and the "b" in the colored cells denote the first and second gene copies at each marker.
+The "a" and the "b" in the colored cells denote the first and second gene copies within an individual at each marker.
+It is worth noting that this representation shows a $2L \times N$ matrix,
+which is the transpose of the $N\times 2L$ matrix used to supply genetic marker data to \gscramble{}.
 {\bf (b)}~Examples of how the {\tt\footnotesize preserve\_individuals} and {\tt\footnotesize preserve\_haplotypes} options of the {\tt\footnotesize segments2markers()}
@@
 The option {\tt\footnotesize preserve\_haplotypes = TRUE} causes each haplotype on a chromosome to be preserved
-while permuting. This option should only be used when the original data set is phased (the phasing is
-reflected in the input order of gene copies at each locus).
+while permuting. This option should only be used when the original data set is phased.
 }

### 2023-10-27 body-text-of-paper.tex
Make more line endings in intro paragraphs

@@
 
-Genetic clustering algorithms are a commonly used approach to describe genetic structure among populations, reflecting the cumulative effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.
-The ability of such algorithms to differentiate populations is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker set and sample sizes from the respective populations.
-Based on an initial description of genetic structure, ecologists are frequently interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.
-Genetic movement has many implications, such as influencing evolutionary processes, maintaining genetic variation through gene flow, and driving disease dynamics \citep{huestis2019windborne} and affecting patterns of invasive species expansion CITE.
+Genetic clustering algorithms are a commonly used approach to describe genetic structure among populations, reflecting the cumulative 
+effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.  The ability of such algorithms to differentiate populations 
+is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker 
+set and sample sizes from the respective populations.  Based on an initial description of genetic structure, ecologists are frequently 
+interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location 
+and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.  Genetic movement has many implications, 
+such as influencing evolutionary processes, maintaining genetic variation through gene flow, and driving disease dynamics 
+\citep{huestis2019windborne} and affecting patterns of invasive species expansion CITE.
 
-	In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of mixed ancestry---
+In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of 
+mixed ancestry---
 characterized as individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference}.
-	a trait often summarized as the proportion of an individual’s genome originating from each of $K$ subpopulations, sensu \citet{pritchard2000inference}.
-Various processes could contribute to such mixed associations.
-Most directly, individuals may represent the influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately representing contributions from multiple sampled populations.
-However, similar patterns may result from functionally different patterns of population structuring.
-For example, such patterns might be expected when clinal systems governed by isolation by distance are described as discrete genetic clusters.
-	NOTE I think a line here about exogenous immigrants might be worthwhile to set up 2 biological processes paired with a statistical process, something like END NOTE
-	Alternatively, individuals of mixed ancestry could reflect immigration from outlying populations (e.g., exogenous immigration), with true source populations not sufficiently
+a trait often summarized as the proportion of an individual’s genome originating from each of $K$ subpopulations, sensu 
+\citet{pritchard2000inference}.  Various processes could contribute to such mixed associations.  Most directly, individuals may represent the 
+influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately 
+representing contributions from multiple sampled populations.  However, similar patterns may result from functionally different patterns of 
+population structuring.  For example, such patterns might be expected when clinal systems governed by isolation by distance are described 
+as discrete genetic clusters.
+NOTE I think a line here about exogenous immigrants might be worthwhile to set up 2 biological processes paired with a statistical process, 
+something like END NOTE
+Alternatively, individuals of mixed ancestry could reflect immigration from outlying populations (e.g., exogenous immigration), with true source 
+populations not sufficiently
 represented in the sample to be among the characterized $K$ subpopulations.
@@
 as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure.
-	 The challenge for researchers then becomes correctly identifying the ecological and/or statistical processes by which observed complex ancestry patterns were created.
+ The challenge for researchers then becomes correctly identifying the ecological and/or statistical processes by which observed complex 
+ ancestry patterns were created.
 
@@
 observed allele frequencies (see CITE, CITE, CITE).
-However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed samples, and
+However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed 
+samples, and
 inflates the perceived ability to correctly assign individuals back to source populations. \citet{anderson2008improved}
@@
 all their ancestry from a single subpopulation.
-, a type of clustering application with the limitation that individuals are fully assigned to single subpopulations (ignoring the possibility of mixed ancestry).
+, a type of clustering application with the limitation that individuals are fully assigned to single subpopulations (ignoring the possibility of 
+mixed ancestry).
 We demonstrate, below, that similar artifacts occur when simulating data to
@@
 Beyond simply evaluating the resolution of discrete populations in a simulated framework, researchers may also be
-interested in how individuals of admixed ancestry are genetically characterized given the statistical power of the available dataset and the analysis methods used.
+interested in how individuals of admixed ancestry are genetically characterized given the statistical power of the available dataset and the 
+analysis methods used.
 Recognizing that connectivity among genetically distinct populations may be relatively rare, additional insight into the
@@
 BCX]).
-	However, the process of recombination during meiosis creates a distribution of expected associations to parental populations among advanced hybrid classes.
-	Treating linked loci as independent can lead to the false interpretation that the association of advanced hybrid classes to parental population occur strictly as discrete distributions.
-	Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of linkage among loci available for analysis.
+However, the process of recombination during meiosis creates a distribution of expected associations to parental populations among
+advanced hybrid classes.
+Treating linked loci as independent can lead to the false interpretation that the association of advanced hybrid classes to parental population 
+occur strictly as discrete distributions.
+Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of
+linkage among loci available for analysis.
 
-	In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without replacement, thus maintaining the genetic differences in the empirical dataset.
-	Further, by simulating individual genotypes based on user-defined pedigrees, \gscramble{} allows for the simulation of admixed genotypes of varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
+In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without 
+replacement, thus maintaining the genetic differences in the empirical dataset.
+Further, by simulating individual genotypes based on user-defined pedigrees, \gscramble{} allows for the simulation of admixed genotypes of 
+varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
 By integrating species-specific recombination rates, \gscramble{} simulates biologically informed genotypes that mirror empirical data.
@@
 $q_A$, the admixture fraction for population $A$, in  $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$. 
-We simulated these genotypes according to the "admixture model" used in {\em structure} \citep{pritchard2000inference}.
-Briefly, at each locus, independently, the origin
+We simulated these genotypes according to the "admixture model" used in {\em structure} \citep{pritchard2000inference}.  Briefly, at each 
+locus, independently, the origin
 of each gene copy was simulated to be from
@@
 We then added these $9n$ simulated individuals to the $N$ original individuals  from $A$ and $N$ from $B$
-and analyzed all the genotypes using ADMIXTURE with $K=2$. On each data set we performed a supervised analysis in which the original $N$ samples from each
+and analyzed all the genotypes using ADMIXTURE with $K=2$. On each data set we performed a supervised analysis in which the original 
+$N$ samples from each
 population were provided as learning samples, and also an
@@
 Because the simulations were done in a na\"{i}ve way that effectively assumes that the population
-allele frequencies are identical to those observed in the samples themselves, we hypothesized that ADMIXTURE would return estimates of $q_A$ that were
+allele frequencies are identical to those observed in the samples themselves, we hypothesized that ADMIXTURE would return estimates of 
+$q_A$ that were
 centered around the true values, even though samples $A$ and $B$ were drawn from the same

### 2023-11-03 body-text-of-paper.tex
Add another GSP.  Also, some tiny edits

@@
 
-Genetic clustering algorithms are a commonly used approach to describe genetic structure among populations, reflecting the cumulative 
+Genetic clustering algorithms are commonly used to describe genetic structure among populations, reflecting the cumulative 
 effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.  The ability of such algorithms to differentiate populations 
@@
 all their ancestry from a single subpopulation.
-, a type of clustering application with the limitation that individuals are fully assigned to single subpopulations (ignoring the possibility of 
-mixed ancestry).
 We demonstrate, below, that similar artifacts occur when simulating data to
@@
 
-The call the structure
+We call the structure
 we use to do this sort of cross-validation simulation of admixed individuals
@@
 In the \gscramble{} package there is a convenience function called {\footnotesize\tt create\_GSP()} that
-will create GSPs for samples up to and including second-generation backcrosses; however, \gscramble{} is designed
+will create GSPs for samples up to and including second-generation backcrosses (an example appears in Figure~\ref{fig:exampleGSPs}); 
+%%%%%
+\begin{figure}
+\newcommand{\gspcaptwo}{\footnotesize
+An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
+admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $BC_1$ in sample node~s9; and two $BC_2$'s in sample node~s10. 
+}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/AllThem.pdf}
+\end{center}
+\caption[\gspcaptwo]{\gspcaptwo}
+\label{fig:gsp1}
+\end{figure}
+%%%%
+however, \gscramble{} is designed
 to handle user-specified GSPs of arbitrary complexity.  When creating one's own GSP, it is useful

### 2023-11-05 body-text-of-paper.tex
matt's updates to methods and results

@@
 In the \gscramble{} package there is a convenience function called {\footnotesize\tt create\_GSP()} that
-will create GSPs for samples up to and including second-generation backcrosses (an example appears in Figure~\ref{fig:exampleGSPs}); 
+will create GSPs for samples up to and including second-generation backcrosses (an example appears in Figure~\ref{fig:gsp2}); 
 %%%%%
@@
 \caption[\gspcaptwo]{\gspcaptwo}
-\label{fig:gsp1}
+\label{fig:gsp2}
 \end{figure}
@@
 first haplotype should occupy the
-first entry for each locus within an individual, while alleles phased together on the second hapltotype should occupy the
+first entry for each locus within an individual, while alleles phased together on the second haplotype should occupy the
 second entry for each locus. 
@@
 
-\subsubsection*{Feral pigs in Missouri}
+\subsubsection*{Invasive wild pigs in Missouri}
 
-Genetic samples were collected from invasive wild pigs across a SPATIAL EXTENT in southeastern Missouri that were lethally removed through population control efforts conducted as a component of the Missouri Feral Hog Elimination Partnership
-by US Department of Agriculture – Animal Plant Health Inspection Services – Wildlife Services, Missouri Department of Conservation, and other cooperative state agencies CHECK.
-DNA was extracted from hair or tissue (collected from animals at the time of euthanasia) with commercially available magnetic bead recovery kits (MagMax DNA, Thermo Fisher Scientific).
-Genotyping was performed with GeneSeek’s Genomic Profiler for Porcine HD which provides 62,128 biallelic SNP loci that are mapped across the 18 autosomal chromosomes.
-We implemented standard genotype quality control metrics, specifically, we pruned genotypes with call rates <95\% and minor allele frequencies <5\%.
-We implemented linkage disequilibrium (LD) pruning using a window size of 50 loci and step size of 5 loci to remove markers above a linkage threshold of R2 > 0.5.
+Genetic samples were collected from 160 invasive wild pigs across a SPATIAL EXTENT in 
+southeastern Missouri that were lethally removed through population control efforts conducted 
+as a component of the Missouri Feral Hog Elimination Partnership
+by US Department of Agriculture – Animal Plant Health Inspection Services – Wildlife Services, Missouri Department of Conservation, 
+and other cooperative state agencies CHECK.
+DNA was extracted from hair or tissue (collected from animals at the time of euthanasia) with 
+commercially available magnetic bead recovery kits (MagMax DNA, Thermo Fisher Scientific).
+Genotyping was performed with GeneSeek’s Genomic Profiler for Porcine HD which provides 
+62,128 biallelic SNP loci that are mapped across the 18 autosomal chromosomes.
+We implemented standard genotype quality control metrics, specifically, we pruned genotypes with 
+call rates <95\% and minor allele frequencies <5\%.
+We implemented linkage disequilibrium (LD) pruning using a window size of 50 loci and step size 
+of 5 loci to remove markers above a linkage threshold of $R^2 > 0.5$.
+
+The 160 invasive wild pig samples represented 120 reference individuals from the core regions of 
+3 populations (40 individuals each from Pop1, Pop2, Pop14) and 40 individuals from the contact regions of these populations. 
+We used ADMIXTURE version 1.3.0 \citep{alexander2009fast} to characterize the genetic structure of the 160 samples (for $K = {1...6}$)
+and identify that indeed the 120 reference individuals represented 3 distinct genetic clusters.
+
+Using the 120 reference individuals, we sought to classify the admixture patterns of the 40 empirical samples 
+from the contact regions using simulated pedigrees from \gscramble{}. To address this objective we simulated 
+pedigrees to represent the likely classes of admixed individuals that would be present in the samples 
+from the contact regions of the 3 reference populations. The types of admixed individuals we considered were 
+all two-population combinations of $F_1$, $F_2$, $B_1$ and $B_2$. We also 
+simulated individuals from each of the founder populations. The first step of this workflow involved 
+simulating admixed individuals for 2 founder populations using one of the preset \gscramble{} 
+GSPs provided by the function {\footnotesize\tt create\_GSP()} for producing each of the types of admixed 
+individuals. While we could have simulated all 4 types of admixed individuals from a single GSP (Figure~\ref{fig:gsp2}), 
+different numbers of simulated individuals are produced for each hybrid class. Therefore, for the sake of the 
+computational efficiency from simulating the exact number of desired individuals, we chose to simulate from 
+a single GSP for each of the hybrid classes by setting the corresponding hybrid parameter in {\footnotesize\tt create\_GSP()} to {\footnotesize\tt TRUE}.
+We used the default settings of the {\footnotesize\tt plink\_map2rec\_rates()} to specify the recombination rates.
+The 120 simulated genotypes were written to PLINK .ped and .map files using {\footnotesize\tt gscramble2plink()}.
+We used PLINK version 1.9 \citep{purcell2007plink} to convert the genotypes to binary files and used these as input into ADMIXTURE for which we specified $K=3$ to produce ancestry values.
+We iteratively
+performed this workflow to produce 1,000 simulated genotypes and ancestry values for each of the 
+4 hybrid classes, as well as founder individuals, for each of the population combinations, for a total of 
+21,000 simulated individuals.
 
@@
 
+\subsection*{Empirical Examples}
+
+\subsubsection*{Rainbow/cutthroat trout hybrids in California}
+
+Boing.
+
+\subsubsection*{Invasive wild pigs in Missouri}
+
+SNP filtering resulted in 14,233 loci that were used for subsequent analysis.
+Using the 160 empirical samples, ADMIXTURE results supported the identification of 3 genetic clusters (SUPPLEMENTAL FIGURE CROSS VALIDATION?).
+Admixture *q* values clearly grouped individuals from each of the core populations into 3 clusters while the 40 individuals from the contact regions were highly admixed (SUPPLEMENTAL FIGURE ADMIXTURE).
+The
+$F_{ST}$ divergence from ADMIXTURE was: $F_{ST} = 0.243$ for Pop1-Pop2, $F_{ST} = 0.145$ for Pop1-Pop14 and $F_{ST} = 0.133$ for Pop2-Pop14.
+
 \section*{Discussion}
@@
 \section*{Acknowledgements}
-We thank etc. etc.   Contribution number  mHAVEAGAS-003
+We thank etc. etc.   Contribution number  mHAVEAGAS-003. 
+This work utilized the Alpine high performance computing resource at the University of Colorado Boulder.
+Alpine is jointly funded by the University of Colorado Boulder, the University of Colorado Anschutz, Colorado State University, and the National Science Foundation (award 2201538).

### 2023-11-16 body-text-of-paper.tex
updated some results, added some figures to folder

@@
 
-Genetic samples were collected from 160 invasive wild pigs across a SPATIAL EXTENT in 
+Genetic samples were collected from 160 invasive wild pigs in 
 southeastern Missouri that were lethally removed through population control efforts conducted 
@@
 We implemented standard genotype quality control metrics, specifically, we pruned genotypes with 
-call rates <95\% and minor allele frequencies <5\%.
+call rates $<95\%$ and minor allele frequencies $<5\%$.
 We implemented linkage disequilibrium (LD) pruning using a window size of 50 loci and step size 
@@
 from the contact regions of the 3 reference populations. The types of admixed individuals we considered were 
-all two-population combinations of $F_1$, $F_2$, $B_1$ and $B_2$. We also 
+all two-population combinations of $F_1$, $F_2$, $BC_1$ and $BC_2$. We also 
 simulated individuals from each of the founder populations. The first step of this workflow involved 
@@
 SNP filtering resulted in 14,233 loci that were used for subsequent analysis. 
-Using the 160 empirical samples, ADMIXTURE results supported the identification of 3 genetic clusters (SUPPLEMENTAL FIGURE CROSS VALIDATION?).
-Admixture *q* values clearly grouped individuals from each of the core populations into 3 clusters while the 40 individuals from the contact regions were highly admixed (SUPPLEMENTAL FIGURE ADMIXTURE).
-The
-$F_{ST}$ divergence from ADMIXTURE was: $F_{ST} = 0.243$ for Pop1-Pop2, $F_{ST} = 0.145$ for Pop1-Pop14 and $F_{ST} = 0.133$ for Pop2-Pop14.
+Using the 160 empirical samples, ADMIXTURE results supported the identification of 
+3 genetic clusters (SUPPLEMENTAL FIGURE). Admixture $q$ values clearly grouped individuals 
+from each of the core populations into 3 clusters while the 40 individuals from the contact regions were 
+highly admixed (SUPPLEMENTAL FIGURE). The
+$F_{ST}$ divergence from ADMIXTURE was: 
+$F_{ST} = 0.24$ for Pop1-Pop2, 
+$F_{ST} = 0.14$ for Pop1-Pop14 and 
+$F_{ST} = 0.13$ for Pop2-Pop14.
+
+The simulated hybrid classes produce 
+the expected range of admixture with $F_1$ and $F_2$ individuals having mean $q=0.5$, with $F_2$ individuals having 
+more variation. Backcrossed individuals have higher ancestry associated with the backcrossed population with $q$ 
+approximately $0.75$ ($BC_1$) and $0.875$ ($BC_2$) (Figure~\ref{fig:pig-sim-qs}).
+
+%%%
+\begin{figure*}
+\newcommand{\pigsimsqscap}{\footnotesize ADMIXTURE estimates of $q$ for $K=3$ of the simulated hybrid classes 
+($F_1$, $F_2$, $BC_1$ and $BC_2$) from all two-population combinations of the invasive wild pig genotypes. 
+Rows and columns provide the population combinations of the hybrids and for backcrossed hybrids the rows 
+specify the backcrossed population. Plots where the rows and columns specify the same population depict the 
+$q$ values for simulated non-hybrid (or "Core") individuals from the population.}
+\includegraphics[width=0.96\textwidth]{figures/simulated-pigs-Q-facet.pdf}
+\caption[\pigsimsqscap]{\pigsimsqscap}
+\label{fig:pig-sim-qs}
+\end{figure*}
+%%%
 
@@
 This work utilized the Alpine high performance computing resource at the University of Colorado Boulder. 
-Alpine is jointly funded by the University of Colorado Boulder, the University of Colorado Anschutz, Colorado State University, and the National Science Foundation (award 2201538).
+Alpine is jointly funded by the University of Colorado Boulder, the University of Colorado Anschutz, Colorado State University, 
+and the National Science Foundation (award 2201538).

### 2023-11-28 body-text-of-paper.tex
MGD added TJS's changes

@@
 and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.  Genetic movement has many implications,
-such as influencing evolutionary processes, maintaining genetic variation through gene flow, and driving disease dynamics 
-\citep{huestis2019windborne} and affecting patterns of invasive species expansion CITE.
+such as influencing evolutionary processes, maintaining genetic variation through gene flow, driving disease dynamics and affecting patterns 
+of invasive species expansion \citep{huestis2019windborne, estoup2010reconstructing}.
 
 In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of
-mixed ancestry---
-characterized as individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference}.
-a trait often summarized as the proportion of an individual’s genome originating from each of $K$ subpopulations, sensu 
-\citet{pritchard2000inference}.  Various processes could contribute to such mixed associations.  Most directly, individuals may represent the 
+mixed ancestry--- individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference}. 
+Various processes could contribute to such mixed associations.  Most directly, individuals may represent the
 influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately
@@
 as discrete genetic clusters.
-NOTE I think a line here about exogenous immigrants might be worthwhile to set up 2 biological processes paired with a statistical process, 
-something like END NOTE
 Alternatively, individuals of mixed ancestry could reflect immigration from outlying populations (e.g., exogenous immigration), with true source
@@
 A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to conduct
-parallel analyses on simulated data where the truth is known, structured to mirror the empirical data set
+parallel analyses on simulated data where the truth is known, yet structured to mirror the empirical data set
 as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
@@
 samples, and
-inflates the perceived ability to correctly assign individuals back to source populations. \citet{anderson2008improved}
-documented this type of artifact in the context of simulations to assess the power for
+inflates the perceived statistical power to correctly assign individuals back to source populations. 
+We refer to this phenomenon as resampling-induced power inflation (RIPI).
+\citet{anderson2008improved} used simulations to document this type of artifact in the context of 
 genetic stock identification (GSI), a type of clustering application in which individuals are constrained to have
 all their ancestry from a single subpopulation.
-We demonstrate, below, that similar artifacts occur when simulating data to
-assess how accurately admixture fractions can be estimated.  We refer to this phenomenon as resampling-induced
-power inflation, or RIPI for short.
-NOTE I would proposed holding on to the next sentence as we build the idea of mixed ancestry in the next paragraph END NOTE
-In this study, we additionally develop and illustrate an approach by which RIPI can be eliminated
-from simulations by constraining the genetic differences among simulated populations to match
-those of populations from which the samples were obtained.
+Similar statistical artifacts occur when simulating data to
+assess how accurately admixture fractions can be estimated.
 
@@
 In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without
-replacement, thus maintaining the genetic differences in the empirical dataset.
+replacement, thus maintaining the genetic differences in the empirical dataset and eliminating the effect of RIPI.
 Further, by simulating individual genotypes based on user-defined pedigrees, \gscramble{} allows for the simulation of admixed genotypes of
@@
 estimation in genetic stock identification (e.g., the cross-validation approach
-desribed in \citealt{moran2019bayesian}).
+described in \citealt{moran2019bayesian}).
 In the cross-validation approach, individuals are removed from the reference
@@
 feature of this approach is that no new genetic material is being created by
-sampling with replacement; however the genetic material of all individuals
-in the original samples is being used---either within the remaining individuals in the
+sampling with replacement; rather the genetic material of all individuals
+in the original sample is being used---either within the remaining individuals in the
 reference samples or within the individuals in the mixture.
@@
 Our scheme, described here, extends the traditional cross-validation approach
-by simulating the descent of chunks of chromosome through a user-specified
+by simulating the descent of chromosomal linkage blocks through a user-specified
 pedigree to create admixed individuals.  By imposing certain constraints
 on the pedigree, our framework ensures that all the genetic material
-within the original individuals is used to either create admixed individuals or to
-remain within reference individuals (i.e., those that are purely of one
+within the original individuals is used to either create admixed individuals or is retained 
+within reference individuals (i.e., those that are purely of one
 putative subopulation), and yet no new genetic material is
@@
 
-We call the structure
-we use to do this sort of cross-validation simulation of admixed individuals
+We refer to the structure used to implement this cross-validation simulation 
+of admixed individuals
 a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
@@
 At the end of this process, the samples simulated (as united pairs of gametes)
-into the sample nodes will contain all of the genetic material found in the founders (and
-no more than that)
-but that material will have been segregated and recombined according to the pedigree.  Thus the
+into the sample nodes will contain all of the genetic material (and no more) found among the founders but that 
+material will have been segregated and recombined according to the pedigree.  Thus the
 samples obtained from a GSP can be simulated with the expected admixture fractions
 for any hybrid category that can be specified by a pedigree.  They have also been simulated
-in a way that 1) respects physical linkage---each individual inherits segments of the genome
+in a way that 1) reflects physical linkage---each individual inherits segments of the genome
 simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
@@
 however, \gscramble{} is designed
-to handle user-specified GSPs of arbitrary complexity.  When creating one's own GSP, it is useful
-to keep in mind a handful of requirements for a GSP to be valid.
+to handle user-specified GSPs of arbitrary complexity.  When creating a unique GSP, 
+the following requirements are necessary for the GSP to be valid 
+and the simulation to execute successfully. 
 Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
@@
 
-These permutations are executed upon the genotypes provided by the user.
+These permutations are implemented with the genotypes provided by the user.
 \gscramble{} accepts genetic data as a matrix with $N$ rows (one for each individual) and
@@
 chromosome pairs of an individual are independently permuted into new positions in the data set.  This form of permutation
-maintains the LD between markers on the same chromosome, which may be desirable in some contexts.
+maintains the linkage disequilibrium (LDL between markers on the same chromosome, which may be desirable in some contexts.
 If {\tt\footnotesize preserve\_individuals = FALSE} then each gene copy within an individual is permuted separately,
 providing the maximal scrambling of the data. The second option, {\tt\footnotesize preserve\_haplotypes}, can take two
-different values, {{\tt\footnotesize FALSE} (the default) and {\tt\footnotesize TRUE}.  This option, which keeps alleles together
+different values, {{\tt\footnotesize FALSE} (the default) and {\tt\footnotesize TRUE}. 
+This option, which keeps alleles together
 on the same haplotype, should be set to true only
@@
 Genetic samples were collected from 160 invasive wild pigs in
-southeastern Missouri that were lethally removed through population control efforts conducted 
+southeastern Missouri, lethally removed through population control efforts conducted
 as a component of the Missouri Feral Hog Elimination Partnership
 by US Department of Agriculture – Animal Plant Health Inspection Services – Wildlife Services, Missouri Department of Conservation,
-and other cooperative state agencies CHECK.
+and other cooperative state agencies.
 DNA was extracted from hair or tissue (collected from animals at the time of euthanasia) with
@@
 62,128 biallelic SNP loci that are mapped across the 18 autosomal chromosomes.
-We implemented standard genotype quality control metrics, specifically, we pruned genotypes with 
+We implemented standard genotype quality control measures; specifically, we pruned loci with
 call rates $<95\%$ and minor allele frequencies $<5\%$.
@@
 The 160 invasive wild pig samples represented 120 reference individuals from the core regions of
-3 populations (40 individuals each from Pop1, Pop2, Pop14) and 40 individuals from the contact regions of these populations. 
-We used ADMIXTURE version 1.3.0 \citep{alexander2009fast} to characterize the genetic structure of the 160 samples (for $K = {1...6}$)
-and identify that indeed the 120 reference individuals represented 3 distinct genetic clusters.
-
-Using the 120 reference individuals, we sought to classify the admixture patterns of the 40 empirical samples 
+3 populations (40 individuals each informed by previous analyses) 
+and 40 individuals from the contact regions of these populations.
+We used ADMIXTURE version 1.3.0 \citep{alexander2009fast} to characterize the genetic structure 
+of the 160 samples (for $K = {1...6}$). Using the 120 reference individuals identified to represent population cores, 
+we sought to classify the admixture patterns of the 40 empirical samples
 from the contact regions using simulated pedigrees from \gscramble{}. To address this objective we simulated
@@
 from the contact regions of the 3 reference populations. The types of admixed individuals we considered were 
-all two-population combinations of $F_1$, $F_2$, $BC_1$ and $BC_2$. We also 
-simulated individuals from each of the founder populations. The first step of this workflow involved 
+$F_1$, $F_2$, $F_{1}BC_1$ and $F_{1}BC_2$ representing all two-population combinations 
+(Pop1 x Pop2, Pop1 x Pop3, Pop2 x Pop3, and reciprocally Pop2 x Pop1, Pop3 x Pop1, Pop3 x Pop2 as required 
+to simulate backcrossed classes). 
+We also simulated individuals from each of the founder populations. The first step of this workflow involved
 simulating admixed individuals for 2 founder populations using one of the preset \gscramble{}
 GSPs provided by the function {\footnotesize\tt create\_GSP()} for producing each of the types of admixed
-individuals. While we could have simulated all 4 types of admixed individuals from a single GSP (Figure~\ref{fig:gsp2}), 
+individuals. Although we could have simulated all 4 types of admixed individuals from a single GSP (Figure~\ref{fig:gsp2}),
 different numbers of simulated individuals are produced for each hybrid class. Therefore, for the sake of the
@@
 21,000 simulated individuals.
-
-\gscramble{} REFERENCE GROUPS
-\gscramble{} PEDIGREES
-ADMIXTURE \gscramble{} VS EMPIRICAL
-FST OF EMPIRICAL AND \gscramble{} CORE POPULATIONS
 
@@
 \section*{Acknowledgements}
-We thank etc. etc.   Contribution number  mHAVEAGAS-003. 
+Research was supported in part by the USDA-APHIS-National Feral Swine Damage 
+Management Program. We are grateful for the numerous Wildlife Services and Missouri 
+Department of Conservation employees that contributed to genetic sampling of wild pigs 
+within southeastern Missouri. Writing of this paper was facilitated by the Mobile High Altitude Venue for Ecological Analysis, 
+Genetics and Statistics, through which ECA, MGD, and TJS were scientists-in-residence on location in 
+Cheyenne, WY for 3 days in October 2023. This is contribution number mHAVEAGAS-003.
 This work utilized the Alpine high performance computing resource at the University of Colorado Boulder.

### 2023-11-28 body-text-of-paper.tex
moved extra figures to supplement directory, if those even end up being used. added new figure of simulations

@@
 $F_{ST} = 0.14$ for Pop1-Pop14 and
-$F_{ST} = 0.13$ for Pop2-Pop14.
-
-The simulated hybrid classes produce
+$F_{ST} = 0.13$ for Pop2-Pop14. The simulated hybrid classes produce
 the expected range of admixture with $F_1$ and $F_2$ individuals having mean $q=0.5$, with $F_2$ individuals having
 more variation. Backcrossed individuals have higher ancestry associated with the backcrossed population with $q$
-approximately $0.75$ ($BC_1$) and $0.875$ ($BC_2$) (Figure~\ref{fig:pig-sim-qs}).
+approximately $0.75$ ($BC_1$) and $0.875$ ($BC_2$) (Figure~\ref{fig:pig-sim-qs}). MATT NEEDS TO AUGMENT AND 
+REVISE THE RESULTS
 
@@
 \begin{figure*}
-\newcommand{\pigsimsqscap}{\footnotesize ADMIXTURE estimates of $q$ for $K=3$ of the simulated hybrid classes
-($F_1$, $F_2$, $BC_1$ and $BC_2$) from all two-population combinations of the invasive wild pig genotypes.
-Rows and columns provide the population combinations of the hybrids and for backcrossed hybrids the rows
-specify the backcrossed population. Plots where the rows and columns specify the same population depict the
-$q$ values for simulated non-hybrid (or "Core") individuals from the population.}
-\includegraphics[width=0.96\textwidth]{figures/simulated-pigs-Q-facet.pdf}
+\newcommand{\pigsimsqscap}{\footnotesize MATT NEEDS TO ADD CAPTION.}
+\includegraphics[width=0.96\textwidth]{figures/simulation-density-multipanel.pdf}
 \caption[\pigsimsqscap]{\pigsimsqscap}

### 2023-11-28 body-text-of-paper.tex
matt's updates to results

@@
 Using the 160 empirical samples, ADMIXTURE results supported the identification of
-3 genetic clusters (SUPPLEMENTAL FIGURE). Admixture $q$ values clearly grouped individuals
-from each of the core populations into 3 clusters while the 40 individuals from the contact regions were
-highly admixed (SUPPLEMENTAL FIGURE). The
-$F_{ST}$ divergence from ADMIXTURE was:
-$F_{ST} = 0.24$ for Pop1-Pop2,
-$F_{ST} = 0.14$ for Pop1-Pop14 and
-$F_{ST} = 0.13$ for Pop2-Pop14. The simulated hybrid classes produce
-the expected range of admixture with $F_1$ and $F_2$ individuals having mean $q=0.5$, with $F_2$ individuals having
-more variation. Backcrossed individuals have higher ancestry associated with the backcrossed population with $q$
-approximately $0.75$ ($BC_1$) and $0.875$ ($BC_2$) (Figure~\ref{fig:pig-sim-qs}). MATT NEEDS TO AUGMENT AND 
-REVISE THE RESULTS
+$K=3$ genetic clusters grouped individuals from each of the core populations into 3 clusters 
+while the 40 individuals from the contact regions were highly admixed. The
+$F_\mathrm{ST}$ divergence from ADMIXTURE was:
+$F_\mathrm{ST} = 0.24$ for Pop1-Pop2,
+$F_\mathrm{ST} = 0.14$ for Pop1-Pop3 and
+$F_\mathrm{ST} = 0.13$ for Pop2-Pop3.
+
+The simulated hybrid classes produce
+the expected range of admixture with $F_1$ and $F_2$ individuals having mean $Q=0.5$, with $F_2$ individuals having
+more variation (Figure~\ref{fig:pig-sim-qs}A). 
+Backcrossed individuals have higher ancestry associated with the backcrossed population with mean $Q$
+approximately $0.75$ ($BC_1$) and $0.87$ ($BC_2$) (Figure~\ref{fig:pig-sim-qs}A). Notably, the simulated 
+$F_1$ hybrids have the most variation in the distribution of the $Q$ values associated with each of the genetic clusters. 
+When the estimated Q values for the $F_1$ hybrids are separated by the population combinations, 
+$F_1$ hybrids for Pop1-Pop2 have the least variation around $q=0.5$ (standard deviation 
+= 0.008), while Pop1-Pop3 and Pop2-Pop3 
+have higher variation (standard deviation = 0.011-0.016) (Figure~\ref{fig:pig-sim-qs}B). Also, all $Q$ 
+distributions for $F_1$ hybrids, regardless of population combination, center around 
+$0.50$ except for the two distributions 
+of Pop3, which have a mean of $Q=0.49$ (Figure~\ref{fig:pig-sim-qs}B).
+The differences among these 
+distributions of $Q$ values for the simulated data reflect the genetic differences observed 
+in the empirical data, specifically, that genetic differentiation is highest between Pop1 and Pop2 and that 
+Pop14 is the most admixed population with ancestry contributions from both Pop1 and Pop2.
 
@@
 \begin{figure*}
-\newcommand{\pigsimsqscap}{\footnotesize MATT NEEDS TO ADD CAPTION.}
+\newcommand{\pigsimsqscap}{\footnotesize (A) The density of the estimated Q values from ADMIXTURE for 
+the simulated runs are 
+grouped across all population-combinations for each hybrid class and colored by the $K$ genetic cluster 
+(e.g., the Pop1 density plot for the $F_{1}BC_1$ plots the Pop1 genetic cluster for both 
+the Pop1-Pop2 and Pop1-Pop3 combinations). The density plots across the $K$ 
+genetic clusters are mostly similar with the most variation among the genetic clusters evident in 
+the $F_1$ plot. (B) Plotting the density of the estimated Q values from ADMIXTURE for the $F_1$ hybrids, 
+separated by the population combinations, reveals the least variance in estimated Q values for 
+Pop1-Pop2 compared to Pop1-Pop3 and Pop2-Pop3. Note that these population combinations group the 
+reciprocal population combination in each plot (e.g., "Pop1_Pop2" has simulations 
+from both Pop1-Pop2 and Pop2-Pop1).}
 \includegraphics[width=0.96\textwidth]{figures/simulation-density-multipanel.pdf}

### 2023-11-28 body-text-of-paper.tex
sampling information

@@
 
-Genetic samples were collected from 160 invasive wild pigs in
-southeastern Missouri, lethally removed through population control efforts conducted
-as a component of the Missouri Feral Hog Elimination Partnership
-by US Department of Agriculture – Animal Plant Health Inspection Services – Wildlife Services, Missouri Department of Conservation,
-and other cooperative state agencies.
+Genetic samples were collected from 160 invasive wild pigs in southeastern
+Missouri that were lethally removed through population control efforts
+conducted as a component of the Missouri Feral Hog Elimination Partnership
+by US Department of Agriculture – Animal Plant Health Inspection Services –
+Wildlife Services, Missouri Department of Conservation, and other cooperative
+federal and state agencies.
 DNA was extracted from hair or tissue (collected from animals at the time of euthanasia) with

### 2023-11-29 body-text-of-paper.tex
Backslash-escape a rogue underscore

@@
 Pop1-Pop2 compared to Pop1-Pop3 and Pop2-Pop3. Note that these population combinations group the
-reciprocal population combination in each plot (e.g., "Pop1_Pop2" has simulations
+reciprocal population combination in each plot (e.g., "Pop1\_Pop2" has simulations
 from both Pop1-Pop2 and Pop2-Pop1).}

### 2023-12-13 body-text-of-paper.tex
Some small additions

@@
 inflates the perceived statistical power to correctly assign individuals back to source populations.
-We refer to this phenomenon as resampling-induced power inflation (RIPI).
+We refer to this phenomenon as resampling-induced, spurious power inflation (RISPI).
 \citet{anderson2008improved} used simulations to document this type of artifact in the context of
@@
 In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without
-replacement, thus maintaining the genetic differences in the empirical dataset and eliminating the effect of RIPI.
+replacement, thus maintaining the genetic differences in the empirical dataset and eliminating the effect of RISPI.
 Further, by simulating individual genotypes based on user-defined pedigrees, \gscramble{} allows for the simulation of admixed genotypes of
@@
 we demonstrate a case in which using a standard sampling-with-replacement
-approach induces RIPI when estimating admixture fractions.
+approach induces RISPI when estimating admixture fractions.
 To construct this simple example, we used the R programming language \citep{Rcoreteam} to
@@
 $q_A$ that were
-centered around the true values, even though samples $A$ and $B$ were drawn from the same
-population.
-We further expected that ADMIXTURE would return more accurate estimates of
-$q_A$ when $N$ was smaller and the number of loci was larger.
+centered around the simulated values, even though samples $A$ and $B$ were drawn from the same
+population. In other words, we expected to observe RISPI.
+We further expected that ADMIXTURE would return more accurate estimates of the simulated
+$q_A$ values when $N$ was smaller and the number of loci was larger.
 We summarized the results by plotting boxplots of the ADMIXTURE-inferred $q_A$ values,
@@
 We refer to the structure used to implement this cross-validation simulation
-of admixed individuals
+of admixed individuals as
 a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
@@
 Squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
-individuals 1--4.  The inverted triangles above each of those founders denotes, in color
+individuals 1--4.  The  triangles above each of those founders denotes, in color
 and text, the population of origin of each of the two genomes within each founder.
@@
 Each non-founder node $x$ will have exactly two edges coming into it from above
-(from its two parents), we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$.
+(from its two parents), and we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$.
 For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
-A non-founder individual may have more or less than two edges coming downward
+A founder individual may have, at most, 2 edges coming downward out of it, while a non-founder individual may have more or less than two edges coming downward
 out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
 non-founder daughter nodes, we denote the number of gametes segregated
-through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example,
+through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example, at
 non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
@@
 node 7. Any non-founder individual node can have a maximum of one edge leading
-to a sample node.  We denote the number of samples emanating from node $x$
-node as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
+to a sample node.  We denote the number of samples emanating from node $x$ as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
 created from a single simulation on this GSP.
@@
 These gametes are united within the founder individual and then
-following recombination between each chromosome
+recombination occurs between each chromosome
 with probability $\frac{1}{2}$ and within each chromosome according to
-a user-specified recombination map,
- one gamete is assorted down each edge from the founder node to its daughter.
+a user-specified recombination map to create two new gametes, which are
+recombined version of the original ones.
+These two gametes are then segregated at random down the edge (or edges) from the founder node.
+Upon each edge, $i$ from founder node $f$, $G^-_{f,i}$ gametes are segregated.
 \item Each non-founder node, $x$, will have exactly two edges coming
@@
 At the end of this process, the samples simulated (as united pairs of gametes)
-into the sample nodes will contain all of the genetic material (and no more) found among the founders but that
+into the sample nodes will contain all of the genetic material (and no more) found among the founders, but that
 material will have been segregated and recombined according to the pedigree.  Thus the
@@
 simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
-genetic material is entirely without replacement so it will not induce RIPI; and 3) no genetic material has been lost---all the genetic material of the founders
-is represented in the samples, maximizing the number of individuals and markers available to
+genetic material is entirely without replacement so it will not induce RISPI; and 3) no genetic material has been lost---all the genetic material of the founders
+is represented in the samples, maximizing the number of simulated individuals and markers available to
 use in downstream analyses.
@@
 variability in the data is created, but, since it is done without sampling with replacement, it does
-not induce RIPI.
+not induce RISPI.
 
@@
 {\footnotesize\tt segments2markers()} is controlled by two main options.  The first, {\tt\footnotesize preserve\_individuals},
-can take three different values.  If {\tt\footnotesize preserve\_individuals = TRUE}, then entire individuals are permuted,
+can take three different values.  If {\tt\footnotesize preserve\_individuals = TRUE}, then entire individuals are the unit that gets permuted within the
+data set,
 such that all the alleles within an individual remain together.  If {\tt\footnotesize preserve\_individuals = "BY\_CHROM"}, then
@@
 chromosome pairs of an individual are independently permuted into new positions in the data set.  This form of permutation
-maintains the linkage disequilibrium (LDL between markers on the same chromosome, which may be desirable in some contexts.
+maintains the linkage disequilibrium (LD) between markers on the same chromosome, which may be desirable in some contexts.
 If {\tt\footnotesize preserve\_individuals = FALSE} then each gene copy within an individual is permuted separately,
@@
 This option, which keeps alleles together
-on the same haplotype, should be set to true only
-if the original genotypes are phased into separate haplotypes.
+on the same haplotype, should be set to {\tt TRUE} only
+if the original genotypes have been phased into separate haplotypes.
+Fig.~\ref{fig:perms} depicts, graphically, the effect of using different settings of
+{\tt\footnotesize preserve\_individuals} and 
+{\tt\footnotesize preserve\_haplotypes}, together.
 
@@
 \caption[\permcap]{\permcap}
+\label{fig:perms}
 \end{figure*}
@@
 \subsection*{Empirical Examples}
+
+We illustrate the use of \gscramble{} in two examples.  The first involves
+simulating a small number ($<\,$70) of markers that are fixed, or nearly
+fixed, for alternate alleles between two species of trout: rainbow trout ({\em 
+Oncorhynchus mykiss}) and cutthroat trout ({\em O. clarkii}). The goal of the 
+simulation study is to determine how accurately individuals of different hybrid
+categories between the species can be identified, and whether the use of linked 
+markers deflates the estimates of uncertainty to an unacceptable degree.  The 
+second example involves many more markers ($\approx$14,000) and 
+investigates the accuracy for
+identifying individuals admixed between closely related groups of feral
+pigs in Missouri.
 

### 2024-01-08 body-text-of-paper.tex
Eric edited the introduction a bit.

@@
 In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of
-mixed ancestry--- individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference}.
-Various processes could contribute to such mixed associations.  Most directly, individuals may represent the
+mixed ancestry---individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference}.
+Various processes could contribute to the appearance of mixed ancestry in a clustering analysis.  Most directly, individuals may represent the
 influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately
@@
 A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to conduct
-parallel analyses on simulated data where the truth is known, yet structured to mirror the empirical data set
+parallel analyses on data that have been simulated---so the truth is known---so as to mirror the empirical data set
 as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
 Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the
-observed allele frequencies (see CITE, CITE, CITE).
+observed allele frequencies (see, for example, \citealt{nielsen2006hybridlab,kinziger2008hybridization}).
 However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed
@@
 BCX]).
-However, the process of recombination during meiosis creates a distribution of expected associations to parental populations among
-advanced hybrid classes.
-Treating linked loci as independent can lead to the false interpretation that the association of advanced hybrid classes to parental population
-occur strictly as discrete distributions.
-Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of
+Evaluating the results of genetic clustering algorithms in the context of data that are
+expected to arrive by such a process of hybridization requires simulation of
+explicit hybrids.  Such simulations are the purpose of the software {\sc hybridlab} 
+\citep{nielsen2006hybridlab}.
+
+{\sc hybridlab} was conceived at a time when
+the data sets used in molecular ecology universally included very few markers, such that
+it was reasonable to assume the markers were unlinked, and, hence, independently segregating.
+Today, however, with data sets of thousands of markers, that assumption is badly violated---alleles
+on the same chromosome are likely to be segregated together into the next generation.
+Simulating linked loci as independent creates simulated hybrids, beyond the $F1$ category, with 
+artificially low variance in admixture fraction, leading to the false interpretation that different
+hybrid categories are more easily distinguished than they are. Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of
 linkage among loci available for analysis.
+This was done recently to assess the power of a SNP panel for identifying descendants of escaped, farm-raised Atlantic salmon \citep{wringe2019development}; however the method used (described at https://eriqande.github.io/SalarHybPower/) was developed in an {\em ad hoc} fashion and was not easily extended to
+complex hybridization scenarios, and was not at all user-friendly.
 
 In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without
-replacement, thus maintaining the genetic differences in the empirical dataset and eliminating the effect of RISPI.
-Further, by simulating individual genotypes based on user-defined pedigrees, \gscramble{} allows for the simulation of admixed genotypes of
-varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
+replacement, thus maintaining genetic differences between the clusters commensurate to those in the empirical dataset, and eliminating the effect of RISPI.
+Further, by simulating individual genotypes based on user-defined pedigrees of arbitrary complexity, \gscramble{} allows for the simulation of admixed genotypes resulting from any conceivable category of hybrid, while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
 By integrating species-specific recombination rates, \gscramble{} simulates biologically informed genotypes that mirror empirical data.
-We develop and illustrate \gscramble{} with use of single nucleotide polymorphic (SNP) genotypes.
+We develop and illustrate \gscramble{} with use of single nucleotide polymorphic (SNP) genotypes, though it is equally applicable to data sets with hundreds or thousands of
+microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites
+\citep{zhan2017megasat}.
 
@@
 
-Our approach for simulating admixed individuals without inflating perceived power for
-the estimation of their admixture fractions is motivated by the traditional cross-validation
+Our approach for simulating admixed individuals without creating RISPI is motivated by the traditional cross-validation
 approach for evaluating the power of mixture proportion
@@
 
-Boing.
+Genotypes for
 
@@
 
+Also, we need to discuss the fact that, because each time a gamete is segregated in the
+pedigree, its complement is also segregated, this is an example of antithetical variate sampling
+in Monte Carlo.
+
 \section*{Acknowledgements}

### 2024-01-11 body-text-of-paper.tex
Added newhybs figure.

@@
 linkage among loci available for analysis.
-This was done recently to assess the power of a SNP panel for identifying descendants of escaped, farm-raised Atlantic salmon \citep{wringe2019development}; however the method used (described at https://eriqande.github.io/SalarHybPower/) was developed in an {\em ad hoc} fashion and was not easily extended to
+This was done recently to assess the power of a SNP panel for identifying descendants of escaped, farm-raised Atlantic salmon \citep{wringe2019development}; however the method used (described at https://eriqande.github.io/SalarHybPower/) was developed in an {\em ad hoc} fashion, was not easily extended to
 complex hybridization scenarios, and was not at all user-friendly.
@@
 for each combination of $L$, $N$, and $n$.
+
+NOTE, WE ALSO DID THESE SIMULATIONS USING GSCRAMBLE AND WE SHOULD NOTE
+HOW THOSE CAME OUT.
+
+\subsection*{Physical Linkage Matters}
+
+ECA STILL NEEDS TO DESCRIBE AND INCLUDE THE SIMULATIONS THAT HE DID ON THIS.
+THAT WORK IS ALL DONE, I JUST NEED TO PUT IT IN HERE, ALONG WITH THE FIGURE.
 
@@
 a user-specified recombination map to create two new gametes, which are
-recombined version of the original ones.
+recombined versions of the original ones.
 These two gametes are then segregated at random down the edge (or edges) from the founder node.
@@
 
-Genotypes for
+We used the archived data from \citet{rizza2023limited}, drawing upon their 65 markers
+that are fixed or are nearly fixed for alternate alleles between rainbow trout and cutthroat trout.
+They designated 634 fish as pure cutthroat trout and 213 as pure steelhead.  We used the
+genotypes of these individuals as input to \gscramble{}.
+
+\citeauthor{rizza2023limited} mapped the 65 markers to the Omyk\_v1.0 genome assembly (RefSeq 
+Accession: GCF\_002163495.1), so we used
+the linkage map described in \citep{pearse2019sex} that was prepared by linkage mapping 
+of $\approx$57,000 markers described in \citet{palti2015development} to Omyk\_v1.0. We mapped the $\approx$57,000 markers to Omyk\_v1.0, and then resolved inconsistencies to produce a map
+of 32,923 markers throughout the genome and rendered it to plink format for use with \gscramble{}.
+
+In the original paper, \citet{rizza2023limited} identified 15 $F_1$s,
+1 $F_2$,
+6 cutthroat backcrosses ($BC1$s) and 
+6 second-generation cutthroat backcrosses ($BC2$s),
+and 1 fish of an unknown hybrid category that may have been the
+product of a $BC2$ crossed with a $BC1$.   In our simulations, we sought to simulate
+a similar collection of hybrid individuals, so, for each replicate, we used
+\gscramble{} to simulate 16 $F_1$s, 4 $F_2$s, 6 $BC1$s, 4 $BC2$s, 4 $BC3$s,  4 individuals
+that were the products of a $BC1$ mating with a $BC2$, and finally, 4 individuals that were simulated pure cutthroat. Individuals not used as founders
+in \gscramble{} were retained as pure cutthroat and pure steelhead.  The goal of the simulation
+was to assess how well these different hybrid categories can be distinguished from one another
+when using either all 65 markers, or just the 30 markers occurring on different chromosome arms
+used by  \citet{rizza2023limited}.
+
+HEY! I SHOULD ALSO DO A SET OF SIMULATIONS IN WHICH THE 65 MARKERS ARE
+SIMULATED AS IF THEY ARE INDEPENDENTLY SEGREGATING---I THINK THAT MEANS
+ASSIGNING EACH TO ITS OWN CHROMOSOME.  BUT MAYBE WORRY ABOUT THAT
+LATER.
+
+We performed 500 replicate simulations with \gscramble{} using either 30 or 65 markers.  Each
+simulated data set was then analyzed using {\sc NewHybrids} \citep{anderson2002model}.
+Individuals not used as founders were given the ``s'' and ``z'' flags, indicating that they were considered training samples.
+The remaining 46 individuals were treated as possible hybrids and their genetic data used to calculate the posterior probability that they belonged to one of 10 hybrid categories as shown in Table~\ref{tab:newhybcats}, using 25,000 MCMC sweeps and discarding the first 5,000 as burn-in.
+%%%%%%%%%%%%
+\begin{table}
+\caption{The 10 hybrid categories considered for inference with {\sc NewHybrids} and the
+expected frequency of their genotypes having 2 ($f(C,C)$), 1 ($f(C,S)$) or 0 ($f(S,S)$) copies of ancestry from cutthroat trout.}
+\label{tab:newhybcats}
+{\small
+\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrr}
+\hline\hline
+Category	& $f(C,C)$ &	$f(C,S)$ & 	$f(S,S)$ \\ \hline
+\input{inputs/newhybs-cats.tsv}
+\end{tabular*}
+}
+\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
+\end{table}
+%%%%%%%%%%%%
 
@@
 
-Boing.
+PRETTY MUCH DONE HERE EXCEPT I WANT TO SIMULATE THINGS
+UNLINKED.
+%%%%%%%%%%
+\begin{figure}
+\newcommand{\nhbpcap}{\footnotesize Gonna need a caption}
+\includegraphics[width=\columnwidth]{figures/newhybs-texed-crop.pdf}
+\caption[\nhbpcap]{\nhbpcap}
+\label{fig:newhyb-sims-barplot}
+\end{figure}
+%%%%%%%%%%
 
@@
 Department of Conservation employees that contributed to genetic sampling of wild pigs
-within southeastern Missouri. Writing of this paper was facilitated by the Mobile High Altitude Venue for Ecological Analysis,
+within southeastern Missouri.
+We thank Yniv Palti for providing flanking sequences and the linkage map for 57K markers for making a recombination map for the cutthroat/rainbow trout example.
+Writing of this paper was facilitated by the Mobile High Altitude Venue for Ecological Analysis,
 Genetics and Statistics, through which ECA, MGD, and TJS were scientists-in-residence on location in

### 2024-01-11 body-text-of-paper.tex
Another big slug of additions.

@@
 
+%%%%%%%%%%%
+\begin{figure}
+\newcommand{\gspcapone}{\footnotesize
+A genomic simulation pedigree (GSP) for simulating $F_2$ individuals.
+Squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
+individuals 1--4.  The  triangles above each of those founders denotes, in color
+and text, the population of origin of each of the two genomes within each founder.
+Thus, individuals~1 and~2 are founders purely from population~$A$, and~3 and~4
+are purely from population~$B$.  The pink hexagon labeled s7 is the sample node.
+It represents samples that are taken from non-founder node~7, which, from the structure
+of the pedigree, clearly represents an $F_2$ hybrid between populations $A$ and $B$.  The red numbers
+on edges above and leading into each non-founder node indicate the number of
+gametes that get segregated from the parent node (at the top of the edge) to the
+daughter node (at the bottom of the edge).
+This emphasizes that,
+in dropping genes through a GSP, in contrast
+to a traditional pedigree, each parent
+node may segregate multiple gametes to a daughter node.  See the text for
+further explanation.
+Each non-founder node $x$ will have exactly two edges coming into it from above
+(from its two parents), and we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$.
+For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
+A founder individual may have, at most, 2 edges coming downward out of it, while a non-founder individual may have more or less than two edges coming downward
+out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
+non-founder daughter nodes, we denote the number of gametes segregated
+through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example, at
+non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
+Finally, the purple number adjacent to the edge leading into the
+sample node indicates how many sample individuals are drawn from individual
+node 7. Any non-founder individual node can have a maximum of one edge leading
+to a sample node.  We denote the number of samples emanating from node $x$ as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
+created from a single simulation on this GSP.
+}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/F2.pdf}
+\end{center}
+\caption[\gspcapone]{\gspcapone}
+\label{fig:gsp1}
+\end{figure}
+%%%%%%%%%%%%
+%%%%%
+\begin{figure}
+\newcommand{\gspcaptwo}{\footnotesize
+An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
+admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $\BC_1$ in sample node~s9; and two $\BC_2$'s in sample node~s10.
+}
+\begin{center}
+\includegraphics[width=\columnwidth]{figures/AllThem.pdf}
+\end{center}
+\caption[\gspcaptwo]{\gspcaptwo}
+\label{fig:gsp2}
+\end{figure}
+%%%%
+
 {\sc hybridlab} was conceived at a time when
@@
 
-\subsection*{An introductory motivating example}
+\subsection*{Two introductory motivating examples}
 
 Before we proceed to a description of the methodology used by \gscramble{},
-we demonstrate a case in which using a standard sampling-with-replacement
-approach induces RISPI when estimating admixture fractions.
+we demonstrate two important points through two simple simulations.  In the first,
+we show that using a standard sampling-with-replacement
+approach induces RISPI when estimating admixture fractions. In the second, we show
+the effect of physical linkage on the distribution of admixture fractions of hybrid individuals,
+which vividly shows why it is not acceptable with large numbers of markers, to simulate
+the genotypes of hybrid individuals assuming no physical linkage.
+
+\subsubsection*{RISPI when estimating admixture fractions}
+
 To construct this simple example, we used the R programming language \citep{Rcoreteam} to
@@
 
-\subsection*{Physical Linkage Matters}
+\subsubsection*{Physical Linkage in Hybrid Individuals}
 
-ECA STILL NEEDS TO DESCRIBE AND INCLUDE THE SIMULATIONS THAT HE DID ON THIS.
-THAT WORK IS ALL DONE, I JUST NEED TO PUT IT IN HERE, ALONG WITH THE FIGURE.
+We first simulated the genotypes of hybrids between two species, $a$ and $b$,
+considering $F_2$'s and first- through
+fourth-generation backcrosses in each direction, erroneously treating each
+locus as if it were independently segregating (as done, for example, in {\sc HybridLab}). Thus,
+each locus independently is simulated to have a genotype of 2, 1, or 0 gene copies originating from species $a$ [{\em e.g.}~genotypes of $(a,a)$, $(a,b)$, or $(b,b)$] according to the expected
+probability of each genotype given the hybrid category (see \citealt{anderson2002model}).
+48,000 hybrids were simulated at each number of loci, $L$, in $\{10^2, 10^3, 10^4, 10^5\}$.
+The admixture fraction of each simulated individual was recorded as the fraction of all $2L$
+gene copies originating from species $a$.
+
+We also simulated the genomes of 48,000 individuals of each
+hybrid category using \gscramble{}'s {\footnotesize\tt segregate()} function, as described
+below, with number of chromosomes and sex-averaged recombination rates
+for pigs, as estimated by from \citet{tortereau2012high}. In this case, we recorded the
+admixture fraction of each simulated individual as the fraction of all genome segments
+that originated from species~$a$.  This provides the distribution of the true admixture fraction
+when physical linkage of markers is not ignored.
+
+All the admixture fraction distributions were summarized by plotting them as histograms
+so they could be compared easily.
 
@@
 uses terminology that is defined in the figure caption.
-%%%%%%%%%%%
-\begin{figure}
-\newcommand{\gspcapone}{\footnotesize
-A genomic simulation pedigree (GSP) for simulating $F_2$ individuals.
-Squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
-individuals 1--4.  The  triangles above each of those founders denotes, in color
-and text, the population of origin of each of the two genomes within each founder.
-Thus, individuals~1 and~2 are founders purely from population~$A$, and~3 and~4
-are purely from population~$B$.  The pink hexagon labeled s7 is the sample node.
-It represents samples that are taken from non-founder node~7, which, from the structure
-of the pedigree, clearly represents an $F_2$ hybrid between populations $A$ and $B$.  The red numbers
-on edges above and leading into each non-founder node indicate the number of
-gametes that get segregated from the parent node (at the top of the edge) to the
-daughter node (at the bottom of the edge).
-This emphasizes that,
-in dropping genes through a GSP, in contrast
-to a traditional pedigree, each parent
-node may segregate multiple gametes to a daughter node.  See the text for
-further explanation.
-Each non-founder node $x$ will have exactly two edges coming into it from above
-(from its two parents), and we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$.
-For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
-A founder individual may have, at most, 2 edges coming downward out of it, while a non-founder individual may have more or less than two edges coming downward
-out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
-non-founder daughter nodes, we denote the number of gametes segregated
-through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example, at
-non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
-Finally, the purple number adjacent to the edge leading into the
-sample node indicates how many sample individuals are drawn from individual
-node 7. Any non-founder individual node can have a maximum of one edge leading
-to a sample node.  We denote the number of samples emanating from node $x$ as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
-created from a single simulation on this GSP.
-}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/F2.pdf}
-\end{center}
-\caption[\gspcapone]{\gspcapone}
-\label{fig:gsp1}
-\end{figure}
-%%%%%%%%%%%%
+
 It is important to understand that, apart from the founder nodes, the individual
@@
 will create GSPs for samples up to and including second-generation backcrosses (an example appears in Figure~\ref{fig:gsp2});
-%%%%%
-\begin{figure}
-\newcommand{\gspcaptwo}{\footnotesize
-An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
-admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $BC_1$ in sample node~s9; and two $BC_2$'s in sample node~s10.
-}
-\begin{center}
-\includegraphics[width=\columnwidth]{figures/AllThem.pdf}
-\end{center}
-\caption[\gspcaptwo]{\gspcaptwo}
-\label{fig:gsp2}
-\end{figure}
-%%%%
 however, \gscramble{} is designed
@@
 
-HEY! I SHOULD ALSO DO A SET OF SIMULATIONS IN WHICH THE 65 MARKERS ARE
-SIMULATED AS IF THEY ARE INDEPENDENTLY SEGREGATING---I THINK THAT MEANS
-ASSIGNING EACH TO ITS OWN CHROMOSOME.  BUT MAYBE WORRY ABOUT THAT
-LATER.
-
 We performed 500 replicate simulations with \gscramble{} using either 30 or 65 markers.  Each
@@
 from the contact regions of the 3 reference populations. The types of admixed individuals we considered were
-$F_1$, $F_2$, $F_{1}BC_1$ and $F_{1}BC_2$ representing all two-population combinations
+$F_1$, $F_2$, $F_{1}\BC_1$ and $F_{1}\BC_2$ representing all two-population combinations
 (Pop1 x Pop2, Pop1 x Pop3, Pop2 x Pop3, and reciprocally Pop2 x Pop1, Pop3 x Pop1, Pop3 x Pop2 as required
@@
 
-\subsection*{An introductory motivating example}
+\subsection*{Two Introductory Motivating Examples}
+
+\subsubsection*{RISPI when estimating admixture fractions}
 
@@
 
+\subsubsection*{Physical Linkage in Hybrid Individuals}
+
+Figure is completed and is in Fig.~\ref{fig:ad-fract-sims}.
+%%%%%%%%%%
+\begin{figure}
+\newcommand{\adfcap}{\footnotesize Admixture fractions of individuals simulated
+assuming no physical linkage (top four panels)
+and when physical linkage is simulated using \gscramble{}.  In the top 4 panels the number
+of simulated loci appears in the panel header.  Failing to simulate markers with physical linkage
+produces distributions of admixture fractions that are far more narrow than reality, giving the
+mistaken impression that admixture fractions of different hybrid categories do not overlap.  }
+\includegraphics[width=\columnwidth]{figures/link-and-nolink-sims-crop.pdf}
+\caption[\adfcap]{\adfcap}
+\label{fig:ad-fract-sims}
+\end{figure}
+%%%%%%%%%%
 \subsection*{Empirical Examples}
@@
 
-PRETTY MUCH DONE HERE EXCEPT I WANT TO SIMULATE THINGS
-UNLINKED.
+Figure is completed and is Fig.~\ref{fig:newhyb-sims-barplot}
 %%%%%%%%%%
 \begin{figure}
-\newcommand{\nhbpcap}{\footnotesize Gonna need a caption}
-\includegraphics[width=\columnwidth]{figures/newhybs-texed-crop.pdf}
+\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
+individuals of different trout hybrid categories.  Each pair of plots shows the results
+for individuals of the true hybrid category listed atop the pair.  (See Table~\protect\ref{tab:newhybcats} for meaning of the names.)  The top panel of each pair
+shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated an additional  35 markers, some of which are linked
+on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
+each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
+are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
+maximum posterior probability. }
+{\centering
+\includegraphics[width=0.98\columnwidth]{figures/newhybs-texed-crop.pdf}
+}
 \caption[\nhbpcap]{\nhbpcap}
@@
 Backcrossed individuals have higher ancestry associated with the backcrossed population with mean $Q$
-approximately $0.75$ ($BC_1$) and $0.87$ ($BC_2$) (Figure~\ref{fig:pig-sim-qs}A). Notably, the simulated
+approximately $0.75$ ($\BC_1$) and $0.87$ ($\BC_2$) (Figure~\ref{fig:pig-sim-qs}A). Notably, the simulated
 $F_1$ hybrids have the most variation in the distribution of the $Q$ values associated with each of the genetic clusters.
@@
 %%%
-\begin{figure*}
+\begin{figure}
 \newcommand{\pigsimsqscap}{\footnotesize (A) The density of the estimated Q values from ADMIXTURE for
@@
 grouped across all population-combinations for each hybrid class and colored by the $K$ genetic cluster
-(e.g., the Pop1 density plot for the $F_{1}BC_1$ plots the Pop1 genetic cluster for both
+(e.g., the Pop1 density plot for the $F_{1}\BC_1$ plots the Pop1 genetic cluster for both
 the Pop1-Pop2 and Pop1-Pop3 combinations). The density plots across the $K$
@@
 from both Pop1-Pop2 and Pop2-Pop1).}
-\includegraphics[width=0.96\textwidth]{figures/simulation-density-multipanel.pdf}
+\includegraphics[width=\columnwidth]{figures/simulation-density-multipanel.pdf}
 \caption[\pigsimsqscap]{\pigsimsqscap}
 \label{fig:pig-sim-qs}
-\end{figure*}
+\end{figure}
 %%%

### 2024-01-11 body-text-of-paper.tex
Rename F1BC to simply BC

@@
 from the contact regions of the 3 reference populations. The types of admixed individuals we considered were
-$F_1$, $F_2$, $F_{1}\BC_1$ and $F_{1}\BC_2$ representing all two-population combinations
+$F_1$, $F_2$, $\BC_1$ and $\BC_2$ representing all two-population combinations
 (Pop1 x Pop2, Pop1 x Pop3, Pop2 x Pop3, and reciprocally Pop2 x Pop1, Pop3 x Pop1, Pop3 x Pop2 as required
@@
 grouped across all population-combinations for each hybrid class and colored by the $K$ genetic cluster
-(e.g., the Pop1 density plot for the $F_{1}\BC_1$ plots the Pop1 genetic cluster for both
+(e.g., the Pop1 density plot for the $\BC_1$ plots the Pop1 genetic cluster for both
 the Pop1-Pop2 and Pop1-Pop3 combinations). The density plots across the $K$

### 2024-01-22 body-text-of-paper.tex
small tweaks
Add an annotation
More things to the examples methods and results.

@@
 \subsubsection*{RISPI when estimating admixture fractions}
+\label{sec:rispi-sim}
 
@@
 
-NOTE, WE ALSO DID THESE SIMULATIONS USING GSCRAMBLE AND WE SHOULD NOTE
-HOW THOSE CAME OUT.
+%NOTE, WE ALSO DID THESE SIMULATIONS USING GSCRAMBLE AND WE COULD NOTE
+%HOW THOSE CAME OUT.  
+% BUT MAYBE JUST HAVE THAT IN CASE A REFEREE ASKS ABOUT IT.
 
@@
 
+%%%%%%%
+\newcommand{\permcap}{\footnotesize An illustration of how \gscramble{} permutes data under
+different options.  {\bf (a)}~A picture
+representing an original data set of $N=12$ diploids sampled from three populations and typed at $L=15$ markers.
+Different colors represent different sampled individuals. Individuals 1--5 are shades of blue and green, indicating that they
+are from a single population.  Individuals 6--9 are from another population, (shades of red and orange), while 10--12
+are from a third population and are colored shades of violet. Thick vertical black lines separate the different
+populations.  Markers are shown in rows. There are 6 markers on the first chromosome, 5 on the second, and 4 on the
+third.
+Each chromosome is delimited by a thick, horizontal black line.
+The "a" and the "b" in the colored cells denote the first and second gene copies within an individual at each marker.
+It is worth noting that this representation shows a $2L \times N$ matrix,
+which is the transpose of the $N\times 2L$ matrix used to supply genetic marker data to \gscramble{}.
+{\bf (b)}~Examples of how the {\tt\footnotesize preserve\_individuals} and {\tt\footnotesize preserve\_haplotypes} options of the {\tt\footnotesize segments2markers()}
+function affect the permutation of
+the genetic data of the original data set.  Panel headers show the option values while the colors and "a" and "b" values
+in the colored cells show the original position of each gene copy in the original sample.  Note that permutation is always
+done within the populations.  When {\tt\footnotesize preserve\_individuals = TRUE},  then just the
+individuals in their entirety are permuted.  When {\tt\footnotesize preserve\_individuals = "BY\_CHROM"}, each pair
+of a single chromosome within an individual remains together when permuted.
+When {\tt\footnotesize preserve\_individuals = FALSE}, gene copies are freely permuted between individuals.
+The option {\tt\footnotesize preserve\_haplotypes = TRUE} causes each haplotype on a chromosome to be preserved
+while permuting. This option should only be used when the original data set is phased.
+}
+\begin{figure*}
+\includegraphics[width=\textwidth]{figures/permutations-plot.png}
+\caption[\permcap]{\permcap}
+\label{fig:perms}
+\end{figure*}
+%%%%%%%
+
 These permutations are implemented with the genotypes provided by the user.
@@
 
-%%%%%%%
-\newcommand{\permcap}{\footnotesize An illustration of how \gscramble{} permutes data under
-different options.  {\bf (a)}~A picture
-representing an original data set of $N=12$ diploids sampled from three populations and typed at $L=15$ markers.
-Different colors represent different sampled individuals. Individuals 1--5 are shades of blue and green, indicating that they
-are from a single population.  Individuals 6--9 are from another population, (shades of red and orange), while 10--12
-are from a third population and are colored shades of violet. Thick vertical black lines separate the different
-populations.  Markers are shown in rows. There are 6 markers on the first chromosome, 5 on the second, and 4 on the
-third.
-Each chromosome is delimited by a thick, horizontal black line.
-The "a" and the "b" in the colored cells denote the first and second gene copies within an individual at each marker.
-It is worth noting that this representation shows a $2L \times N$ matrix,
-which is the transpose of the $N\times 2L$ matrix used to supply genetic marker data to \gscramble{}.
-{\bf (b)}~Examples of how the {\tt\footnotesize preserve\_individuals} and {\tt\footnotesize preserve\_haplotypes} options of the {\tt\footnotesize segments2markers()}
-function affect the permutation of
-the genetic data of the original data set.  Panel headers show the option values while the colors and "a" and "b" values
-in the colored cells show the original position of each gene copy in the original sample.  Note that permutation is always
-done within the populations.  When {\tt\footnotesize preserve\_individuals = TRUE},  then just the
-individuals in their entirety are permuted.  When {\tt\footnotesize preserve\_individuals = "BY\_CHROM"}, each pair
-of a single chromosome within an individual remains together when permuted.
-When {\tt\footnotesize preserve\_individuals = FALSE}, gene copies are freely permuted between individuals.
-The option {\tt\footnotesize preserve\_haplotypes = TRUE} causes each haplotype on a chromosome to be preserved
-while permuting. This option should only be used when the original data set is phased.
-}
-\begin{figure*}
-\includegraphics[width=\textwidth]{figures/permutations-plot.png}
-\caption[\permcap]{\permcap}
-\label{fig:perms}
-\end{figure*}
-%%%%%%%
+% I THOUGHT ABOUT PUTTING THESE RESULTS IN, BUT I
+% THINK WE ARE FINE WITHOUT THEM.
+%\subsection{Simulation: \gscramble{} Eliminates RISPI}
+%
+%We present a brief simulation similar to those in
+%section ``RISPI when estimating admixture fractions,'' but conducted using
+%\gscramble{} rather than sampling with replacement.  We focused on two
+%scenarios that
 
-\subsection*{Empirical Examples}
+\subsection*{Examples}
 
-We illustrate the use of \gscramble{} in two examples.  The first involves
+We illustrate the use of \gscramble{} in three example cases. The first is
+a simple simulatino to show that, unlike sampling with replacement, the
+approach of \gscramble{} does not produce RISPI.
+The next two are examples using empirical data sets.
+The first of these involves
 simulating a small number ($<\,$70) of markers that are fixed, or nearly
@@
 pigs in Missouri.
+
+\subsubsection*{No RISPI when using \gscramble{}}
+
+For this example, we repeat the simulation described in \nameref{sec:rispi-sim};
+however, after simulating individuals in the two samples drawn from the same population,
+we create data sets of admixed individuals using \gscramble{}, rather than sampling
+gene copies with replacement.  As before, individuals were simulated with fractions of
+ancestry from sample $A$ in $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$, but these were simulated from pedigrees using gscramble.  We limited
+this simulation to one of the scenarios that showed extreme RISPI when sampling gene
+copies with replacement: $L=1,000$ markers with sizes of samples $A$ and $B$ being
+$N=50$. This simulation was repeated 240 times, and the results of each analyzed with
+ADMIXTURE to estimate the ancestry fractions of each simulated individual.
 
@@
 
-Figure is completed and is in Fig.~\ref{fig:ad-fract-sims}.
+When hybrids were simulated in a way that accounts for the physical linkage of
+markers, the variance in admixture fractions was considerably higher than when
+markers were simulated under the assumption that none of them are physically
+linked. (Fig.~\ref{fig:ad-fract-sims}).
 %%%%%%%%%%
@@
 %%%%%%%%%%
-\subsection*{Empirical Examples}
+If markers are assumed to be unlinked, then the variance around the $Q$ estimates of different 
+hybrid categories continually declines as more markers are added, from 100, to 100,000, which 
+makes it appear that different hybrid categories become increasingly well-resolved.  For example, 
+looking at the distributions for 100,000 unlinked markers one would think that it would be easy to 
+distinguish between the $BC_1$ through $BC_4$ categories on the basis of the admixture fractions.
+This is, however, not the case.  The bottom panel of Fig.\ref{fig:ad-fract-sims} shows the distribution
+of the true admixture fraction (calculated as the total length of genome from each of the two
+different species) of each individual when simulated with physical linkage using \gscramble{}.
+When physical linkage is not ignored, it is quite clear that there is considerable overlap in the
+admixture fractions between the $BC_1$ and $BC_2$ categories, and that $BC_2$, $BC_3$, and 
+$BC_4$ all have admixture fraction distributions that overlap greatly.  This shows that,
+in real life, when physical linkage cannot be ignored, it may be quite difficult to distinguish
+between different backcross categories on the basis of their admixture fractions.
+
+\subsection*{Examples}
+
+\subsubsection*{No RISPI when using \gscramble{}}
+
+As expected, there was no RISPI apparent in the ADMIXTURE-estimated
+ancestry proportions  when simulating individuals
+using \gscramble{} (Fig.~\ref{fig:norispi}).  On average, the estimated $Q$ values
+were around 0.5, which is expected when the samples are drawn from the same
+population.
+%%%
+\begin{figure}
+\newcommand{\gssimcap}{\footnotesize ADMIXTURE estimates of ancestry
+fractions, $Q$, for individuals simulated from samples from the same population,
+simulated using \gscramble{}.  There is a considerable variation in the estimates,
+but the central tendency of them all is around 0.5, as would be expected in this case
+within which there is not genetic basis for distinguishing individuals from the different
+samples.  Contrast this to the corresponding panel ($L=1,000, N=50$) in Fig.~\ref{fig:bias-sims}(b),
+in which the RISPI is readily apparent.}
+\includegraphics[width=\columnwidth]{figures/gscram_supervised_Qs.pdf}
+\caption[\gssimcap]{\gssimcap}
+\label{fig:norispi}
+\end{figure}
+%%%
 
@@
 
-Figure is completed and is Fig.~\ref{fig:newhyb-sims-barplot}
+The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
+probabilities of the different hybrid categories.
+
 %%%%%%%%%%
@@
 for individuals of the true hybrid category listed atop the pair.  (See Table~\protect\ref{tab:newhybcats} for meaning of the names.)  The top panel of each pair
-shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated an additional  35 markers, some of which are linked
+shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
+plus an additional  35 (65 total), some of which are linked
 on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
@@
 %%%%%%%%%%
+These are similar to ADMIXTURE barplots;
+however, rather than plotting the ancestry fraction, the different colors are showing the
+posterior probabilities of different hybrid categories.  Each individual is represented by a thin
+vertical line, and the individuals have been arranged from left to right and the colors have been
+arranged from top to bottom throughout different sections of the plot to make it easy to see
+the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
+by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
+category changes from one to another.  It is apparent over all individuals simulated to be
+either CCT, $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
+hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
+were used. By contrast, for the individuals simulated to be of the complex hybrid category
+$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
+than from 
+30 markers.  However, it is also apparent that using 65 markers, some of which are physically
+linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
+estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
+incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
+$BC_2$--CCT category,
+or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
+maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
+are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
+than does using the 30 unlinked markers. So, as expected, ignoring physical linkage the way
+{\sc NewHybrids} does can yield a poor estimate of the certainty around different inferences.
 

### 2024-02-03 body-text-of-paper.tex
started writing the discussion- hey! 3 paragraphs!

@@
 any inbreeding loop indicates that more than one copy of a gene in an ancestor may
-be created amongst its descendants, which is a type of sampling without
+be created amongst its descendants, which is a type of sampling with
 replacement.  Furthermore, in a GSP, the amount of genetic material taken from the
@@
 
+Here, we present \gscramble{}, a simulation approach for creating biologically informed individual 
+genotypes of different hybrid classes. Two integral features of \gscramble{} are: 
+1) alleles are sampled from populations without replacement and 
+2) alleles are segregated based on species-specific recombination rates. 
+Our results show that both 
+of these features are especially important for accurately reflecting empirical data from large genetic 
+marker sets (thousands of markers), 
+which are common from next generation sequencing methods. Sampling alleles without 
+replacement, and thereby preserving the genetic differences (or lack thereof) of populations is 
+essential for avoiding the issue of RISPI---an issue that is exacerbated as the number of 
+alleles sampled for an individual increases. Additionally, we show the importance of accounting for 
+the physical linkage of markers when large marker sets are used and that treating loci as 
+independent results in erroneously reduced variation in the estimation of individuals' admixture 
+fractions.
+
+\gscramble{} is conducive for numerous types of contemporary marker data sets. As mentioned above, 
+one key feature of \gscramble{} is the integration of recombination rates, thereby accounting for physical 
+linkage which is expected among larger contemporary data sets. Furthermore, \gscramble{} performs marker 
+segregation with computational efficiency---for the four different GSPs used with the invasive wild pig data set, 
+the mean run time of each simulation with 14,233 SNP loci was 2.3 seconds (range: 1.9-5.7 $s$). 
+While \gscramble{} is equally applicable to data sets with 
+microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites \citep{zhan2017megasat}, 
+it is not suited for genotype likelihood data (i.e. from low-coverage whole genome sequencing). 
+The \gscramble{} simulation approach involves the segregation of gametes and therefore requires 
+that genotypes are resolved.
+
+We envision \gscramble{} to be relevant for a diverse range of potential applications. In our 
+provided empirical examples with trout and wild pig data, we demonstrate the basics of 
+creating hybrid classes. Similarly, \gscramble{} has recently been used to demonstrate the 
+power for hybrid classification of a population genetics method used to differentiate wild pigs with 
+wildboar introgression from domestic pigs \citep{smyser2024probgen}. In \citet{smyser2024probgen}, 
+researchers made use of the segregation options in \gscramble{} in order to preserve founder individuals' 
+genotypes within populations (in contrast to \textit{a priori} shuffling alleles within populations) in order to 
+retain the subpopulation structure present within certain populations---thereby, better reflecting the 
+actual admixture variation expected from wildboar and domestic pig hybrids. We think \gscramble{} 
+may also be useful for assessing relatedness estimates using pedigrees of varying complexity. Here, it 
+is important to note that each time a gamete is segregated in a GSP, its complement is also 
+segregated---an example of antithetical variate sampling in Monte Carlo methods. Thus, in the 
+example of relatedness, fixed specified founders should be repeatedly "mated" with one offspring 
+sampled per iteration to avoid sampling of anticorrelated genotypes.
+
+\textbf{Eric's notes}: 
 Somewhere here I want to mention that the Anderson et al. (2008) approach for simulating

### 2024-02-05 abstract.tex
added draft abstract

%!TEX main.tex

While a best practice for evaluating the behavior of genetic clustering algorithms 
on empirical data is to conduct parallel analyses on simulated data, these types 
of simulation techniques often involve sampling genetic data with replacement. 
In this paper we demonstrate that sampling with replacement, especially with large 
marker sets, inflates the perceived statistical power to correctly assign individuals 
back to source populations---a phenomenon we refer to as resampling-induced, 
spurious power inflation (RISPI). To address this issue, we present \gscramble{}, a 
simulation approach in R for creating biologically informed individual genotypes from 
empirical data that: 1) samples alleles from populations \textit{without} replacement, 
2) segregates alleles based on species-specific recombination rates. \gscramble{} is 
a flexible simulation approach that allows users to create pedigrees of varying complexity 
in order to simulate admixed genotypes and it allows users to track haplotype blocks 
from the source populations through the pedigrees. We demonstrate the functionality of 
\gscramble{} with both simulated and empirical data sets and highlight additional uses of 
the package that users may find valuable.

### 2024-02-05 body-text-of-paper.tex
Add a and b to Fig 4, plus minor edits to discussion

@@
 is a bias resulting from the use of sampling with replacement to simulate new, admixed genotypes to
-test in ADMIXTURE.  In each figure, the different columns represent different numbers of markers
+test in ADMIXTURE.  ({\bf a}) Results from the supervised analysis in ADMIXTURE, where 
+"ref-2" and "ref-1" show the $Q$-values estimated by ADMIXTURE for the reference samples
+(even though they are treated by the program as being of known origin). ({\bf b}) Results for the
+unsupervised analysis in ADMIXTURE. In both {\bf a} and {\bf b}, the different columns represent different numbers of markers
 from 10 to 100,000, while different rows represent different original sample sizes, $N$, taken. Colors
@@
 marker sets (thousands of markers), 
-which are common from next generation sequencing methods. Sampling alleles without 
+which are common from  both custom SNP-chip genotyping and next generation sequencing methods. Sampling alleles without 
 replacement, and thereby preserving the genetic differences (or lack thereof) of populations is 
-essential for avoiding the issue of RISPI---an issue that is exacerbated as the number of 
+essential for avoiding RISPI---an issue that is exacerbated as the number of 
 alleles sampled for an individual increases. Additionally, we show the importance of accounting for 
 the physical linkage of markers when large marker sets are used and that treating loci as 
-independent results in erroneously reduced variation in the estimation of individuals' admixture 
-fractions.
+independent results in erroneously reduced variation in the admixture 
+fractions of simulated individuals.
 

### 2024-02-12 body-text-of-paper.tex
Eric edited through line 262 in body-text-of-paper.tex

@@
 samples, and
-inflates the perceived statistical power to correctly assign individuals back to source populations.
+inflates the perceived statistical power to correctly assign individuals (or segments of the genome) back to source populations.
 We refer to this phenomenon as resampling-induced, spurious power inflation (RISPI).
@@
 frequency of immigration can be gained by integrating across generations to identify both direct migrants and descendants of migrants
- (i.e., F1 hybrids or advanced hybrids reflecting the effect of backcrossing across multiple generations [BC1, \ldots,
-BCX]).
+ (i.e., F1 hybrids or advanced hybrids reflecting the effect of backcrossing across multiple generations $[BC_1, \ldots,
+BC_X])$.
 Evaluating the results of genetic clustering algorithms in the context of data that are
@@
 This was done recently to assess the power of a SNP panel for identifying descendants of escaped, farm-raised Atlantic salmon \citep{wringe2019development}; however the method used (described at https://eriqande.github.io/SalarHybPower/) was developed in an {\em ad hoc} fashion, was not easily extended to
-complex hybridization scenarios, and was not at all user-friendly.
+complex hybridization scenarios, and is not user-friendly.
 
@@
 replacement, thus maintaining genetic differences between the clusters commensurate to those in the empirical dataset, and eliminating the effect of RISPI.
-Further, by simulating individual genotypes based on user-defined pedigrees of arbitrary complexity, \gscramble{} allows for the simulation of admixed genotypes resulting from any conceivable category of hybrid, while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
+Further, by simulating individual genotypes based on user-defined pedigrees of arbitrary complexity, \gscramble{} allows for the simulation of admixed genotypes resulting from any conceivable category of (non-inbred) hybrid, while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
 By integrating species-specific recombination rates, \gscramble{} simulates biologically informed genotypes that mirror empirical data.
-We develop and illustrate \gscramble{} with use of single nucleotide polymorphic (SNP) genotypes, though it is equally applicable to data sets with hundreds or thousands of
+We develop and illustrate \gscramble{} through the use of single nucleotide polymorphism (SNP) genotypes, though it is equally applicable to data sets with hundreds or thousands of
 microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites
@@
 the effect of physical linkage on the distribution of admixture fractions of hybrid individuals,
-which vividly shows why it is not acceptable with large numbers of markers, to simulate
+which vividly shows why it is not acceptable, with large numbers of markers, to simulate
 the genotypes of hybrid individuals assuming no physical linkage.
@@
 for pigs, as estimated by from \citet{tortereau2012high}. In this case, we recorded the
-admixture fraction of each simulated individual as the fraction of all genome segments
-that originated from species~$a$.  This provides the distribution of the true admixture fraction
+admixture fraction of each simulated individual as the cumulative length of all genome segments
+that originated from species~$a$, divided by the total genome length.  This provides the distribution of the true admixture fraction
 when physical linkage of markers is not ignored.
@@
 Our scheme, described here, extends the traditional cross-validation approach
-by simulating the descent of chromosomal linkage blocks through a user-specified
+by simulating the descent of chromosomal segments through a user-specified
 pedigree to create admixed individuals.  By imposing certain constraints
@@
 samples at the sample nodes should be equal to the amount of material in the founders,
-to ensure that all available data is being used to estimate power for the estimation of
-admixture fractions.
+to ensure that all available data are being used in the simulations to predict the power for
+estimating admixture fractions.
 

### 2024-02-26 body-text-of-paper.tex
Eric edited through the end of the methods.

@@
 uses terminology that is defined in the figure caption.
-
 It is important to understand that, apart from the founder nodes, the individual
@@
 to handle user-specified GSPs of arbitrary complexity.  When creating a unique GSP,
-the following requirements are necessary for the GSP to be valid
-and the simulation to execute successfully.
+there are several requirements to ensure that the GSP is valid
+and the simulation can execute successfully.
 Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
@@
 \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\
+ S_x/2  &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfounds \\
  S_x/2 + \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfoundse.  \\
@@
 \end{equation}
+In words, the separate lines above specify that 1) the number of samples
+must equal the number of founders; 2) exactly two gametes must be
+segregated from each founder; 3) the number of gametes incoming on each of the
+two edges coming into a non-founder (from its parents) must be equal; 
+4) the number of gametes going downward, out of a non-founder with no sample
+daughters, must equal the number of gametes coming into that non-founder;
+5)  the total number of gametes entering a non-founder with only sample daughters
+must be half the number of sample daughters; and 6) the number of gametes
+entering non-founders with both sample daughters and non-founder daughters must
+equal half the number of its sample daughters plus the sum of the number of gametes segregated
+along each edge to all of its non-founder daughters.
+
 Before simulating from a GSP, \gscramble{} checks to ensure the GSP
@@
 We illustrate the use of \gscramble{} in three example cases. The first is
-a simple simulatino to show that, unlike sampling with replacement, the
+a simple simulation to show that, unlike sampling with replacement, the
 approach of \gscramble{} does not produce RISPI.
@@
 
-For this example, we repeat the simulation described in \nameref{sec:rispi-sim};
+For this example, we repeat the simulation described in {\em \nameref{sec:rispi-sim}};
 however, after simulating individuals in the two samples drawn from the same population,
@@
 copies with replacement: $L=1,000$ markers with sizes of samples $A$ and $B$ being
-$N=50$. This simulation was repeated 240 times, and the results of each analyzed with
+$N=50$. This simulation was repeated 240 times, and the results of each replicate were analyzed with
 ADMIXTURE to estimate the ancestry fractions of each simulated individual.
@@
 the linkage map described in \citep{pearse2019sex} that was prepared by linkage mapping 
-of $\approx$57,000 markers described in \citet{palti2015development} to Omyk\_v1.0. We mapped the $\approx$57,000 markers to Omyk\_v1.0, and then resolved inconsistencies to produce a map
+of $\approx$57,000 markers described in \citet{palti2015development} to Omyk\_v1.0. We mapped these $\approx$57,000 markers to Omyk\_v1.0, and then resolved inconsistencies to produce a map
 of 32,923 markers throughout the genome and rendered it to plink format for use with \gscramble{}.
@@
 1 $F_2$,
-6 cutthroat backcrosses ($BC1$s) and 
-6 second-generation cutthroat backcrosses ($BC2$s),
+6 cutthroat backcrosses ($BC_1$s) and 
+6 second-generation cutthroat backcrosses ($BC_2$s),
 and 1 fish of an unknown hybrid category that may have been the
-product of a $BC2$ crossed with a $BC1$.   In our simulations, we sought to simulate
+product of a $BC_2$ crossed with a $BC_1$.   In our simulations, we sought to simulate
 a similar collection of hybrid individuals, so, for each replicate, we used
-\gscramble{} to simulate 16 $F_1$s, 4 $F_2$s, 6 $BC1$s, 4 $BC2$s, 4 $BC3$s,  4 individuals
-that were the products of a $BC1$ mating with a $BC2$, and finally, 4 individuals that were simulated pure cutthroat. Individuals not used as founders
+\gscramble{} to simulate 16 $F_1$s, 4 $F_2$s, 6 $BC_1$s, 4 $BC_2$s, 4 $BC_3$s,  4 individuals
+that were the products of a $BC_1$ mating with a $BC_2$, and finally, 4 individuals that were simulated pure cutthroat. Individuals not used as founders
 in \gscramble{} were retained as pure cutthroat and pure steelhead.  The goal of the simulation
@@
 $F_1$, $F_2$, $\BC_1$ and $\BC_2$ representing all two-population combinations
-(Pop1 x Pop2, Pop1 x Pop3, Pop2 x Pop3, and reciprocally Pop2 x Pop1, Pop3 x Pop1, Pop3 x Pop2 as required
-to simulate backcrossed classes).
+(Pop1 x Pop2, Pop1 x Pop3, Pop2 x Pop3, and reciprocally Pop2 x Pop1, Pop3 x Pop1, Pop3 x Pop2).
 We also simulated individuals from each of the founder populations. The first step of this workflow involved

### 2024-03-04 body-text-of-paper.tex
These are edits that Rachael sent me
Text and fig caps in Results to end of mykiss stuff.
Just have the pigs left to do in the results.

@@
 hybrid category using \gscramble{}'s {\footnotesize\tt segregate()} function, as described
-below, with number of chromosomes and sex-averaged recombination rates
+below, with 18 chromosomes (2.35 gigabases of physical genome length)  and
+sex-averaged recombination rates
 for pigs, as estimated by from \citet{tortereau2012high}. In this case, we recorded the
@@
 \gscramble{} accepts genetic data as a matrix with $N$ rows (one for each individual) and
-$2L$ columns (two for each locus). If the data have been phased, then the phased alleles from the
+$2L$ columns (two alleles for each locus). If the data have been phased, then the phased alleles from the
 first haplotype should occupy the
@@
 is a bias resulting from the use of sampling with replacement to simulate new, admixed genotypes to
-test in ADMIXTURE.  ({\bf a}) Results from the supervised analysis in ADMIXTURE, where 
+test in ADMIXTURE (e.g., these simulated samples suffer from RISPI).  ({\bf a}) Results from the supervised analysis in ADMIXTURE, where 
 "ref-2" and "ref-1" show the $Q$-values estimated by ADMIXTURE for the reference samples
@@
 markers were simulated under the assumption that none of them are physically
-linked. (Fig.~\ref{fig:ad-fract-sims}).
+linked (Fig.~\ref{fig:ad-fract-sims}).
 %%%%%%%%%%
@@
 assuming no physical linkage (top four panels)
-and when physical linkage is simulated using \gscramble{}.  In the top 4 panels the number
+and when physical linkage is simulated using \gscramble{} assuming a genetic map with 18 chromosomes spanning 2.35 gigabases with recombination fractions like those found in pigs
+\citep{tortereau2012high}.  In the top 4 panels the number
 of simulated loci appears in the panel header.  Failing to simulate markers with physical linkage
@@
 distinguish between the $BC_1$ through $BC_4$ categories on the basis of the admixture fractions.
-This is, however, not the case.  The bottom panel of Fig.\ref{fig:ad-fract-sims} shows the distribution
+This is, however, not the case.  The bottom panel of Fig.~\ref{fig:ad-fract-sims} shows the distribution
 of the true admixture fraction (calculated as the total length of genome from each of the two
@@
 admixture fractions between the $BC_1$ and $BC_2$ categories, and that $BC_2$, $BC_3$, and 
-$BC_4$ all have admixture fraction distributions that overlap greatly.  This shows that,
+$BC_4$ all have admixture-fraction distributions that overlap greatly.  This shows that,
 in real life, when physical linkage cannot be ignored, it may be quite difficult to distinguish
@@
 fractions, $Q$, for individuals simulated from samples from the same population,
-simulated using \gscramble{}.  There is a considerable variation in the estimates,
+simulated using \gscramble{}.  There is considerable variation in the estimates,
 but the central tendency of them all is around 0.5, as would be expected in this case
-within which there is not genetic basis for distinguishing individuals from the different
-samples.  Contrast this to the corresponding panel ($L=1,000, N=50$) in Fig.~\ref{fig:bias-sims}(b),
+within which the different
+samples are drawn from the same population.  Contrast this to the corresponding panel ($L=1,000, N=50$) in Fig.~\ref{fig:bias-sims}(b),
 in which the RISPI is readily apparent.}
@@
 probabilities of the different hybrid categories. 
-
 %%%%%%%%%%
@@
 individuals of different trout hybrid categories.  Each pair of plots shows the results
-for individuals of the true hybrid category listed atop the pair.  (See Table~\protect\ref{tab:newhybcats} for meaning of the names.)  The top panel of each pair
+for individuals of the true hybrid category listed atop the pair  (see
+ Table~\protect\ref{tab:newhybcats} for the meanings of the names).  The top panel of each pair
 shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
@@
 %%%%%%%%%%
-These are similar to ADMIXTURE barplots;
+These are similar to STRUCTURE or ADMIXTURE barplots;
 however, rather than plotting the ancestry fraction, the different colors are showing the
@@
 than does using the 30 unlinked markers. So, as expected, ignoring physical linkage the way
-{\sc NewHybrids} does can yield a poor estimate of the certainty around different inferences.
+{\sc NewHybrids} does can yield a poor estimate of the uncertainty around different inferences.
 

### 2024-03-18 body-text-of-paper.tex
Small tweaks

@@
 3 populations (40 individuals each informed by previous analyses)
-and 40 individuals from the contact regions of these populations.
+and 40 individuals from contact regions between these populations.
 We used ADMIXTURE version 1.3.0 \citep{alexander2009fast} to characterize the genetic structure
 of the 160 samples (for $K = {1...6}$). Using the 120 reference individuals identified to represent population cores,
-we sought to classify the admixture patterns of the 40 empirical samples
+we sought to contextualize the admixture patterns of the 40 samples
 from the contact regions using simulated pedigrees from \gscramble{}. To address this objective we simulated
@@
 individuals. Although we could have simulated all 4 types of admixed individuals from a single GSP (Figure~\ref{fig:gsp2}),
-different numbers of simulated individuals are produced for each hybrid class. Therefore, for the sake of the
-computational efficiency from simulating the exact number of desired individuals, we chose to simulate from
-a single GSP for each of the hybrid classes by setting the corresponding hybrid parameter in {\footnotesize\tt create\_GSP()} to {\footnotesize\tt TRUE}.
-We used the default settings of the {\footnotesize\tt plink\_map2rec\_rates()} to specify the recombination rates.
+different numbers of simulated individuals are produced for each hybrid class. Therefore, for ease of simulating the exact number of desired individuals, we chose to simulate from
+a single GSP for each of the hybrid classes by setting the corresponding hybrid category (i.e.
+{\tt F1}, {\tt F2}, etc.) parameter in {\footnotesize\tt create\_GSP()} to {\footnotesize\tt TRUE} with all others set to {\footnotesize\tt FALSE}.
+We used the default settings of the {\footnotesize\tt plink\_map2rec\_rates()}  function to specify the recombination rates.
 The 120 simulated genotypes were written to PLINK .ped and .map files using {\footnotesize\tt gscramble2plink()}.
@@
 Using the 160 empirical samples, ADMIXTURE results supported the identification of
-$K=3$ genetic clusters grouped individuals from each of the core populations into 3 clusters
+$K=3$ genetic clusters that grouped individuals from each of the core populations into 3 clusters
 while the 40 individuals from the contact regions were highly admixed. The
@@
 Backcrossed individuals have higher ancestry associated with the backcrossed population with mean $Q$
-approximately $0.75$ ($\BC_1$) and $0.87$ ($\BC_2$) (Figure~\ref{fig:pig-sim-qs}A). Notably, the simulated
-$F_1$ hybrids have the most variation in the distribution of the $Q$ values associated with each of the genetic clusters.
-When the estimated Q values for the $F_1$ hybrids are separated by the population combinations,
+approximately $0.75$ ($\BC_1$) and $0.87$ ($\BC_2$) (Figure~\ref{fig:pig-sim-qs}A).
+Interestingly, the distribution of the $Q$ values for $F_1$'s varies more across the different genetic clusters than it does for the other
+hybrid categories.
+When the estimated $Q$ values for the $F_1$ hybrids are separated by the population combinations,
 $F_1$ hybrids for Pop1-Pop2 have the least variation around $q=0.5$ (standard deviation
@@
 in the empirical data, specifically, that genetic differentiation is highest between Pop1 and Pop2 and that
-Pop14 is the most admixed population with ancestry contributions from both Pop1 and Pop2.
+Pop3 is the most admixed population with ancestry contributions from both Pop1 and Pop2.
+The results emphasize difficulties that may arise in estimating admixture fractions between
+groups that have experienced historical gene flow.
 

### 2024-03-25 body-text-of-paper.tex
added the supplement

@@
 Using the 160 empirical samples, ADMIXTURE results supported the identification of
-$K=3$ genetic clusters that grouped individuals from each of the core populations into 3 clusters
-while the 40 individuals from the contact regions were highly admixed. The
+$K=3$ genetic clusters (Figure S1) that grouped individuals from each of the core populations into 3 clusters
+while the 40 individuals from the contact regions were highly admixed (Figure S2). The
 $F_\mathrm{ST}$ divergence from ADMIXTURE was:

### 2024-04-15 body-text-of-paper.tex
Small edits Tim had for us in Monday meeting!

@@
 
-A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to conduct
-parallel analyses on data that have been simulated---so the truth is known---so as to mirror the empirical data set
-as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
+A best practice for evaluating the behavior of genetic clustering algorithms on empirical data is to conduct
+parallel analyses on simulated data that mirror the empirical data set
+as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine} while providing
+the benefit of known parameters (e.g.~$F_\mathrm{ST}$, admixture fractions, etc.) that can
+be compared to estimates.
 Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the
@@
 BC_X])$.
-Evaluating the results of genetic clustering algorithms in the context of data that are
+Evaluating the results and resolution of genetic clustering algorithms in the context of data that are
 expected to arrive by such a process of hybridization requires simulation of
@@
 from the alleles at that locus in sample $A$ ($B$).
-We then added these $9n$ simulated individuals to the $N$ original individuals  from $A$ and $N$ from $B$
-and analyzed all the genotypes using ADMIXTURE with $K=2$. On each data set we performed a supervised analysis in which the original
+We then analyzed all the genotypes ($9n + N + N$) using ADMIXTURE with $K=2$. On each data set we performed a supervised analysis in which the original
 $N$ samples from each
@@
 Our approach for simulating admixed individuals without creating RISPI is motivated by the traditional cross-validation
-approach for evaluating the power of mixture proportion
-estimation in genetic stock identification (e.g., the cross-validation approach
-described in \citealt{moran2019bayesian}).
+approach for assessing the power of mixed stock analysis (e.g., the ``Monte Carlo cross-validation'' approach
+described in \citealt{moran2019bayesian}, p.~555).
 In the cross-validation approach, individuals are removed from the reference
-samples and placed in a mixture sample of "unknowns", whose origin
+samples and placed in a mixture sample of ``unknowns,'' whose origin
 is inferred using the remaining individuals in the reference samples.  A key
@@
 in the original sample is being used---either within the remaining individuals in the
-reference samples or within the individuals in the mixture.
+reference samples or within the individuals set aside as a test set in the
+simulated mixture.
 

### 2024-05-07 body-text-of-paper.tex
made some edits

@@
 all their ancestry from a single subpopulation.
-Similar statistical artifacts occur when simulating data to
-assess how accurately admixture fractions can be estimated.
 
@@
 Evaluating the results and resolution of genetic clustering algorithms in the context of data that are
-expected to arrive by such a process of hybridization requires simulation of
+expected to reflect a history of hybridization requires simulation of
 explicit hybrids.  Such simulations are the purpose of the software {\sc hybridlab} 
@@
 
-To construct this simple example, we used the R programming language \citep{Rcoreteam} to
+To construct this simple illustration of sampling-with-replacement induced RISPI, 
+we used the R programming language \citep{Rcoreteam} to
 simulate allele frequencies for $L$ biallelic

### 2024-05-07 body-text-of-paper.tex
new changes

@@
 We further expected that ADMIXTURE would return more accurate estimates of the simulated
-$q_A$ values when $N$ was smaller and the number of loci was larger.
-We summarized the results by plotting boxplots of the ADMIXTURE-inferred $q_A$ values,
+$q_A$ values when $N$ was smaller and the number of loci was larger - 
+attributes that magnify the effect of RISPI.
+We summarized the results with boxplots of the ADMIXTURE-inferred $q_A$ values,
  for the 480 newly simulated individuals (and for the original samples, $A$ and $B$, in the
@@
 must equal the number of founders; 2) exactly two gametes must be
-segregated from each founder; 3) the number of gametes incoming on each of the
-two edges coming into a non-founder (from its parents) must be equal; 
+segregated from each founder; 3) the number of gametes to a non-founder along the 
+edges from its parents must be equal; 
 4) the number of gametes going downward, out of a non-founder with no sample
@@
 different numbers of simulated individuals are produced for each hybrid class. Therefore, for ease of simulating the exact number of desired individuals, we chose to simulate from
-a single GSP for each of the hybrid classes by setting the corresponding hybrid category (i.e.
-{\tt F1}, {\tt F2}, etc.) parameter in {\footnotesize\tt create\_GSP()} to {\footnotesize\tt TRUE} with all others set to {\footnotesize\tt FALSE}.
+a single GSP for each of the hybrid classes by setting the corresponding hybrid category (i.e., {\tt F1}, {\tt F2}, etc.) parameter in {\footnotesize\tt create\_GSP()} to {\footnotesize\tt TRUE} with all others set to {\footnotesize\tt FALSE}.
 We used the default settings of the {\footnotesize\tt plink\_map2rec\_rates()}  function to specify the recombination rates.
@@
 using na\"{i}ve simulations improperly indicates that admixture fractions of individuals can be
-accurately estimated, even when there are no genetic differences, whatsoever between the populations that
+accurately estimated, even when there are no genetic differences between the populations from which
 samples $A$ and $B$ are drawn.
@@
 
-As expected, there was no RISPI apparent in the ADMIXTURE-estimated
+As expected, there was no observed RISPI in the ADMIXTURE-estimated
 ancestry proportions  when simulating individuals
@@
 category changes from one to another.  It is apparent over all individuals simulated to be
-either CCT, $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
+either CCT (cut-throat trout), $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
 hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
@@
 are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
-than does using the 30 unlinked markers. So, as expected, ignoring physical linkage the way
-{\sc NewHybrids} does can yield a poor estimate of the uncertainty around different inferences.
+than does using the 30 unlinked markers. So, as expected, ignoring physical linkage (e.g., as 
+implemented in {\sc NewHybrids}) 
+can yield a poor estimate of the uncertainty around different inferences.
 
@@
 essential for avoiding RISPI---an issue that is exacerbated as the number of 
-alleles sampled for an individual increases. Additionally, we show the importance of accounting for 
-the physical linkage of markers when large marker sets are used and that treating loci as 
-independent results in erroneously reduced variation in the admixture 
+alleles sampled for an individual increases. Additionally, we demonstrate the importance of accounting for 
+the physical linkage of markers when large marker sets are used and that treating physically linked loci as 
+independent erroneously reduces the variation in the admixture 
 fractions of simulated individuals.
 
-\gscramble{} is conducive for numerous types of contemporary marker data sets. As mentioned above, 
+\gscramble{} is conducive for analysis of numerous types of marker sets available from contemporary data generation 
+techniques. As mentioned above, 
 one key feature of \gscramble{} is the integration of recombination rates, thereby accounting for physical 
@@
 it is not suited for genotype likelihood data (i.e. from low-coverage whole genome sequencing). 
-The \gscramble{} simulation approach involves the segregation of gametes and therefore requires 
-that genotypes are resolved.
+Specifically, simulations conducted within \gscramble{} involve the segregation of gametes within the 
+specified pedigrees and, therefore, requires genotypes to be resolved.
 
@@
 provided empirical examples with trout and wild pig data, we demonstrate the basics of 
-creating hybrid classes. Similarly, \gscramble{} has recently been used to demonstrate the 
-power for hybrid classification of a population genetics method used to differentiate wild pigs with 
-wildboar introgression from domestic pigs \citep{smyser2024probgen}. In \citet{smyser2024probgen}, 
-researchers made use of the segregation options in \gscramble{} in order to preserve founder individuals' 
-genotypes within populations (in contrast to \textit{a priori} shuffling alleles within populations) in order to 
-retain the subpopulation structure present within certain populations---thereby, better reflecting the 
-actual admixture variation expected from wildboar and domestic pig hybrids. We think \gscramble{} 
+creating hybrid classes. In much the same manner, \gscramble{} has been used to evaluate the statistical power 
+of hybrid classification used to differentiate wild boar hybrids from domestic pigs 
+\citep{smyser2024probgen}. Specifically \citet{smyser2024probgen} used the segregation options in \gscramble{} 
+ to preserve founder individuals' genotypes within populations in contrast to \textit{a priori} shuffling alleles 
+within populations. This retains the subpopulation structure present within certain populations---thereby, better reflecting the 
+actual admixture variation expected from wild boar and domestic pig hybrids. We think \gscramble{} 
 may also be useful for assessing relatedness estimates using pedigrees of varying complexity. Here, it 

### 2024-05-07 body-text-of-paper.tex
added "lack-of-recombination-map" discussion piece

@@
 one key feature of \gscramble{} is the integration of recombination rates, thereby accounting for physical 
-linkage which is expected among larger contemporary data sets. Furthermore, \gscramble{} performs marker 
+linkage which is expected among high density markers sets. Furthermore, \gscramble{} performs marker 
 segregation with computational efficiency---for the four different GSPs used with the invasive wild pig data set, 
@@
 \citep{smyser2024probgen}. Specifically \citet{smyser2024probgen} used the segregation options in \gscramble{} 
- to preserve founder individuals' genotypes within populations in contrast to \textit{a priori} shuffling alleles 
+ to preserve founder individuals' genotypes in contrast to \textit{a priori} shuffling alleles 
 within populations. This retains the subpopulation structure present within certain populations---thereby, better reflecting the 
@@
 sampled per iteration to avoid sampling of anticorrelated genotypes.
+
+Here, we have demonstrated the consequences of failing to consider physical linkage among loci 
+given the availability of a recombination map. However, in the absence of a recombination map, 
+researchers with a basic understanding of the genome size and the number of chromosomes in 
+the study organism could still benefit from using gscramble to simulate assumed linkage. To do 
+so, users of gscramble would need to randomly distribution markers across, and within, the 
+known number of chromosomes to produce a conceptual marker map (*.map within the PLINK 
+file structure), and assume basic rates of recombination (e.g., 1 cM per 1 Mb). The resulting 
+simulations from gscramble would produce admixture fractions that capture the expected 
+variation from known linkage patterns, in contrast to the reduced variation produced from 
+assuming independence (Figure 5). Characterizing uncertainty would provide caution for 
+researchers to not over interpret conclusions drawn in the absence of a recombination map.
 
@@
 
-Also, we need to discuss the fact that, because each time a gamete is segregated in the
-pedigree, its complement is also segregated, this is an example of antithetical variate sampling
-in Monte Carlo.
-
 \section*{Acknowledgements}

### 2024-05-13 abstract.tex
Eric final edits

@@
-%!TEX main.tex
+%!TEX root = main.tex
 
@@
 marker sets, inflates the perceived statistical power to correctly assign individuals
+(or the alleles that they carry)
 back to source populations---a phenomenon we refer to as resampling-induced, 
@@
 empirical data that: 1) samples alleles from populations \textit{without} replacement, 
-2) segregates alleles based on species-specific recombination rates. \gscramble{} is 
-a flexible simulation approach that allows users to create pedigrees of varying complexity 
-in order to simulate admixed genotypes and it allows users to track haplotype blocks 
-from the source populations through the pedigrees. We demonstrate the functionality of 
+2) segregates alleles based on species-specific recombination rates. This framework
+makes it possible to simulate admixed individuals in a way that respects the physical
+linkage between markers on the same chromosome and which does not suffer
+from RISPI.  This is achieved in \gscramble{} by allowing 
+users to specify pedigrees of varying complexity 
+in order to simulate admixed genotypes, segregating and tracking haplotype blocks 
+from different source populations through those pedigrees, and then sampling---using
+a variety of permutation schemes---alleles from empirical data into those haplotype blocks.
+We demonstrate the functionality of 
 \gscramble{} with both simulated and empirical data sets and highlight additional uses of 

### 2024-05-13 body-text-of-paper.tex
Eric final edits

--- 2024-05-07 body-text-of-paper.tex
@@
 represented in the sample to be among the characterized $K$ subpopulations.
-Furthermore, the characterization of admixed individuals may reflect a limitation of the statistical power of the genetic data,
-as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure.
+Furthermore, the characterization of admixed individuals may reflect a limitation of the statistical power of the genetic data (as a function of both the marker set and sample size) to resolve the true underlying patterns of genetic structure.
  The challenge for researchers then becomes correctly identifying the ecological and/or statistical processes by which observed complex
@@
 
-In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without
+In response to these needs, we developed \gscramble{}---a simulation approach that samples alleles from respective populations without
 replacement, thus maintaining genetic differences between the clusters commensurate to those in the empirical dataset, and eliminating the effect of RISPI.
@@
 population. In other words, we expected to observe RISPI.
-We further expected that ADMIXTURE would return more accurate estimates of the simulated
-$q_A$ values when $N$ was smaller and the number of loci was larger - 
-attributes that magnify the effect of RISPI.
+We further expected that ADMIXTURE would return estimates of $q_A$ that were closer
+to the simulated values when $N$ was smaller and the number of loci was larger---attributes that magnify the effect of RISPI.
 We summarized the results with boxplots of the ADMIXTURE-inferred $q_A$ values,
@@
 These two gametes are then segregated at random down the edge (or edges) from the founder node.
-Upon each edge, $i$ from founder node $f$, $G^-_{f,i}$ gametes are segregated.
+Upon each edge, $i$ from founder node $f$,  a number, $G^-_{f,i}$, of gametes are segregated.
 \item Each non-founder node, $x$, will have exactly two edges coming
@@
 simulating a small number ($<\,$70) of markers that are fixed, or nearly
-fixed, for alternate alleles between two species of trout: rainbow trout ({\em 
+fixed, for alternate alleles between two species of trout: steelhead trout ({\em 
 Oncorhynchus mykiss}) and cutthroat trout ({\em O. clarkii}). The goal of the 
@@
 gene copies with replacement.  As before, individuals were simulated with fractions of
-ancestry from sample $A$ in $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$, but these were simulated from pedigrees using gscramble.  We limited
+ancestry from sample $A$ in $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$, but these were simulated from pedigrees using \gscramble{}.
+We limited
 this simulation to one of the scenarios that showed extreme RISPI when sampling gene
@@
 
-\subsubsection*{Rainbow/cutthroat trout hybrids in California}
+\subsubsection*{Steelhead/cutthroat trout hybrids in California}
 
 We used the archived data from \citet{rizza2023limited}, drawing upon their 65 markers
-that are fixed or are nearly fixed for alternate alleles between rainbow trout and cutthroat trout.
+that are fixed or are nearly fixed for alternate alleles between steelhead trout and cutthroat trout.
 They designated 634 fish as pure cutthroat trout and 213 as pure steelhead.  We used the
@@
 1 $F_2$,
-6 cutthroat backcrosses ($BC_1$s) and 
-6 second-generation cutthroat backcrosses ($BC_2$s),
+6 cutthroat backcrosses ($BC_1$--CCTs) and 
+6 second-generation cutthroat backcrosses ($BC_2$--CCTs),
 and 1 fish of an unknown hybrid category that may have been the
@@
 
-\subsubsection*{Rainbow/cutthroat trout hybrids in California}
+\subsubsection*{Steehlead/cutthroat trout hybrids in California}
 
@@
 category changes from one to another.  It is apparent over all individuals simulated to be
-either CCT (cut-throat trout), $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
+either CCT (cutthroat trout), $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
 hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
@@
 The simulated hybrid classes produce
-the expected range of admixture with $F_1$ and $F_2$ individuals having mean $Q=0.5$, with $F_2$ individuals having
+the expected range of admixture with $F_1$ and $F_2$ individuals having mean $Q=0.5$, and $F_2$ individuals having
 more variation (Figure~\ref{fig:pig-sim-qs}A).
@@
 approximately $0.75$ ($\BC_1$) and $0.87$ ($\BC_2$) (Figure~\ref{fig:pig-sim-qs}A).
-Interestingly, the distribution of the $Q$ values for $F_1$'s varies more across the different genetic clusters than it does for the other
+Interestingly, the distribution of the $Q$ values for $F_1$ hybrids varies more across the different genetic clusters than it does for the other
 hybrid categories.
@@
 
-Here, we present \gscramble{}, a simulation approach for creating biologically informed individual 
+Here, we present \gscramble{}, an R package implementing a simulation
+approach for creating biologically informed individual 
 genotypes of different hybrid classes. Two integral features of \gscramble{} are: 
@@
 essential for avoiding RISPI---an issue that is exacerbated as the number of 
-alleles sampled for an individual increases. Additionally, we demonstrate the importance of accounting for 
-the physical linkage of markers when large marker sets are used and that treating physically linked loci as 
+genetic markers increases. Additionally, we demonstrate the importance of accounting for 
+the physical linkage of markers when large marker sets are used, and that treating physically linked loci as 
 independent erroneously reduces the variation in the admixture 
@@
 
-\gscramble{} is conducive for analysis of numerous types of marker sets available from contemporary data generation 
+\gscramble{} is suitable for analysis of numerous types of marker sets available from contemporary data generation 
 techniques. As mentioned above, 
 one key feature of \gscramble{} is the integration of recombination rates, thereby accounting for physical 
-linkage which is expected among larger contemporary data sets. Furthermore, \gscramble{} performs marker 
+linkage which is expected among high-density markers sets. Furthermore, \gscramble{} performs marker 
 segregation with computational efficiency---for the four different GSPs used with the invasive wild pig data set, 
@@
 microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites \citep{zhan2017megasat}, 
-it is not suited for genotype likelihood data (i.e. from low-coverage whole genome sequencing). 
+it is not built to use genotype likelihoods (i.e., from low-coverage whole genome sequencing). 
 Specifically, simulations conducted within \gscramble{} involve the segregation of gametes within the 
-specified pedigrees and, therefore, requires genotypes to be resolved.
+specified pedigrees and, therefore, require called genotypes (though missing data is acceptable).
 
@@
 is important to note that each time a gamete is segregated in a GSP, its complement is also 
-segregated---an example of antithetical variate sampling in Monte Carlo methods. Thus, in the 
-example of relatedness, fixed specified founders should be repeatedly "mated" with one offspring 
+segregated---an example of antithetical variate sampling in Monte Carlo methods \citep{hammersley}.
+Thus, in the  example of relatedness, fixed specified founders should be repeatedly "mated" with one offspring 
 sampled per iteration to avoid sampling of anticorrelated genotypes.
 
-\textbf{Eric's notes}: 
-Somewhere here I want to mention that the Anderson et al. (2008) approach for simulating
-mixtures of genotypes to assess how well one can estimate mixture proportions---the fraction
-of individuals from each of a set of baseline or reference populations---allows the user to
-simulate, if desired, more genotypes than occur in the reference samples.  This is because
-a trick is done in the inference stage to remove the correlation between the simulated individual
-and the reference sample, relative to the true population, by removing the simulated individual's
-genotypes from its own reference population when calculating the likelihood that its genotype
-originated from each of the reference populations.  To do the same when doing simulations for
-assessing power using ADMIXTURE or STRUCTURE would require hacking the underlying code
-of those programs, which is not desirable.  So, for these cases we have used the sampling
-without replacement approach.
-
-Also, we need to discuss the fact that, because each time a gamete is segregated in the
-pedigree, its complement is also segregated, this is an example of antithetical variate sampling
-in Monte Carlo.
+Here, we have demonstrated the consequences of failing to consider physical linkage among loci 
+given the availability of a recombination map. However, in the absence of a recombination map, 
+researchers with a basic understanding of the genome size and the number of chromosomes in 
+the study organism could still benefit from using \gscramble{} to simulate genetic data under
+an assumed, approximate linkage map. To do 
+so, users of \gscramble{} could randomly distribute markers across, and within, the 
+known number of chromosomes to produce a conceptual marker map (*.map within the PLINK 
+file structure), and assume basic rates of recombination (e.g., 1 cM per 1 Mb). The resulting 
+simulations from \gscramble{} would produce admixture fractions that capture the expected 
+variation from known linkage patterns, in contrast to the reduced variation produced from 
+assuming independence (Figure 5). Characterizing uncertainty in this way would provide caution for 
+researchers to not overinterpret conclusions drawn in the absence of a recombination map.
 
@@
 within southeastern Missouri.
-We thank Yniv Palti for providing flanking sequences and the linkage map for 57K markers for making a recombination map for the cutthroat/rainbow trout example.
+We thank Yniv Palti for providing flanking sequences and the linkage map for 57K markers for making a recombination map for the steelhead/cutthroat trout example.
 Writing of this paper was facilitated by the Mobile High Altitude Venue for Ecological Analysis,
-Genetics and Statistics, through which ECA, MGD, and TJS were scientists-in-residence on location in
+Genetics and Statistics, through which ECA, MGD, and TJS were scientists-in-residence on location near
 Cheyenne, WY for 3 days in October 2023. This is contribution number mHAVEAGAS-003.

### 2024-05-13 data-accessibility.tex
Clean up to make it a GitHub repo
Eric final edits

@@
 
-{\sc microhaplot} is available as an R package with these associated links:
 \begin{itemize}
-\item Stable version available on CRAN (when available): \url{https://CRAN.R-project.org/package=gscramble}
+\item Stable version of \gscramble{} available on CRAN: \url{https://CRAN.R-project.org/package=gscramble}
 \item Development version and entire revision history on GitHub: \url{https://github.com/eriqande/gscramble}
 \item Online version of all package documentation: \url{https://eriqande.github.io/gscramble}
-\item Data repository with full data from pig, Chinook, and bird (maybe?) data sets \url{https://dryad.something.or.other}
+\item Data and code repository with simulation analyses and steelhead-cutthroat analyses: \url{https://github.com/eriqande/gscramble-paper-sims-and-analyses} (To be transferred to Dryad or Zenodo upon acceptance).
+\item Data and code repository for pig analyses: {NEED THE URL FOR THIS} (To be transferred to Dryad or Zenodo upon acceptance)
 \end{itemize}

### 2024-05-13 body-text-of-paper.tex
Disable endfloat for one-column version
To send to Tony

@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/F2.pdf}
+\includegraphics[width=0.48\textwidth]{figures/F2.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figures/AllThem.pdf}
+\includegraphics[width=0.48\textwidth]{figures/AllThem.pdf}
 \end{center}
@@
 {\small
-\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}} lrrr}
+\begin{tabular*}{0.48\textwidth}{@{\extracolsep{\fill}} lrrr}
 \hline\hline
@@
 mistaken impression that admixture fractions of different hybrid categories do not overlap.  }
-\includegraphics[width=\columnwidth]{figures/link-and-nolink-sims-crop.pdf}
+\includegraphics[width=0.48\textwidth]{figures/link-and-nolink-sims-crop.pdf}
 \caption[\adfcap]{\adfcap}
@@
 in which the RISPI is readily apparent.}
-\includegraphics[width=\columnwidth]{figures/gscram_supervised_Qs.pdf}
+\includegraphics[width=0.48\textwidth]{figures/gscram_supervised_Qs.pdf}
 \caption[\gssimcap]{\gssimcap}
@@
 {\centering
-\includegraphics[width=0.98\columnwidth]{figures/newhybs-texed-crop.pdf}
+\includegraphics[width=0.48\textwidth]{figures/newhybs-texed-crop.pdf}
 }
@@
 from both Pop1-Pop2 and Pop2-Pop1).}
-\includegraphics[width=\columnwidth]{figures/simulation-density-multipanel.pdf}
+\includegraphics[width=0.48\textwidth]{figures/simulation-density-multipanel.pdf}
 \caption[\pigsimsqscap]{\pigsimsqscap}

### 2024-05-22 body-text-of-paper.tex
matt revised some pig pieces re: Toni's comments

@@
 investigates the accuracy for
-identifying individuals admixed between closely related groups of feral
+identifying admixed individuals of different hybrid classes between 
+closely related groups of invasive wild 
 pigs in Missouri.
@@
 federal and state agencies.
-DNA was extracted from hair or tissue (collected from animals at the time of euthanasia) with
+DNA was extracted from hair or tissue (collected from animals at the time of control) with
 commercially available magnetic bead recovery kits (MagMax DNA, Thermo Fisher Scientific).
@@
 
-The simulated hybrid classes produce
-the expected range of admixture with $F_1$ and $F_2$ individuals having mean $Q=0.5$, and $F_2$ individuals having
+The simulated hybrid classes of wild pig populations demonstrate how \gscramble{} can be 
+used to accurately produce 
+the expected variation in the admixture fractions of different hybrid classes. 
+$F_1$ and $F_2$ individuals had the expected range of admixture with 
+mean $Q=0.5$, and $F_2$ individuals having
 more variation (Figure~\ref{fig:pig-sim-qs}A).
@@
 \newcommand{\pigsimsqscap}{\footnotesize (A) The density of the estimated Q values from ADMIXTURE for
-the simulated runs are
+the simulated runs of wild pigs are
 grouped across all population-combinations for each hybrid class and colored by the $K$ genetic cluster
@@
 and the National Science Foundation (award 2201538). 
+The findings and conclusions in this publication are those of the author(s) and should not be construed to represent any official USDA or U.S.
+Government determination or policy.
+This research was supported in part by the U.S.
+Department of Agriculture, Animal Plant Inspection Service, Wildlife Services, National Wildlife Research Center.

### 2024-05-22 body-text-of-paper.tex
Eric completed the edits from Tony's comments

@@
 and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.  Genetic movement has many implications,
-such as influencing evolutionary processes, maintaining genetic variation through gene flow, driving disease dynamics and affecting patterns
+such as influencing evolutionary processes, maintaining genetic variation through gene flow, driving disease dynamics, and affecting patterns
 of invasive species expansion \citep{huestis2019windborne, estoup2010reconstructing}.
@@
 Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the
-observed allele frequencies (see, for example, \citealt{nielsen2006hybridlab,kinziger2008hybridization}).
-However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed
+observed allele frequencies (see, for example, \citealt{nielsen2006hybridlab,kinziger2008hybridization}). In this
+approach, allele frequencies are estimated for each population (or cluster) from the samples, and then new
+genotypes are simulated at each locus by drawing, for each simulated diploid individual, two gene copies from a 
+multinomial distribution with two trials
+and cell probabilities given by the estimated allele frequencies.  Such a simulation is akin to assuming that the
+population is of infinite size with allele frequencies given by those observed in the sample; however the allele
+frequencies in the sample differ from those in the actual population because of the variation from random sampling.
+As a consequence, this sort of sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed
 samples, and
-inflates the perceived statistical power to correctly assign individuals (or segments of the genome) back to source populations.
+inflates the perceived statistical power to correctly assign individuals (or segments of the genome) back to source 
+populations.  In effect, simulations that sample with replacement from the samples are assessing how well one could 
+assign individuals to populations
+that have the same allele frequencies as the samples, rather than the same allele frequencies as the populations
+{\em from which the samples were drawn}, which is what is needed.
 We refer to this phenomenon as resampling-induced, spurious power inflation (RISPI).
@@
 $q_A$, the admixture fraction for population $A$, in  $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$.
-We simulated these genotypes according to the "admixture model" used in {\em structure} \citep{pritchard2000inference}.  Briefly, at each
+We simulated these genotypes according to the ``admixture model'' used in {\em structure} \citep{pritchard2000inference}.  Briefly, at each
 locus, independently, the origin
@@
 hybrid category using \gscramble{}'s {\footnotesize\tt segregate()} function, as described
-below, with 18 chromosomes (2.35 gigabases of physical genome length)  and
-sex-averaged recombination rates
-for pigs, as estimated by from \citet{tortereau2012high}. In this case, we recorded the
+below, taking as an example, a species with 18 chromosomes across 2.35 gigabases of total physical genome length
+and sex-averaged recombination rates the same as for pigs, as estimated by from \citet{tortereau2012high}. (We chose
+a pig-like genome here, as wild pigs form one of our example empirical data sets for illustration, below.)  In this case, we recorded the
 admixture fraction of each simulated individual as the cumulative length of all genome segments
-that originated from species~$a$, divided by the total genome length.  This provides the distribution of the true admixture fraction
+that originated from species~$a$, divided by the total genome length.
+This provides the distribution of the true admixture fraction
 when physical linkage of markers is not ignored.
@@
 Accession: GCF\_002163495.1), so we used
-the linkage map described in \citep{pearse2019sex} that was prepared by linkage mapping 
-of $\approx$57,000 markers described in \citet{palti2015development} to Omyk\_v1.0. We mapped these $\approx$57,000 markers to Omyk\_v1.0, and then resolved inconsistencies to produce a map
+the linkage map described in \citep{pearse2019sex} that was prepared 
+with $\approx$57,000 markers described in \citet{palti2015development} mapped to Omyk\_v1.0. We mapped these $\approx$57,000 markers to Omyk\_v1.0, and then resolved inconsistencies to produce a map
 of 32,923 markers throughout the genome and rendered it to plink format for use with \gscramble{}.
@@
 segregated---an example of antithetical variate sampling in Monte Carlo methods \citep{hammersley}.
-Thus, in the  example of relatedness, fixed specified founders should be repeatedly "mated" with one offspring 
+Thus, in the  example of relatedness, fixed specified founders should be repeatedly ``mated'' with one offspring 
 sampled per iteration to avoid sampling of anticorrelated genotypes.

### 2024-07-11 body-text-of-paper.tex
Formmating stuff with endfloat
Fixing quotation marks and urls...

@@
 \end{center}
-\caption[\gspcapone]{\gspcapone}
+\caption[\gspcapone]{\sometimes{\gspcapone}}
 \label{fig:gsp1}
@@
 \end{center}
-\caption[\gspcaptwo]{\gspcaptwo}
+\caption[\gspcaptwo]{\sometimes{\gspcaptwo}}
 \label{fig:gsp2}
@@
 Each chromosome is delimited by a thick, horizontal black line.
-The "a" and the "b" in the colored cells denote the first and second gene copies within an individual at each marker.
+The ``a'' and the ``b'' in the colored cells denote the first and second gene copies within an individual at each marker.
 It is worth noting that this representation shows a $2L \times N$ matrix,
@@
 function affect the permutation of
-the genetic data of the original data set.  Panel headers show the option values while the colors and "a" and "b" values
+the genetic data of the original data set.  Panel headers show the option values while the colors and ``a'' and ``b'' values
 in the colored cells show the original position of each gene copy in the original sample.  Note that permutation is always
@@
 \begin{figure*}
+\begin{center}
 \includegraphics[width=\textwidth]{figures/permutations-plot.png}
-\caption[\permcap]{\permcap}
+\end{center}
+\caption[\permcap]{\sometimes{\permcap}}
 \label{fig:perms}
@@
 \label{tab:newhybcats}
+\begin{center}
 {\small
@@
 \vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
+\end{center}
 \end{table}
@@
 test in ADMIXTURE (e.g., these simulated samples suffer from RISPI).  ({\bf a}) Results from the supervised analysis in ADMIXTURE, where 
-"ref-2" and "ref-1" show the $Q$-values estimated by ADMIXTURE for the reference samples
+``ref-2'' and ``ref-1'' show the $Q$-values estimated by ADMIXTURE for the reference samples
 (even though they are treated by the program as being of known origin). ({\bf b}) Results for the
@@
 replicate.}
+\begin{center}
 \includegraphics[width=0.96\textwidth]{figures/bias-sims-unsup-and-sup.pdf}
-\caption[\biassimscap]{\biassimscap}
+\end{center}
+\caption[\biassimscap]{\sometimes{\biassimscap}}
 \label{fig:bias-sims}
@@
 mistaken impression that admixture fractions of different hybrid categories do not overlap.  }
+\begin{center}
 \includegraphics[width=0.48\textwidth]{figures/link-and-nolink-sims-crop.pdf}
-\caption[\adfcap]{\adfcap}
+\end{center}
+\caption[\adfcap]{\sometimes{\adfcap}}
 \label{fig:ad-fract-sims}
@@
 in which the RISPI is readily apparent.}
+\begin{center}
 \includegraphics[width=0.48\textwidth]{figures/gscram_supervised_Qs.pdf}
-\caption[\gssimcap]{\gssimcap}
+\end{center} 
+\caption[\gssimcap]{\sometimes{\gssimcap}}
 \label{fig:norispi}
@@
 maximum posterior probability. }
-{\centering
+\begin{center}
 \includegraphics[width=0.48\textwidth]{figures/newhybs-texed-crop.pdf}
-}
-\caption[\nhbpcap]{\nhbpcap}
+\end{center}
+\caption[\nhbpcap]{\sometimes{\nhbpcap}}
 \label{fig:newhyb-sims-barplot}
@@
 Pop1-Pop2 compared to Pop1-Pop3 and Pop2-Pop3. Note that these population combinations group the
-reciprocal population combination in each plot (e.g., "Pop1\_Pop2" has simulations
+reciprocal population combination in each plot (e.g., ``Pop1\_Pop2'' has simulations
 from both Pop1-Pop2 and Pop2-Pop1).}
+\begin{center}
 \includegraphics[width=0.48\textwidth]{figures/simulation-density-multipanel.pdf}
-\caption[\pigsimsqscap]{\pigsimsqscap}
+\end{center}
+\caption[\pigsimsqscap]{\sometimes{\pigsimsqscap}}
 \label{fig:pig-sim-qs}

### 2024-07-11 cover-letter/gscramble-cover-letter.tex
First draft of cover letter

\documentclass[12pt]{letter}
\usepackage{graphicx}
%\usepackage{fontspec}
%\setmainfont{Times New Roman}
%\setmainfont{TeX Gyre Termes}

\textwidth = 6.5 in
\textheight = 9.0 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = 0.0 in
\headsep = 0.0 in
\parskip = 0.2in
\parindent = 0.0in

\date{}

\address{\mbox{}\vspace*{-1.1in}
\mbox{}\hfill
\includegraphics*{/Users/eriq/Documents/work/nonprj/NMFS_FORMS/NMFS_letterhead_mcallister.pdf}\\}

\signature{Eric C. Anderson \\ \mbox{} \\ eric.anderson@noaa.gov \\ ph: 831-227-4239}

\begin{document}

\begin{letter}{ \today

Editorial Board \\
Molecular Ecology Resources
}

\pagestyle{empty}
\opening{To the Editor:}

I am delighted to submit the accompanying manuscript,
``{\sc gscramble}: Simulation of admixed individuals without reuse of genetic material,''
to {\em Molecular Ecology Resources}. While it has been understood since the exposition
in Anderson, Waples, \& Kalinowski (2008; {\em Can.\ J.\ Fish.\ Aquat.\ Sci.})\ that care must be
taken when simulating data sets to assess the power of a marker panel for population
assignment and/or genetic stock identification, it has been less widely appreciated that the
same holds true for assessing how much power is available for estimating admixture fractions
and identifying hybrids between closely related groups.  With the advent of data sets
with thousands, or even hundreds of thousands, of genetic markers, this problem
becomes severe.  At the same time, with so many markers, it is critical to account
for physical linkage when assessing how accurately different hybrid categories
can be resolved.

In our paper, we start with several simple simulations to show how dramatic the bias
of power estimation can be with modern data sets, a phenomenon that we refer to
as resampling-induced spurious power inflation, or RISPI.  We then describe a novel,
pedigree-based approach to sampling alleles to create hybrid and admixed individuals
to properly estimate the power for admixture analysis and hybrid identification that avoids
RISPI, and takes full account of physical linkage. We are confident that this approach provides
state-of-the-art methodology for rigorously comparing the outcome of
genetic clustering approaches, such as STRUCTURE, and ADMIXTURE, to results on
simulated data sets that are commensurate with the original data.

In the remainder of the paper, we provide two examples of how the approach can
be used to assess the power for hybrid identification and admixture proportion estimation
in case studies involving Pacific trouts and feral pigs.

The approach we have developed is implemented in the R package `gscramble' which
is currently available for download from the Comprehensive R Archive
Network (CRAN), and includes full documenation and vignettes.
Members of our team have used `gscramble' already in one published project and
the package is currently in use in several more.   We predict that the package will be
widely used in the future, and as a consequence, the paper will be highly cited.

Thank you for the opportunity to submit our work to {\em Molecular Ecology Resources}.

\closing{Sincerely,}
\setlength{\topmargin}{0in}
\end{letter}

 \end{document}

### 2024-07-11 cover-letter/gscramble-cover-letter.tex
Update gscramble-cover-letter.tex
typo

@@
 is currently available for download from the Comprehensive R Archive
-Network (CRAN), and includes full documenation and vignettes.
+Network (CRAN), and includes full documentation and vignettes.
 Members of our team have used `gscramble' already in one published project and

### 2024-07-12 data-accessibility.tex
pig git repo added
Fixing quotation marks and urls...
Add author contributions

@@
-\section*{Data Accessibility}
+\section*{Author Contributions}
+
+The gscramble simulation approach grew from discussions between TJS and ECA\@.
+ECA, RMG, MGD, and TJS wrote the R package, `gscramble.'   Simulations of bias due to RISPI and
+of variance in admixture fractions due to physical linkage were conducted by ECA.
+The empirical analyses using `gscramble' were conducted
+by MGD and RMG with data from TJS and by ECA with previously published data.  All authors contributed to writing and
+editing the manuscript.
+
+\section*{Data Availability Statement}
 
@@
 \item Data and code repository with simulation analyses and steelhead-cutthroat analyses: \url{https://github.com/eriqande/gscramble-paper-sims-and-analyses} (To be transferred to Dryad or Zenodo upon acceptance).
-\item Data and code repository for pig analyses: {NEED THE URL FOR THIS} (To be transferred to Dryad or Zenodo upon acceptance)
+\item Data and code repository for pig analyses: \url{https://github.com/mgdesaix/gscramble-paper-pigs} (To be transferred to Dryad or Zenodo upon acceptance)
 \end{itemize}

### 2024-07-12 body-text-of-paper.tex
Update author contributions

@@
 segregated---an example of antithetical variate sampling in Monte Carlo methods \citep{hammersley}.
-Thus, in the  example of relatedness, fixed specified founders should be repeatedly ``mated'' with one offspring 
+Thus, in the  example of relatedness, fixed specified founders should be repeatedly ``mated,'' with one offspring 
 sampled per iteration to avoid sampling of anticorrelated genotypes.

### 2024-07-15 body-text-of-paper.tex
small edit

@@
 We thank Yniv Palti for providing flanking sequences and the linkage map for 57K markers for making a recombination map for the steelhead/cutthroat trout example.
-Writing of this paper was facilitated by the Mobile High Altitude Venue for Ecological Analysis,
+Writing of this paper was facilitated by the mobile High Altitude Venue for Ecological Analysis,
 Genetics and Statistics, through which ECA, MGD, and TJS were scientists-in-residence on location near

### 2024-07-15 author-title-etc.tex
MGD added TJS's changes
added draft abstract
Tiny tweaks before submitting

@@
 % All caps tends to look better
-\newcommand{\myAuthors}{Eric C. Anderson$^{*,\dagger,\ddag,\S}$, Rachel M. Giglio$^{\#}$, Matthew G. DeSaix$^{\ddag,\#}$ Timothy J. Smyser$^{\#}$}
-\author{Eric C. Anderson$^{*,\dagger,\ddag,\S}$, Rachel M. Giglio$^{\#}$, Matthew G. DeSaix$^{\ddag,\#}$ Timothy J. Smyser$^{\#}$}
+\newcommand{\myAuthors}{Eric C. Anderson$^{*,\dagger,\ddag,\S}$, Rachael M. Giglio$^{\#}$, Matthew G. DeSaix$^{\ddag,\#}$ Timothy J. Smyser$^{\#}$}
+\author{Eric C. Anderson$^{*,\dagger,\ddag,\S}$, Rachael M. Giglio$^{\#}$, Matthew G. DeSaix$^{\ddag,\#}$ Timothy J. Smyser$^{\#}$}
 
@@
 
-\renewcommand{\KeyWords}{genetic stock identification, mixture deconvolution, permutation testing, power analysis, blah, blah}
+\renewcommand{\KeyWords}{genetic stock identification, mixture deconvolution, permutation testing, power analysis, R package}
 
@@
 % here you can put your very own copyright notice
-\newcommand{\myCopyright}{\copyright US Federal Government work in the public domain in the USA}
+\newcommand{\myCopyright}{}
 

### 2024-07-15 CompileForSubmission.sh
Compile script for making submission easier

@@ -1,0 +1,27 @@
+
+
+# Ths just puts things into a folder called TMP-SUBMIT that I can then
+# test to make sure it compiles.
+
+# When I see that it does compile, I re-run this, then I zip the TMP-SUBMIT
+# folder and submit to Wiley as a Latex Source Folder
+
+rm -rf TMP-SUBMIT
+mkdir TMP-SUBMIT
+
+cp main.{tex,bbl} TMP-SUBMIT/
+cp eca_molecolres.sty TMP-SUBMIT/
+
+cp abstract.tex \
+  author-title-etc.tex \
+  body-text-of-paper.tex \
+  data-accessibility.tex \
+  newcomms.tex \
+  TMP-SUBMIT/
+
+cp -r figures TMP-SUBMIT/
+cp -r inputs TMP-SUBMIT/
+cp -r images TMP-SUBMIT/
+
+rm TMP-SUBMIT/figures/permutations-plot.pdf
+

### 2024-07-17 CompileForSubmission.sh
Update submission script to do one-col version

@@ -6,10 +6,14 @@
 # When I see that it does compile, I re-run this, then I zip the TMP-SUBMIT
 # folder and submit to Wiley as a Latex Source Folder
 
+# which version?
+BASE="main-one-col"
+#BASE="main"
+
 rm -rf TMP-SUBMIT
 mkdir TMP-SUBMIT
 
-cp main.{tex,bbl} TMP-SUBMIT/
+cp $BASE.{tex,bbl} TMP-SUBMIT/
 cp eca_molecolres.sty TMP-SUBMIT/
 
 cp abstract.tex \

### 2024-07-18 body-text-of-paper.tex
Add Sam Rizza to acknowledgements

@@
 within southeastern Missouri.
-We thank Yniv Palti for providing flanking sequences and the linkage map for 57K markers for making a recombination map for the steelhead/cutthroat trout example.
-Writing of this paper was facilitated by the mobile High Altitude Venue for Ecological Analysis,
+We thank Yniv Palti for providing flanking sequences and the linkage map for 57K markers for making a recombination map for the steelhead/cutthroat trout example, and Sam Rizza for providing additional insights
+into those same data. Writing of this paper was facilitated by the mobile High Altitude Venue for Ecological Analysis,
 Genetics and Statistics, through which ECA, MGD, and TJS were scientists-in-residence on location near

### 2024-12-03 cover-letter/responses-to-editor-and-referees.Rmd
Revisions for pub/accept

---
title: "Responses to Editor and Referees"
output: pdf_document

---

## Editor Comments to Author:

Two reviewers have provided comments on this manuscript and both are favorably impressed.  They have raised a few minor issues that should be addressed in a revision.  In particular, it would be good to better address the limitations when there is inbreeding present.  I look forward to reading a revised version.

_Thank you.  We address the issue of inbreeding in the revised version.  See below.
Our replies and comments are in italic text throughout._

Subject Editor, Molecular Ecology Resources

## Reviewer Comments to Author:

## Reviewer: 1

### General comments:

In this work, Anderson and colleagues present a new R package; gscramble which can be used to simulate genotypes across hybrid classes. The package differs from other commonly used programs such as ADMIXTURE by, sampling alleles without replacement and using recombination rates to account for linkage among large genetic datasets.

Both these considerations are important and represent improvements to current methods. I do have some questions about how the package is implemented and what some of the requirements mean for real-world applications detailed below. Once clarified, I believe this work is suitable for publication in Molecular Ecology Resources.

### Detailed comments:

L90-91: Are the first two figures supposed to appear one after the other in the Introduction? They aren’t referred to in the text until line 202 in the methods.

_Thank you for catching that.  They should not appear there.  That was a byproduct of using the same
\LaTeX{} input for both a typeset-2-column version and our submission version.  We have moved the
figure callouts to their appropriate locations in the paper._

L120-123: Is it not possible to simply filter the loci set based on linkage disequilibrium? Do you have evidence that this will not alleviate the problem?

_No, it is not possible. Genomic material is inherited in chunks that are typically orders of magnitude longer than
the typical range over which LD decays.  Thus, pairs of markers that are
on the same chromosome, are much more likely to be co-inherited than markers that
are not on the same chromosome, even if the two markers are far enough apart on the
same chromosome that they are not in LD. We have noted this in the discussion for any other readers that may
wonder whether LD pruning will solve the problem of physical linkage._

L196-198: How would the requirement of having no inbreeding loops in the GSP affect the applicability of these simulations to real-life scenarios? Within most wild populations, especially threatened species, inbreeding is a regular occurrence.

_This is a reasonable question and we address it now in the Discussion. In short, gscramble works fine
for inbred populations.  Partly this is because if the founders are related (as they would be in a tiny population)
the offspring produced by the GSP will by inbred despite the absence of inbreeding loops in the GSP._

L202-204: Figure captions should not contain required information for the text only. The terminology should be described in the text at the first mention.

_We pulled that long caption out of the figure and include it in the text. That improves things.
Thank you!_

L227-229: Will recombination rates be something the user will have to input?

_Yes, the user has to put those in, though we already note in the discussion that even if an
accurate recombination map is not known, it is still possible to use one simply with
average rates across each chromosome._

L363-365: My understanding was that gscramble accounted for the bias associated with linked loci without the need to filter, so why would an empirical dataset need to be filtered for linked loci? What benefit then does gscramble present over other methods in this regard?

_As noted above, LD and physical linkage are quite different.  Where we now address that in the discussion, we also
note that for applications of the data after it has been run through gscramble, like using ADMIXTURE, it can be
advantageous to have LD-pruned data._

Figure 7: How did the results of gscramble differ from the original paper? How do you know what is the truth? Or do the results simply differ?

_The point of this simulation was to see whether the original authors were justified in limiting their data set to just the 30
markers on separate chromosome arms. We now emphasize in the results section that this choice is justified by the simulation results._

L482: Is there a function in gscramble for users to estimate the “specific-specific recombination rates” from their empirical data?

_No, there is not._

## Reviewer: 2

### Comments to the Author

The manuscript "gscramble: Simulation of admixed individuals without reuse of genetic material" presents a new R package to simulate genetically admixed individuals. The main advancement presented is that this simulation avoids the issue of resampling and the spurious power inflation that can result from this.

The method allows users to upload their own recombination maps and specify complex hybridization scenarios. To test their method the authors present simulations with trout and pig data which showcase the strength of their method.

Overall the paper is well written. It takes complex concepts and presents them in plain english. I commend the authors for the clear organization and concise writing.

The authors do present some of the limitations of their method, but I think they could do a little bit more of this. This might be outside the scope of the paper but I wonder how this might apply in the field of human genetics.

_This is an interesting question. We do feel that gscramble could be applicable in some
contexts within human genetics.  The ready availability of phased dense data in humans
would be appropriate for use with the `preserve_haplotypes = TRUE` option. So, gscramble
could be used to assess how well admixture from different poulations could be detected.
Given the emphasis on molecular ecology in the outlet, however, we agree with the
reviewer that including discussion of human genetics might be outside of the scope of the
paper._

The authors mention "genetic clustering algorithms" multiple times in the text, it seems to be a major motivator for their software, however they only use ADMIXTURE. I wonder if the authors could make these results more robust by presenting results using another clustering algorithm (could even just be K-means) to show that their results hold up using more than other clustering algorithm.

_I think it is fairly well understood from the explanations in Anderson, Waples, and Kalinowski (2008) that
RISPI would also affect something like K-means clustering.  We considered adding more simulations to the paper, but felt that
the paper was already quite long as it is.  We focus on ADMIXTURE in the paper because it provides
an explicit estimate of the admixture fraction, and one of the main points of the paper is that
the variance of admixture fractions within recent hybrids is strongly affected by physical linkage._

The graphics of the paper could be improved with larger text sizes. This is an aesthetic suggestion, but I wonder if the 4 by 4 grid plots could be reduced to a 3 by 3 grid to increase the overall size of each plot while preserving the visualization of the trends.

_Yes, being able to make those 3 x 3 grids could make them more readable; however we do
feel it is important to have those four magnitudes of marker numbers available, since they
do attempt to span the range of the sorts of data sets that might be encountered in molecular
ecology today: 100 = Fluidigm chips and amplicon sequencing; 1000 = marker capture;
10,000 = RAD seq and small SNP chips; 100,000+ RAD seq, large SNP chips and whole genome
sequencing. Ultimately, we chose to keep them 4 x 4 to preserve that range, and symmetry
in the vertical dimension._

## Production Checklist

To hasten the appearance of your manuscript online, I urge you to complete the following prior to returning your revised manuscript:

1. Ensure data is archived in a publicly-accessible repository. Please see the Data Storage and Documentation section in the author guidelines for details of what data should be archived, appropriate public archives for specific data types and suitable file formats. Exemptions to the data archiving policy may be granted for cases involving sensitive information such as human subject data or the location of endangered species.

_Done. Data and code are archived and publicly accessible._

2. Complete Data Accessibility and Benefit-Sharing section in your manuscript. For correct formatting of this section, please see our author guidelines. Please include DOIs or accession numbers in your Data Accessibility Statement. Data Accessibility statements will be checked post acceptance but prior to publication for completeness and that information such as DOIs or accession numbers are included.

_Done. Data and code are archived and publicly accessible, and that is described in the statement._

Molecular Ecology Resources strongly encourages authors to include a Benefit-Sharing Statement, which should disclose non-monetary benefits generated by your research in the context of the Nagoya Protocol (https://www.cbd.int/abs/). Please review the benefits associated with this research and describe them within the Data Accessibility and Benefit-Sharing section in your manuscript. For more information, please see our editorial https://onlinelibrary.wiley.com/doi/10.1111/mec.15702.

_We have added a Benefit-Sharing Statement_

3. Consolidate Supplemental Information into as few files as possible.

    * Use the attached Molecular Ecology Resources Supplemental Information Template if possible (not required)
    * If the file is quite long, create a Table of Contents for the first page
    * Make sure the title of the paper and the names of all authors appear on the Supplemental document.
    * Your Supplemental Information will not be copy-edited.  Do not leave in track-changes and other editing marks.  The document will be posted "as is.”
    * Save as PDF if possible.

_The Supplemental Information has been saved as a PDF_

4. Provide your manuscript in an editable format (not PDF!)

5. Provide tables in an editable format (not images)

6. Make sure all table and figure captions are present and editable

_All will be submitted in LaTeX format, and we will ensure proper rendering_

The title and abstract of your article are very important to increasing its visibility. We encourage you to make your article more discoverable online with the following tips:

https://authorservices.wiley.com/author-resources/Journal-Authors/Prepare/writing-for-seo.html

_Thank you for those SEO tips!_

To submit your revised manuscript:

1. Log in to your Scholar One account and click on the 'manuscripts with decisions' link
2. Click 'create a revision' and follow the on-screen instructions. (Do not upload your revised manuscripts as a new submission.)
3. Please include a point-by-point response to the reviewer comments, including a detailed list of changes made and a rebuttal to any comments with which you disagree, being as specific as possible in your responses.
4. Please provide a version of your manuscript with the changes tracked or highlighted.
5. Your original files will be available to you when you upload your revised manuscript. Please delete any outdated files before completing the submission.

In order to schedule publication efficiently, we impose a deadline of 30 days for the return of revised manuscripts. Manuscripts returned after the deadline may be regarded as new submissions, with a new receipt date, and may be subject to the full peer review process.

Wiley Editing Services Available to All Authors
Should you be interested, Wiley Editing Services offers expert help with manuscript, language, and format editing, along with other article preparation services. You can learn more about this service option at www.wileyauthors.com/eeo/preparation. You can also check out Wiley’s collection of free article preparation resources for general guidance about writing and preparing your manuscript at www.wileyauthors.com/eeo/prepresources.

This journal offers a number of license options for published papers; information about this is available here: https://authorservices.wiley.com/author-resources/Journal-Authors/licensing/index.html. The submitting author has confirmed that all co-authors have the necessary rights to grant in the submission, including in light of each co-author’s funder policies. If any author’s funder has a policy that restricts which kinds of license they can sign, for example if the funder is a member of Coalition S, please make sure the submitting author is aware.

Thank you for submitting your research to Molecular Ecology Resources.

Best Wishes,

Benjamin Sibbett PhD
Executive Editor, Molecular Ecology Resources
E-mail: molecol@wiley.com

### 2024-12-03 body-text-of-paper.tex
Revisions for pub/accept

@@
 \citep{nielsen2006hybridlab}.
-
-%%%%%%%%%%%
-\begin{figure}
-\newcommand{\gspcapone}{\footnotesize
-A genomic simulation pedigree (GSP) for simulating $F_2$ individuals.
-Squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
-individuals 1--4.  The  triangles above each of those founders denotes, in color
-and text, the population of origin of each of the two genomes within each founder.
-Thus, individuals~1 and~2 are founders purely from population~$A$, and~3 and~4
-are purely from population~$B$.  The pink hexagon labeled s7 is the sample node.
-It represents samples that are taken from non-founder node~7, which, from the structure
-of the pedigree, clearly represents an $F_2$ hybrid between populations $A$ and $B$.  The red numbers
-on edges above and leading into each non-founder node indicate the number of
-gametes that get segregated from the parent node (at the top of the edge) to the
-daughter node (at the bottom of the edge).
-This emphasizes that,
-in dropping genes through a GSP, in contrast
-to a traditional pedigree, each parent
-node may segregate multiple gametes to a daughter node.  See the text for
-further explanation.
-Each non-founder node $x$ will have exactly two edges coming into it from above
-(from its two parents), and we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$.
-For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
-A founder individual may have, at most, 2 edges coming downward out of it, while a non-founder individual may have more or less than two edges coming downward
-out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
-non-founder daughter nodes, we denote the number of gametes segregated
-through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example, at
-non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
-Finally, the purple number adjacent to the edge leading into the
-sample node indicates how many sample individuals are drawn from individual
-node 7. Any non-founder individual node can have a maximum of one edge leading
-to a sample node.  We denote the number of samples emanating from node $x$ as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
-created from a single simulation on this GSP.
-}
-\begin{center}
-\includegraphics[width=0.48\textwidth]{figures/F2.pdf}
-\end{center}
-\caption[\gspcapone]{\sometimes{\gspcapone}}
-\label{fig:gsp1}
-\end{figure}
-%%%%%%%%%%%%
-%%%%%
-\begin{figure}
-\newcommand{\gspcaptwo}{\footnotesize
-An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
-admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $\BC_1$ in sample node~s9; and two $\BC_2$'s in sample node~s10.
-}
-\begin{center}
-\includegraphics[width=0.48\textwidth]{figures/AllThem.pdf}
-\end{center}
-\caption[\gspcaptwo]{\sometimes{\gspcaptwo}}
-\label{fig:gsp2}
-\end{figure}
-%%%%
 
@@
 Figure~\ref{fig:gsp1} shows an example GSP for the simple case
-of simulating $F_2$ individuals. The figure and its caption
-should be studied, as the following section
-uses terminology that is defined in the figure caption.
+of simulating $F_2$ individuals.
+%%%%%%%%%%%
+\begin{figure}
+\newcommand{\gspcapone}{\footnotesize
+A genomic simulation pedigree (GSP) for simulating $F_2$ individuals. See text
+for detailed explanation.
+}
+\begin{center}
+\includegraphics[width=0.48\textwidth]{figures/F2.pdf}
+\end{center}
+\caption[\gspcapone]{\sometimes{\gspcapone}}
+\label{fig:gsp1}
+\end{figure}
+%%%%%%%%%%%%
+In this figure, squares denote founder and non-founder nodes in the pedigree. There are four founders in the pedigree,
+individuals 1--4.  The  triangles above each of those founders denotes, in color
+and text, the population of origin of each of the two genomes within each founder.
+Thus, individuals~1 and~2 are founders purely from population~$A$, and~3 and~4
+are purely from population~$B$.  The pink hexagon labeled s7 is the sample node.
+It represents samples that are taken from non-founder node~7, which, from the structure
+of the pedigree, clearly represents an $F_2$ hybrid between populations $A$ and $B$.  The red numbers
+on edges above and leading into each non-founder node indicate the number of
+gametes that get segregated from the parent node (at the top of the edge) to the
+daughter node (at the bottom of the edge).
+This emphasizes that,
+in dropping genes through a GSP, in contrast
+to a traditional pedigree, each parent
+node may segregate multiple gametes to a daughter node.  See the text for
+further explanation.
+Each non-founder node $x$ will have exactly two edges coming into it from above
+(from its two parents), and we denote the numbers associated with those edges as $G^+_{x,1}$ and $G^+_{x,2}$.
+For example, node 6 has $G^+_{6,1}=G^+_{6,2} = 2$.
+A founder individual may have, at most, 2 edges coming downward out of it, while a non-founder individual may have more or less than two edges coming downward
+out of it.  Letting $e_x$ be the number of downward edges out of node $x$ to its $e_x$
+non-founder daughter nodes, we denote the number of gametes segregated
+through each edge as $G^{-}_{x, 1},\ldots,G^{-}_{x, e_x}$.  For example, at
+non-founder node~5, $e_5 = 1$ and $G^{-}_{5,1}=4$.
+Finally, the purple number adjacent to the edge leading into the
+sample node indicates how many sample individuals are drawn from individual
+node 7. Any non-founder individual node can have a maximum of one edge leading
+to a sample node.  We denote the number of samples emanating from node $x$ as $S_x$. In our example $S_7=4$ indicates that four $F_2$ individuals are
+created from a single simulation on this GSP.
 It is important to understand that, apart from the founder nodes, the individual
@@
 will create GSPs for samples up to and including second-generation backcrosses (an example appears in Figure~\ref{fig:gsp2});
+%%%%%
+\begin{figure}
+\newcommand{\gspcaptwo}{\footnotesize
+An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
+admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $\BC_1$ in sample node~s9; and two $\BC_2$'s in sample node~s10.
+}
+\begin{center}
+\includegraphics[width=0.48\textwidth]{figures/AllThem.pdf}
+\end{center}
+\caption[\gspcaptwo]{\sometimes{\gspcaptwo}}
+\label{fig:gsp2}
+\end{figure}
+%%%%
 however, \gscramble{} is designed
@@
 
-In the original paper, \citet{rizza2023limited} identified 15 $F_1$s,
+In the original paper, \citet{rizza2023limited} used 30 of the 65 markers that
+occur on different chromosome arms to identify 15 $F_1$s,
 1 $F_2$,
@@
 implemented in {\sc NewHybrids}) 
-can yield a poor estimate of the uncertainty around different inferences.
+can yield a poor estimate of the uncertainty around different inferences, and this analysis
+supports the choice of \citet{rizza2023limited} to limit markers to those
 
@@
 independent erroneously reduces the variation in the admixture 
-fractions of simulated individuals.
+fractions of simulated individuals.  In connection with this, we also stress that simply
+pruning markers to eliminate or reduce linkage disequilibrium (LD)
+will not solve the problem posed by physical linkage.
+Genomic material is inherited in chunks that are typically orders of magnitude longer than
+the typical range over which LD decays.  Thus, pairs of markers that are
+on the same chromosome, are much more likely to be co-inherited than markers that
+are not on the same chromosome, even if the two markers are far enough apart on the
+same chromosome that they are not in LD.  That said, for some downstream analyses (after
+running \gscramble{}),  LD pruning can be advantageous; for example, ADMIXTURE will run faster on a smaller, pruned data set.
 
@@
 sampled per iteration to avoid sampling of anticorrelated genotypes.
+
+As noted in the methods, the GSP used to simulate with \gscramble{} must not contain
+inbreeding loops, since the occurrence of such loops means genes would be
+sampled with replacement from the founders, potentially inducing RISPI and incorrectly
+inflating the
+perceived power for admixture fraction estimation.  One might reasonably ask whether this
+limits the application of \gscramble{} in small populations of conservation concern where 
+inbreeding occurs; we submit that it does not.  If genetic structure between two
+locales is driven by family structure (as it often may be in very small and inbred populations), that
+family structure will be reflected in samples taken from each locale---the samples within a locale 
+may be close relatives of one another.  If related individuals are used as founders in a GSP, then
+the offspring they produce will essentially be inbred, even though the GSP itself contains no
+inbreeding loops, because the founders are related.
+The inclusion of related founders in this way, then, also appropriately reflects the genetic drift that is causing allele frequency differences between the locales.
 

### 2024-12-03 data-accessibility.tex
Update author contributions
Tiny tweaks before submitting
Revisions for pub/accept

@@
+\section*{Data Accessibility and Benefit-Sharing}
+
+\subsection*{Data Accessibility Statement}
+
+A stable version of the `gscramble' R package is available on CRAN: \url{https://CRAN.R-project.org/package=gscramble}. The development version and entire revision history of
+\gscramble{} is available on GitHub: \url{https://github.com/eriqande/gscramble}, and all package documentation is available online in HTML format: \url{https://eriqande.github.io/gscramble}
+
+Data and code for the simulation analyses and steelhead-cutthroat analyses are available from \url{https://github.com/eriqande/gscramble-paper-sims-and-analyses} and are archived at \url{https://doi.org/10.5281/zenodo.14270660}.
+Data and code for the pig analyses are available from \url{https://github.com/mgdesaix/gscramble-paper-pigs} and are archived
+at \url{Get-from-Matt}.
+
+\subsection*{Benefit-sharing Statement}
+
+Benefits Generated: Benefits from this research accrue from the sharing of our data and results on public databases and the distribution of a novel and useful R package as described above.
+
 \section*{Author Contributions}
@@
 of variance in admixture fractions due to physical linkage were conducted by ECA.
-The empirical analyses using `gscramble' were conducted
-by MGD and RMG with data from TJS and by ECA with previously published data.  All authors contributed to writing and
-editing the manuscript.
-
-\section*{Data Availability Statement}
-
-\begin{itemize}
-\item Stable version of \gscramble{} available on CRAN: \url{https://CRAN.R-project.org/package=gscramble}
-\item Development version and entire revision history on GitHub: \url{https://github.com/eriqande/gscramble}
-\item Online version of all package documentation: \url{https://eriqande.github.io/gscramble}
-\item Data and code repository with simulation analyses and steelhead-cutthroat analyses: \url{https://github.com/eriqande/gscramble-paper-sims-and-analyses} (To be transferred to Dryad or Zenodo upon acceptance).
-\item Data and code repository for pig analyses: \url{https://github.com/mgdesaix/gscramble-paper-pigs} (To be transferred to Dryad or Zenodo upon acceptance)
-\end{itemize}
-\mbox{}
+The empirical analyses using `gscramble' were conducted by MGD, RMG, TJS, and ECA with previously published data.
+All authors contributed to writing and editing the manuscript.

### 2024-12-03 GitLatexDiffIt.sh

@@ -1,0 +1,26 @@
+
+
+if [ $# -ne 1 ]; then
+    echo " Syntax: GitLatexDiffIt.sh  ref "; echo
+    echo "   where ref is a hash of tag for the starting commit you want.  "; echo
+    echo "  It will then be compared to the current working copy (even if uncommitted) "
+    echo
+    echo
+    echo " Example: GitLatexDiffIt.sh toCoauthors "
+    echo
+fi
+
+
+mkdir -p diffs
+
+M="diffs/main-one-col-diff-to-$1.pdf"
+
+# get the main document (note we have to latex supplement.tex first for xr)
+git-latexdiff \
+    --bibtex \
+    --main main-one-col.tex \
+    --output  $M \
+    $1  -- && \
+    open $M;
+
+

### 2024-12-10 cover-letter/responses-to-editor-and-referees.Rmd
Revised and resubmitted version

@@
 ---
-title: "Responses to Editor and Referees"
 output: pdf_document
+---
 
----
+Benjamin Sibbett PhD  
+Executive Editor, Molecular Ecology Resources  
+4 December 2024
+
+Dear Dr Sibbett,
+
+Thank you for the opportunity to revise our work for consideration of acceptance
+in _Molecular Ecology Resources_.  We appreciate the efforts and helpful
+questions and suggestions from the associate editor and the two referees.  We have
+addressed all queries in the decision letter and are pleased with how doing so
+has improved the manuscript.
+
+On the following pages of thie PDF document, please find:
+
+- a PDF version of our responses to the decision letter (these are also pasted in 
+into the text box on manuscript central)
+- a latexdiff PDF file showing changes we made to between the original submission and our
+new revised version of the manuscript.
+
+Thank you again for considering our work.  As always, the MER editorial process has
+been a model of efficiency and clarity.  Thank you for all your efforts and leadership.
+
+Best regards,
+
+\vspace*{4em}
+
+Eric C. Anderson, PhD
+
+\newpage
+
+# Responses to Editor and Referees
 
@@
 
-_Thank you.  We address the issue of inbreeding in the revised version.  See below.
-Our replies and comments are in italic text throughout._
+_Thank you.  We address the issue of inbreeding in the revised version.  See below. Our replies and comments are in italic text throughout._
 
@@
 
-_Thank you for catching that.  They should not appear there.  That was a byproduct of using the same
-\LaTeX{} input for both a typeset-2-column version and our submission version.  We have moved the
-figure callouts to their appropriate locations in the paper._
+_Thank you for catching that.
+They should not appear there.
+That was a byproduct of using the same \LaTeX{} input for both a typeset-2-column version and our submission version.
+We have moved the figure callouts to their appropriate locations in the paper._
 
@@
 
-_No, it is not possible. Genomic material is inherited in chunks that are typically orders of magnitude longer than
-the typical range over which LD decays.  Thus, pairs of markers that are
-on the same chromosome, are much more likely to be co-inherited than markers that
-are not on the same chromosome, even if the two markers are far enough apart on the
-same chromosome that they are not in LD. We have noted this in the discussion for any other readers that may
-wonder whether LD pruning will solve the problem of physical linkage._
+_No, it is not possible.
+Genomic material is inherited in chunks that are typically orders of magnitude longer than the typical range over which LD decays.
+Thus, pairs of markers that are on the same chromosome, are much more likely to be co-inherited than markers that are not on the same chromosome, even if the two markers are far enough apart on the same chromosome that they are not in LD.
+We have noted this in the discussion for any other readers that may wonder whether LD pruning will solve the problem of physical linkage._
 
@@
 
-_This is a reasonable question and we address it now in the Discussion. In short, gscramble works fine
-for inbred populations.  Partly this is because if the founders are related (as they would be in a tiny population)
-the offspring produced by the GSP will by inbred despite the absence of inbreeding loops in the GSP._
+_This is a reasonable question and we address it now in the Discussion.
+In short, gscramble works fine for inbred populations.
+Partly this is because if the founders are related (as they would be in a tiny population) the offspring produced by the GSP will be inbred despite the absence of inbreeding loops in the GSP._
 
@@
 
-_We pulled that long caption out of the figure and include it in the text. That improves things.
-Thank you!_
+_We pulled that long caption out of the figure and include it in the text. That improves things. Thank you!_
 
@@
 
-_Yes, the user has to put those in, though we already note in the discussion that even if an
-accurate recombination map is not known, it is still possible to use one simply with
-average rates across each chromosome._
+_Yes, the user has to put those in, though we already note in the discussion that even if an accurate recombination map is not known, it is still possible to use one simply with average rates across each chromosome._
 
@@
 
-_As noted above, LD and physical linkage are quite different.  Where we now address that in the discussion, we also
-note that for applications of the data after it has been run through gscramble, like using ADMIXTURE, it can be
-advantageous to have LD-pruned data._
+_As noted above, LD and physical linkage are quite different.
+Where we now address this issuein the discussion, we also note that for analyses of simulated data produced with gscramble, i.e., the use of ADMIXTURE, it can be advantageous to conduct LD-pruning._
 
@@
 
-_The point of this simulation was to see whether the original authors were justified in limiting their data set to just the 30
-markers on separate chromosome arms. We now emphasize in the results section that this choice is justified by the simulation results._
+_The point of this simulation was to see whether the original authors were justified in limiting their data set to just the 30 markers on separate chromosome arms.
+We now emphasize in the results section that this choice is justified by the simulation results._
 
@@
 
-_This is an interesting question. We do feel that gscramble could be applicable in some
-contexts within human genetics.  The ready availability of phased dense data in humans
-would be appropriate for use with the `preserve_haplotypes = TRUE` option. So, gscramble
-could be used to assess how well admixture from different poulations could be detected.
-Given the emphasis on molecular ecology in the outlet, however, we agree with the
-reviewer that including discussion of human genetics might be outside of the scope of the
-paper._
+_This is an interesting question.
+We do feel that gscramble could be applicable in some contexts within human genetics.
+The ready availability of phased dense data in humans would be appropriate for use with the `preserve_haplotypes = TRUE` option.
+So, gscramble could be used to assess how well admixture from different poulations could be detected.
+Given the emphasis on molecular ecology in the outlet, however, we agree with the reviewer that including discussion of human genetics might be outside of the scope of the paper._
 
@@
 
-_I think it is fairly well understood from the explanations in Anderson, Waples, and Kalinowski (2008) that
-RISPI would also affect something like K-means clustering.  We considered adding more simulations to the paper, but felt that
-the paper was already quite long as it is.  We focus on ADMIXTURE in the paper because it provides
-an explicit estimate of the admixture fraction, and one of the main points of the paper is that
-the variance of admixture fractions within recent hybrids is strongly affected by physical linkage._
+_I think it is fairly well understood from the explanations in Anderson, Waples, and Kalinowski (2008) that RISPI would also affect something like K-means clustering.
+We considered adding more simulations to the paper, but felt that the paper was already quite long as it is.
+We focus on ADMIXTURE in the paper because it provides an explicit estimate of the admixture fraction, and one of the main points of the paper is that the variance of admixture fractions within recent hybrids is strongly affected by physical linkage._
 
@@
 
-_Yes, being able to make those 3 x 3 grids could make them more readable; however we do
-feel it is important to have those four magnitudes of marker numbers available, since they
-do attempt to span the range of the sorts of data sets that might be encountered in molecular
-ecology today: 100 = Fluidigm chips and amplicon sequencing; 1000 = marker capture;
-10,000 = RAD seq and small SNP chips; 100,000+ RAD seq, large SNP chips and whole genome
-sequencing. Ultimately, we chose to keep them 4 x 4 to preserve that range, and symmetry
-in the vertical dimension._
+_Yes, being able to make those 3 x 3 grids could make them more readable; however we do feel it is important to have those four magnitudes of marker numbers available, since they do attempt to span the range of the sorts of data sets that might be encountered in molecular ecology today: 100 = Fluidigm chips and amplicon sequencing; 1000 = marker capture; 10,000 = RAD seq and small SNP chips; 100,000+ RAD seq, large SNP chips and whole genome sequencing.
+Ultimately, we chose to keep them 4 x 4 to preserve that range, and symmetry in the vertical dimension._
 
@@
 4. Provide your manuscript in an editable format (not PDF!)
-
 5. Provide tables in an editable format (not images)
-
 6. Make sure all table and figure captions are present and editable

### 2024-12-10 abstract.tex
Revised and resubmitted version

@@
 simulation approach in R for creating biologically informed individual genotypes from 
-empirical data that: 1) samples alleles from populations \textit{without} replacement, 
+empirical data that: 1) samples alleles from populations \textit{without} replacement and 
 2) segregates alleles based on species-specific recombination rates. This framework

### 2024-12-10 body-text-of-paper.tex
Revised and resubmitted version

@@
 \begin{center}
-\includegraphics[width=\textwidth]{figures/permutations-plot.png}
+\includegraphics[width=\textwidth]{figures/permutations-plot.pdf}
 \end{center}
@@
 }
-\vspace*{-2.3ex}\hrule\vspace*{0.3ex}\hrule
+\vspace*{-0.3ex}\hrule\vspace*{0.3ex}\hrule
 \end{center}
@@
 maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
-are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
+are quite high ($>0.90$).  In fact, using 65 markers yields more posteriors $>0.90$ for an {\em incorrect} hybrid category
 than does using the 30 unlinked markers. So, as expected, ignoring physical linkage (e.g., as 
@@
 can yield a poor estimate of the uncertainty around different inferences, and this analysis
-supports the choice of \citet{rizza2023limited} to limit markers to those
+supports the choice of \citet{rizza2023limited} to limit markers to those on different chromosome arms during their {\sc NewHybrids} analysis.
 
@@
 same chromosome that they are not in LD.  That said, for some downstream analyses (after
-running \gscramble{}),  LD pruning can be advantageous; for example, ADMIXTURE will run faster on a smaller, pruned data set.
+running \gscramble{}),  LD pruning can be advantageous; for example, ADMIXTURE assumes independent loci and will run faster on a smaller, pruned data set.
 
@@
 
-As noted in the methods, the GSP used to simulate with \gscramble{} must not contain
+As noted in the methods, the GSP used to conduct simulations with \gscramble{} must not contain
 inbreeding loops, since the occurrence of such loops means genes would be
@@
 family structure will be reflected in samples taken from each locale---the samples within a locale 
-may be close relatives of one another.  If related individuals are used as founders in a GSP, then
-the offspring they produce will essentially be inbred, even though the GSP itself contains no
-inbreeding loops, because the founders are related.
+may be close relatives of one another.
+If related individuals are used as founders in a GSP, then the offspring they produce will essentially be inbred because the founders are related even though the GSP contains to inbreeding loops.
 The inclusion of related founders in this way, then, also appropriately reflects the genetic drift that is causing allele frequency differences between the locales.

### 2024-12-10 data-accessibility.tex
Revised and resubmitted version

@@
 Data and code for the pig analyses are available from \url{https://github.com/mgdesaix/gscramble-paper-pigs} and are archived
-at \url{Get-from-Matt}.
+at \url{https://doi.org/10.5281/zenodo.14271038}.
 
